### 2020-11-06 article_rvTH15_6.Rnw
change ref bib

\documentclass[article]{jss}

%% -- LaTeX packages and custom commands ---------------------------------------

%% recommended packages
\usepackage{thumbpdf,lmodern}
\usepackage{float,tikz}
%% another package (only for this demo article)
\usepackage{framed}
%% author pakckage
\usepackage{xcolor,caption, float}

\graphicspath{ {./imgs/} }

%% new custom commands
\newcommand{\class}[1]{`\code{#1}'}
\newcommand{\fct}[1]{\code{#1()}}

<<preliminaries, echo=FALSE, include=FALSE, cache=FALSE >>=
# options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
library("extrafont") #;library("ggplot")
library("igraph")
opts_chunk$set(prompt=TRUE, fig.pos = 'H', results='asis', highlight=FALSE)
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
# options(replace.assign=TRUE, width=90, prompt="R> ")
opts_knit$set(concordance=TRUE, self.contained=FALSE, background = "#FFFFFF")
# knitr::opts_chunk$set(fig.pos = 'H')
loadfonts(device = "win")
@

%% -- Article metainformation (author, title, ...) -----------------------------

%% - \author{} with primary affiliation
%% - \Plainauthor{} without affiliations
%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
%% - \AND starts a new line, \And does not.
\author{Thomas Huet\\UMR 5140}
%   \And Second Author\\Plus Affiliation}
\Plainauthor{Thomas Huet}

%% - \title{} in title case
%% - \Plaintitle{} without LaTeX markup (if any)
%% - \Shorttitle{} with LaTeX markup (if any), used as running title
\title{Modelling Prehistorical Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
\Plaintitle{Modelling Prehistorical Iconographic Compositions. The R package 'decorr'}
\Shorttitle{Modelling Prehistorical Iconographic Compositions}

%% - \Abstract{} almost as usual
\Abstract{
  By definition, Prehistorical societies are characterised by the absence of a writing system. During, the largest part of human history -- from far -- symbolic expressions belong to illiterate societies which express themself with rock-art paintings, pottery decorations, figurines and statuary, etc., and a lot of now disappeared carved woods, textil design, etc. At the composition level, recognition of meaningful associations of signs and reccurent patterns indicate clearly the existence of social conventions in the way to display and to read these expressions.
  We present the \pkg{decorr} \proglang{R} package which grounds concepts, methods and tools to analyse any ancient graphical systems. Our asumption is that i) any graphical system is a spatial distribution of features, and ii) these features have possibly any meaningful relationships one with another depending on their pairwise spatial proximities. To model the graphical compositions we employ concepts coming from the Graph Theory. To ensure the feasability of this type of analysis, we propose a GIS-based method for inputs and a serie of functions for data management.   
}

\Keywords{Iconography, Prehistory, Graph Theory, Graph Drawing, Spatial Analysis, \proglang{R}}
\Plainkeywords{Iconography, Prehistory, Graph Theory, Graph Drawing, Spatial Analysis, R}

\Address{
  Thomas Huet\\
  CNRS-UMR 5140 \\
  Archeologie des Societes Mediterraneennes\\
  Universite Paul Valery\\
  route de Mende\\
  Montpellier 34199, France\\
  E-mail: \email{thomashuet7@gmail.com}\\
%  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
}

\begin{document}
% \SweaveOpts{concordance=TRUE}
% \Sweave2knitr("article_rvTH14_1.Rnw")
% \SweaveOpts{concordance=TRUE}

\section[Introduction]{Introduction} \label{sec:intro}

Symbolic practices is a characteristic trait of human societies. Even discussed, such practices seems to start between 233,000 to 800,00 BC \citep{dErrico00}, covering more than 97\% of total human societies time span. Symbolism cover a large range of practices, from ochre deposit in a tomb, to menhir alignements, passing through wall fresco. This latter, what might be called "iconographical practices", probably shows the most complex and interesting testimonies of past societies symbolism. For decades, its study was linked to history of religion because commonly seen as closely linked to cultual practices and believes. Since the \textit{New Archaeology} developpement during the 60's \citep{Clarke14}, symbolic expressions start to be studied with the same formal methods (statistics, seriations, distribution maps, etc.) as any another aspect of social organisation: settlement patterns, tools \emph{chaine op\'eratoire}, susbsitence strategies, etc. \citep{Renfrew91}. But unlike many aspects of the material culture where technological requirements and technical efficiency determine the choice of the raw material and the of the object -- a flint blade for cutting, a pottery for containing, a house for living --, the function of an iconographic composition cannot be drawn directly from itself. Whether these last decades study of ancient iconography had  undergone significative improvements at the site scale -- with GIS/database statistics, paleoenvironmental restitutions, etc. -- and at the sign scale with the development of archaeological sciences -- radiocarbon dating, use-wear analysis, elemental analysis, etc. --, these improvement do not necessarly help to understand the semantic content of the iconography.

Semantics or semiotics can be defined as a system of conventional features -- called signs -- organised also in conventional manners. Until our days, formal methods to study ancient iconography semantics have been mostlty been grounded  -- explicitly or not -- on the prime principle of Saussurian linguistic: the '\emph{linearity of the signifier}' \citep{Saussure89}.  
Writing is one of the most rational semiographical system with a clear distinction between signified and signifier and the development of the signified on a linear axis. Even if we do not understand the meaning of the signs, writing can easily be modelled with Graph theory and recurent patterns can be identified. For example, the 3-letters word "\code{art}" can be modelled with three vertices (\code{a}, \code{r}, \code{t}) and two edges (one between \code{a} and \code{r}, the other between \code{r} and \code{t}). In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat}.

<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=3, fig.cap="\\label{fig:figs}concatenate of graphical units (GUs) is art">>=
# par(mar=c(0,0,0,0))
g <- graph_from_literal(a-+r, r-+t)
plot(g,
     layout = cbind(seq(-1,1,length.out = gorder(g)), 0),
     vertex.color="white",
     vertex.frame.color="white",
     vertex.size=60,
     vertex.label.cex=1.5,
     vertex.label.color="black",
     vertex.label.family="Courier New",
     edge.arrow.size=.5)
@

But, as stated, in Prehistorical the writing system does not exists. Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise GUs can be in any order.

<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:figs}Potential spatial relations between GUs.">>=
par(mar=c(0.1,0.1,0.1,0.1) )
ugs <- data.frame(name=c("a","r","t"))
lks <- data.frame(from=c("a", "r", "t"),to=c("r", "t", "a"),type=c("=","=","="))
g <- igraph::graph_from_data_frame(lks, directed=FALSE, vertices= ugs)
plot(g,
     vertex.color="white",
     vertex.frame.color="white",
     vertex.size=80,
     vertex.label.cex=1,
     vertex.label.color="black",
     vertex.label.family="Courier New"
     )
@

Because of the inherent variability of iconography, and because graphical and spatial proximities between GUs are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
\begin{itemize}
\setlength\itemsep{.05em}
  \item unexplicit groupings of graphical units -- like graphical units grouped into \emp{figures},  \emp{figures} grouped into \emp{patterns}, \emp{patterns} grouped into \emp{motives}, etc. -- with tedious number of groups
  \item consistency, proximities and relationships between these groups are often implicit and not quantified
  \item studies develop proper descriptive vocabularies, singular relationships of grouped, idosyncratic methods in a site-dependend or period-dependend scales 
\end{itemize}

These issues limit drastically the possibility to conduct cross-cultural comparisons and to draw a synthesis of humankind's symbolism at a large scale and over the long-term. 
\\
\\
In this article we present the \proglang{R} package \pkg{decorr}. Its purpose is to formalise a method based on relative neighborhood graphs to analyse any graphical content. As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities, as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}. The principal idea of our model is that any graphical system can be represented by features connected (or not) to each other depending on their spatial proximity. To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory. This package has been grounded on the seminal work of \citet{Alexander08} and its first IT implementation by \citet{Huet18a}.

\section[Model]{Graph theory Model} \label{sec:model}

Graph theory offers a conceptual framework and indices (global at the entire graph scale, local at the vertex scale) to deal with notions of networks, relationships and neighbourhoods. Graphical units (GUs) can be modelled as vertices (nodes) and their spatial relations can be modelled as edges. The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.

Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface. Indeed, for large series of graphical decorations, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates. These \emph{x} and \emph{y} coordinates, measured in pixels, are relative to the decoration figure which is open in the first place in a new GIS project without any projection system. The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis. The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive. The entire decoration image is tiled into different GUs' area of influence like a Voronoi diagram of the support where the Voronoi seeds are the GUs. Exist a link between a pair of GUs when their area of influence share a common border. A RNG -- also called a geometric graph, or a planar graph -- is constructed from GUs (nodes) and their proximity links (edges). The graph itself is as a subgraph of a Delaunay triangulation.

\begin{figure}[H]
\centering
\includegraphics{gis}
\caption{\label{fig:gis} GIS interface. A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stele (drawing: \cite{DiazGuardamino10}) with its extent (\code{xmin, xmax, ymin, ymax}); B) After the polygonisation of the GUs, including the border of the stelae, the Voronoi cells, the centroid of GUs and the links between GUs having adjacent cells (ie, sharing a border) are calculated; C) For each GUs, \emph{x} and \emph{y} are calculated; D) At least one variable, like the \code{type} of the GUs is defined in order to compute composition analysis. A simplier solution will be to create directly centroids (\code{POINTS}) on the GUs and to draw the edges manually}
\end{figure}

This model has a minimal of \emph{a priori} definitions. Those definitions only concern the intrinsic properties of GUs (type, technology, color, orientation, size, etc.) and the types of relations they share. Here, we will only consider one property for the nodes, its type (column \code{type}), and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.

\begin{itemize}
  \item \emph{normal} edges
\end{itemize}

By convention, two different GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='} and represented with a plain line. The textual notation of such an edge is \code{'-=-'}. For example: \code{1 -=- 4} means that the nodes \code{1} and \code{4} have a common border.

\begin{itemize}
  \item \emph{attribute} edges
\end{itemize}

It occurs frequently that a GU can be divided into a \emph{main unit} (eg, a character) and one or various \emph{attribute units} (eg, a helmet, male sex). Broadly speaking, for further statistical analysis, it is better to use this \emph{attribute} method than to multiply the categories of GUs. To record this information, a new type of edge, tagged with \code{'+'}, is be introduced. This type of edges is directed and, by convention, displayed with a dashed line. Its starts from the \emph{main unit} and ends with the \emph{attribute units}. The textual notation of such an edge is \code{'-+-'}. For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6}.

\begin{figure}[H] 
\centering
\includegraphics[scale=0.15]{gis_1}
\caption{\label{fig:gis1} GIS interface. The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two attributes of the GU \code{personnage} (character).}
\end{figure}

\begin{itemize}
  \item \emph{overlapping} edges
\end{itemize}

Finally, it is quite common that a graphical composition shows superimpositions between different UGs permit to distinguish different decoration phases for a single support. So, at first, the analyse must be performed on each different phases of decoration separately. This stratigraphical information (\code{A} \emph{over} \code{B}, or \code{B} \emph{under} \code{A}) helps to understand the relative chronology between GUs and must be recorded. A simple way to achieve this is to introduce the new tag \code{'>'} for the for the type of edge. This type of edges is directed. The textual notation of such an edge is \code{'->-'}. For example \code{A ->- B} means that \code{A} crosses \code{B} (ie, \code{A} overlaps \code{B} in the stratigraphical sense).

\section{The R package decorr} \label{sec:models}

The \pkg{decorr} package can be downloaded from GitHub
%
\begin{CodeChunk}
\begin{CodeInput}
R>   devtools::install_github("zoometh/iconr", build_vignettes=TRUE)
\end{CodeInput}
\end{CodeChunk}
%

\subsection{External package} \label{sec:ext_pck}

The \pkg{decorr} package imports the following packages:
\begin{itemize}
\setlength\itemsep{.1em}
  \item \pkg{magick} for image manipulation \citep{Ooms18}
  \item \pkg{igraph} for graph and network analysis \citep{Csardi06} 
  \item \pkg{rgdal}  to read \emph{shapefiles} of nodes and edges \citep{Bivand19}
  \item \pkg{grDevices} for colors and font plotting, \pkg{graphics} for graphics, \pkg{utils} and \pkg{methods} for formally defined methods and \emph{varia} methods (all combinations, etc.) \citep{R19}
\end{itemize}

\subsection{Data} \label{sec:ext_data}

A training dataset (nodes and edges coordinates, decoration images) is stored in the \code{extdata} folder of the \pkg{decorr} package

\begin{itemize}
\item The \code{imgs} dataframe
\end{itemize}

The dataframe storing the inventory of decorations is \code{imgs}. The field \code{imgs$idf} is the short name of the decoration, useful during statistical analysis. The primary key of each decoration is the concatenate of \code{imgs$site} and \code{imgs$decor}. These keys will allow joints with the other dataframes (\code{nodes} and \code{edges})\\
%
<<results='asis', echo=FALSE>>=
library("decorr")
library("igraph")
library("magick")
# par(mar=c(0,0,0,0))
imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
                   sep="\t", stringsAsFactors = FALSE)
print(xtable::xtable(imgs,
                     caption="The studied corpus, the \\code{imgs.tsv} dataframe",
                     label="Test_table",
                     size=6,
                     align=c(
                       "|p{0.01cm}|",
                       "|p{0.5cm}|",
                       "p{2.5cm}|",
                       "p{2.5cm}|",
                       "p{8.5cm}|")),
      table.placement="H",
      include.rownames=FALSE)
@
%
The \pkg{decorr} package training dataset is composed by four stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}. At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package.
%
<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=13, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
# library("magick")
pth <- system.file("extdata", package = "decorr")
imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
                   sep="\t", stringsAsFactors = FALSE)
lims <- list()
for(i in 1:nrow(imgs)){
  i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
  lbl.txt <- paste0(imgs[i,"idf"],"\n",
                    imgs[i,"site"],"\n",
                    imgs[i,"decor"],"\n",
                    imgs[i,"img"],"\n",
                    image_info(i1)$width,"*",image_info(i1)$height," px")
  i1 <- image_annotate(i1,lbl.txt,location = "northwest",
                       size = 25, color = "red")
  lims[[length(lims)+1]]<- i1
}
out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
                          image_append(c(lims[[3]],lims[[4]]))),
                        stack = TRUE)
plot(out.img)
@
%
To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load \code{nodes} and \code{edges} dataframes.
%
<<echo=TRUE>>=
nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
                    sep=";",stringsAsFactors = FALSE)
edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
                    sep=";",stringsAsFactors = FALSE)
@

\begin{itemize}
\item The \code{nodes} dataframe
\end{itemize}

It contains the required minimum variables for the analysis.

<<results='asis', echo=TRUE>>=
caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
print(xtable::xtable(head(nodes),
                     caption = caption),
      table.placement="H")
@

The primary key of the decoration is based on two fields: \code{nodes$site} and \code{nodes$decor}. The site is the current unit of analysis in Prehistory and Archaeology, but since a site can have various decorated objects, a primary key on two fields is necessary. The \code{nodes$id} is the identifier of the node. The \code{nodes$type} field is the default variable for further statistical analysis. Here, \code{nodes$type} refers to the typology of the GUs (anthropomorph, weapons, etc.). The \code{nodes$x} and \code{nodes$y} columns refer to the \emph{x} and \emph{}y coordinates of the nodes. As said, in the first place theses coordinates come from the GIS. But, in a GIS, the coordinates origin (0, 0) is the bottom-left corner and exist negative values, while this origin is top-left for any \proglang{R} matrices (rasters, grids, dataframes, etc.) with only positive values. To recover the correct the \emph{y} value of GUs nodes and edges, that is to say the \emph{y} value on the decoration image, the \pkg{decorr} calculate the absolute \emph{y} value and used the image height as a constant offset.

\begin{itemize}
\item The \code{edges} dataframe
\end{itemize}

The \code{edges} dataframe is quite similar to the \code{nodes} dataframe.

<<edgesdf, results='asis', echo=TRUE>>=
caption <- "Edges (from \\code{edges.csv} dataframe)}"
print(xtable::xtable(head(edges),
               caption=caption,
               label="edgesdf"),
      table.placement="H")
@

Fields \code{edges$site} and \code{edges$decor} are the primary key of decoration. The fields \code{edges$a} and \code{edges$b} are the equivalent to columns \emph{from} and \emph{to} in Graph theory. Even if undirected graphs will the most common models in further studies, this direction helps to distinguish between nodes. The first column \code{edges$a} is the identifier of starting node or \emph{main node} or \emph{overlapping node}. The second column \code{edges$b} is the identifier of the ending node or \emph{attribue node} or \emph{overlapped node}. The \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the starting node and the ending node. There is no need to get the coordinates of the edges, these coordinates are calculated from the \code{nodes} dataframe. For example, Table \ref{edgesdf} shows that the first edge of the \emph{Cerro Muriano 1} decoration connects the nodes \code{1} and \code{8} (respectively in column \code{edges$a} and \code{edges$b}). A way to retrieve these connected nodes' coordinates will be:

<<results='asis', echo=TRUE>>=
cm.1 <- subset(nodes, decor == "Cerro Muriano 1" & id == 1)[,c("x","y")]
cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
cat(as.numeric(cm.1),";",as.numeric(cm.8))
@

Once these three dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.

\subsection{\code{list\_dec()} function} \label{sec:funlistdec}

The \code{list\_dec()} function allows to calculate graphs for all decorations stored into \code{nodes}, \code{edges} and \code{images}. The result is a list of decoration graph. The first graph of can be plotted

<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:figs}Plot of the first graph of the list">>=
# par(mar=c(0.1,0.1,0.1,0.1))
# library("decorr")
# library("magick")
# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
#                       sep="\t", stringsAsFactors = FALSE)
# nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
#                     sep="\t",stringsAsFactors = FALSE)
# edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
#                     sep="\t",stringsAsFactors = FALSE)
lgrph <- list_dec(imgs,nodes,edges,var="type")
plot(lgrph[[1]],
     vertex.color = "orange",
     vertex.frame.color="orange",
     vertex.label.color = "black",
     vertex.size = 8,
     vertex.label.cex = 0.6,
     edge.color = "orange",
     vertex.label.family="Helvetica"
)
@

The others \pkg{decorr} package functions can be divided into:
\begin{enumerate}
 \item graphical functions
 \item single decoration functions
 \item comparisons between different decorations functions
\end{enumerate}

\subsection{Graphical functions} \label{sec:functions_gr}

The \pkg{decorr} package has three graphical functions
\begin{itemize}
\setlength\itemsep{.1em}
\item \code{labels_shadow()} function is a re-use of the \code{shadowtext()} function from the \pkg{TeachingDemos} package \citep{Snow20}.
\item \code{side_plot_nds()} and \code{side_plot_eds()} allow to plot figures side-by-side for nodes or edges comparisons
\end{itemize}

\subsection{Single decoration functions} \label{sec:functions_one}

Functions allowing to create a RNG for a single decoration are:

\begin{itemize}
\setlength\itemsep{.1em}
  \item \code{read_nds()} and \code{read_eds()} functions allow to read respectively a file of nodes and a file of edges (\code{.tsv} or \code{.shp} files)
\end{itemize}

The \code{read_nds()} function is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.

The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe. For example, the first \emph{Torrejon Rubio 1} edge, between the nodes \code{6} and \code{5}, has the starting point (xa = 366.7001, ya = -563.1358) and the ending point (xb = 490.1195, yb = -513.2428)

<<results='asis', echo=TRUE>>=
# library(decorr)
sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
print(xtable::xtable(eds.df[1,],
                     caption="first edge of the dataframe",
                     label="Test_table_1",
                     size=7),
      table.placement="H")
@
\end{figure}

\begin{itemize}
\setlength\itemsep{.1em}
  \item \code{plot_dec_grph ()} allows to plot a RNG over a decoration image
\end{itemize}

Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted, here for the \emph{Torrejon Rubio 1} decoration. The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field.

<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
# library("decorr")
par(mar=c(0,0,0,0))
sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
                   doss = system.file("extdata", package = "decorr"))
img.graph <- plot_dec_grph(nds.df = nds.df,
                           eds.df = eds.df,
                           site = sit,
                           decor = dec,
                           doss = system.file("extdata", package = "decorr"),
                           lbl.txt = "type",
                           lbl.size=1.7,
                           shw = c("nodes","edges"))
plot(image_trim(img.graph))
@

\subsection{Decoration comparison functions} \label{sec:functions_one}

Functions allowing to compare different decorations with RNG are

\begin{itemize}
\setlength\itemsep{.1em}
  \item \code{list_nds_compar()} and \code{list_eds_compar()} functions allow to compare respectively the common nodes and the common edges between two decorations
\end{itemize}

Comparisons between pairwise of decorations are first stored into a list. These comparisons are performed for nodes and/or edges. There are four (4) decorations in the default dataset, so there is $\frac{4!}{(4-2)!2!}={6}$ pairwise comparisons

<<results='asis', echo=TRUE>>=
# set wd to data folder
# setwd(system.file("extdata", package = "decorr"))
# library("decorr")
g.compar <- list_eds_compar(lgrph,"type")
df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
                                       g.compar[[2]][[1]]$decor,
                                       g.compar[[3]][[1]]$decor,
                                       g.compar[[4]][[1]]$decor,
                                       g.compar[[5]][[1]]$decor,
                                       g.compar[[6]][[1]]$decor),
                              decor.B=c(g.compar[[1]][[2]]$decor,
                                       g.compar[[2]][[2]]$decor,
                                       g.compar[[3]][[2]]$decor,
                                       g.compar[[4]][[2]]$decor,
                                       g.compar[[5]][[2]]$decor,
                                       g.compar[[6]][[2]]$decor))
print(xtable::xtable(df.edges.compar,
                     caption="Pairwise comparisons dataframe between decor.A and decor.B",
                     label="Test_table_1",
                     size=7),
      table.placement="H")
@

\begin{itemize}
\setlength\itemsep{.1em}
  \item \code{plot_nds_compar()} and \code{plot_eds_compar()} functions allow to plot and save two figures side-by-side for a decorations pairwise with, respectively, common nodes and common edges identified
\end{itemize}

The \code{plot_nds_compar()}and \code{plot_eds_compar()} functions create a \code{.png} image of two decorations plotted side-by-side with common nodes or edges identified. Functions returns also the name of the image. The common edges or nodes are displayed in red by default.
Let us choose the decorations \code{1} (\emph{Cerro Muriano 1}) and \code{4} (\emph{Zarza de Montsanchez}) and identify common edges.

<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.height=8, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
par(mar=c(0,0,0,0))
eds_compar <- plot_eds_compar(g.compar, c(1,4),
                              doss = system.file("extdata", package = "decorr"))
plot(image_trim(image_read(eds_compar)))
@

The comparison shows that \code{1} (\emph{Cerro Muriano 1}) and \code{4} (\emph{Zarza de Montsanchez}) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.

\begin{itemize}
\setlength\itemsep{.1em}
  \item \code{same_nds()} and \code{same_eds()} functions allow to repectively count matching nodes and matching edges between decoration pairwises
\end{itemize}

\code{same_nds()} and \code{same_eds()} allow to repectively count matching nodes and matching edges between decoration pairwises. The result is a square matrix between all pairwise comparisons with the number of common nodes or edges in the cells. For example, we can compute the matrix of common edges.

<<results='asis', echo=TRUE>>=
df.same_edges <- same_eds(lgrph,"type")
caption <- "Number of same edges between all decoration pairwise comparisons"
print(xtable::xtable(df.same_edges,
                     caption = caption,
                     label = "Test_table_2",
                     size = 8,
                     digits = c(0)),
      table.placement="H",
      include.rownames = TRUE)
@

For these two last exemples, the edges comparisons between the decoration \code{1} and the decoration \code{4} show that they have two (2) common edges.

\section{Illustrations} \label{sec:illustrations}

As said, since the precise location of the GUs is usualy not registred, the most commonly used method in statistical analysis on prehistorical iconography is the presence of common nodes. In order to demonstrate the first insight of a graph-based analysis of the decorations, we will compare two classifications, the first one based on the "classic" presence of common nodes, the second one based on the presence of common edges.

\begin{table}[H]
\begin{minipage}{.45\textwidth}
\centering
<<echo=FALSE,results='asis' >>=
library("xtable")
# library("MASS");library(decorr);library(xtable);library(magick)
# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
#                    sep="\t",stringsAsFactors = FALSE)
# nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
#                     sep="\t",stringsAsFactors = FALSE)
# edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
#                     sep="\t",stringsAsFactors = FALSE)
# lgrph <- list_dec(imgs,nodes,edges,"type")
df.same_edges <- same_eds(lgrph,"type")
df.same_nodes<- same_nds(lgrph,"type")
# library("xtable")
print(xtable(df.same_nodes, digits=c(0)),
      floating=FALSE,
      hline.after=NULL)
@
\captionof{table}{Common nodes table}
\end{minipage}
\begin{minipage}{.45\textwidth}
\centering
<<echo=FALSE,results='asis'>>=
print(xtable(df.same_edges, digits=c(0)),
      floating=FALSE,
      hline.after=NULL)
@
\captionof{table}{Common edges table}
\end{minipage}
\end{table}

Once the heatmap matrices calculated, the native \code{dist()} and \code{hclust()} functions \citep{R19} are calculated from the inverse matrices with the function \code{inv()} of the \pkg{matlib} package \citep{Friendly20}

\begin{figure}[H]
<<echo=TRUE, warning=FALSE>>=
library("matlib")
par(mfrow=c(1,2))
dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
dist.edges  <- dist(inv(as.matrix(df.same_edges)))
plot(hclust(dist.nodes), hang = -1, main = "common nodes")
plot(hclust(dist.edges), hang = -1, main = "common edges")
@
\end{figure}

Results of classifications show that for both common nodes and common edges, the most different decorations are \code{1} and \code{4}. These two decorations share four (4) common nodes and, as previously seen, only two (2) common edges.
In any cases decorations \code{2} and \code{3} are closer to decoration \code{4} than to decoration \code{1}, but their classifications changes depending on counting of common nodes or common edges. While decorations \code{2} and \code{4} have 7 common edges and 7 common nodes, plotting the comparisons for \code{3} and \code{4}, helps to understand the differences between the two classifications.

<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.cap="\\label{fig:figs}Decoration comparisons between 3 and 4">>=
par(mar=c(0,0,0,0))
par(mfrow=c(2,1))
g.compar <- list_nds_compar(lgrph,"type")
nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
                                  doss = system.file("extdata", package = "decorr"))
plot(image_read(nds_compar.3.4))
g.compar <- list_eds_compar(lgrph,"type")
eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
                              doss = system.file("extdata", package = "decorr"))
plot(image_read(eds_compar.3.4))
@

Decorations \code{3} and \code{4} share four (4) common GUs (\code{bouclier, epee, lance, miroir}) but  these GUs have different spatial organisations with only one common edge (\code{bouclier -=- lance}). At the opposite, decorations \code{2} and \code{4} show more properly the same compositions.

%% -- Summary/conclusions/discussion -------------------------------------------

\section{Summary and discussion} \label{sec:summary}

In this example we propose the iconographical \code{nodes$type} (character, weapon, etc.) GUs as the studied variable, but the user of the package can create and choose any other study variable: color for a painting, technique of realisation, size, etc.
\code{edges$type} can also be extended to other types than normal, attribute, overlapping.
The background is considered as homogeneous but a crack, a pit, a something can also be considered
The plasticity of Graph Theory allows to develop conventions in order to quote the different types of relations s.
Its geographical equivalent is a Thiessen polygon

\begin{table}[H]
% \begin{center}
  \centering
 \begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
 \hline
  % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
 node 1 & edge type & node 2 & (un)directed & \emph{birel} & description \\
 \hline
  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint, \code{A} and \code{B} can be contemporaneous \\
  \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} can be more recent than \code{B} \\
 % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
 \hline
\end{tabular}
\caption{Synthesis for the different types of relations between GUs}\label{tab1}
% \end{center}
\end{table}

\section*{Acknowledgments}

\begin{leftbar}
All acknowledgments (note the AE spelling) should be collected in this
unnumbered section before the references. It may contain the usual information
about funding and feedback from colleagues/reviewers/etc. Furthermore,
information such as relative contributions of the authors may be added here
(if any).
\end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{bibref}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

\newpage

\begin{appendix}

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

\begin{leftbar}
References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
formats of author-year citations and allow to include additional details (e.g.,
pages, chapters, \dots) in brackets. In case you are not familiar with these
commands see the JSS style FAQ for details.

Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
when acquiring the entries automatically from mixed online sources. However,
it is important that informations are complete and presented in a consistent
style to avoid confusions. JSS requires the following format.
\begin{itemize}
  \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
    be used in the references.
  \item Titles should be in title case.
  \item Journal titles should not be abbreviated and in title case.
  \item DOIs should be included where available.
  \item Software should be properly cited as well. For \proglang{R} packages
    \code{citation("pkgname")} typically provides a good starting point.
\end{itemize}
\end{leftbar}

\end{appendix}

%% -----------------------------------------------------------------------------

\end{document}

### 2020-11-06 article_rvTH15_6.Rnw
neww changes

@@
 %% - \AND starts a new line, \And does not.
-\author{Thomas Huet\\UMR 5140}
+\author{Thomas Huet\\UMR 5140 \And Craig Alexander \And Jose Pozo}
 %   \And Second Author\\Plus Affiliation}
-\Plainauthor{Thomas Huet}
+\Plainauthor{Thomas Huet, Craig Alexander, Jose Pozo}
 
@@
 \includegraphics[scale=0.15]{gis_1}
-\caption{\label{fig:gis1} GIS interface. The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two attributes of the GU \code{personnage} (character).}
+\caption{\label{fig:gis1} GIS interface. The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the GU \code{personnage} (character).}
 \end{figure}

### 2020-11-22 article_rvTH15_6.tex
gis_2

--- 2020-11-06 article_rvTH15_6.Rnw
@@
-\documentclass[article]{jss}
+\documentclass[article]{jss}\usepackage{knitr}
 
@@
 \newcommand{\fct}[1]{\code{#1()}}
-
-<<preliminaries, echo=FALSE, include=FALSE, cache=FALSE >>=
-# options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-library("MASS")
-library("extrafont") #;library("ggplot")
-library("igraph")
-opts_chunk$set(prompt=TRUE, fig.pos = 'H', results='asis', highlight=FALSE)
-options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-# options(replace.assign=TRUE, width=90, prompt="R> ")
-opts_knit$set(concordance=TRUE, self.contained=FALSE, background = "#FFFFFF")
-# knitr::opts_chunk$set(fig.pos = 'H')
-loadfonts(device = "win")
-@
 
@@
 }
-
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
 \begin{document}
@@
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=3, fig.cap="\\label{fig:figs}concatenate of graphical units (GUs) is art">>=
-# par(mar=c(0,0,0,0))
-g <- graph_from_literal(a-+r, r-+t)
-plot(g,
-     layout = cbind(seq(-1,1,length.out = gorder(g)), 0),
-     vertex.color="white",
-     vertex.frame.color="white",
-     vertex.size=60,
-     vertex.label.cex=1.5,
-     vertex.label.color="black",
-     vertex.label.family="Courier New",
-     edge.arrow.size=.5)
-@
+\begin{figure}[H]
+
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-1-1}
+
+}
+
+\caption{\label{fig:figs}concatenate of graphical units (GUs) is art}\label{fig:unnamed-chunk-1}
+\end{figure}
 
@@
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:figs}Potential spatial relations between GUs.">>=
-par(mar=c(0.1,0.1,0.1,0.1) )
-ugs <- data.frame(name=c("a","r","t"))
-lks <- data.frame(from=c("a", "r", "t"),to=c("r", "t", "a"),type=c("=","=","="))
-g <- igraph::graph_from_data_frame(lks, directed=FALSE, vertices= ugs)
-plot(g,
-     vertex.color="white",
-     vertex.frame.color="white",
-     vertex.size=80,
-     vertex.label.cex=1,
-     vertex.label.color="black",
-     vertex.label.family="Courier New"
-     )
-@
+\begin{figure}[H]
+
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-2-1}
+
+}
+
+\caption{\label{fig:figs}Potential spatial relations between GUs.}\label{fig:unnamed-chunk-2}
+\end{figure}
 
@@
 %
-<<results='asis', echo=FALSE>>=
-library("decorr")
-library("igraph")
-library("magick")
-# par(mar=c(0,0,0,0))
-imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-                   sep="\t", stringsAsFactors = FALSE)
-print(xtable::xtable(imgs,
-                     caption="The studied corpus, the \\code{imgs.tsv} dataframe",
-                     label="Test_table",
-                     size=6,
-                     align=c(
-                       "|p{0.01cm}|",
-                       "|p{0.5cm}|",
-                       "p{2.5cm}|",
-                       "p{2.5cm}|",
-                       "p{8.5cm}|")),
-      table.placement="H",
-      include.rownames=FALSE)
-@
+\begin{kframe}
+
+{\ttfamily\noindent\color{warningcolor}{\#\# Warning: package 'magick' was built under R version 3.6.3}}
+
+{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Linking to ImageMagick 6.9.9.14\\\#\# Enabled features: cairo, freetype, fftw, ghostscript, lcms, pango, rsvg, webp\\\#\# Disabled features: fontconfig, x11}}\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Sun Nov 22 13:33:53 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{|p{0.5cm}|p{2.5cm}|p{2.5cm}|p{8.5cm}|}
+  \hline
+idf & site & decor & img \\ 
+  \hline
+  1 & Cerro Muriano & Cerro Muriano 1 & Cerro\_Muriano.Cerro\_Muriano\_1.jpg \\ 
+    2 & Torrejon Rubio & Torrejon Rubio 1 & Torrejon\_Rubio.Torrejon\_Rubio\_1.jpg \\ 
+    3 & Brozas & Brozas & Brozas.Brozas.jpg \\ 
+    4 & Zarza de Montanchez & Zarza De Montanchez & Zarza\_de\_Montanchez.Zarza\_De\_Montanchez.jpg \\ 
+    5 & Ibahernando & Ibahernando & Ibahernando.Ibahernando.jpg \\ 
+   \hline
+\end{tabular}
+\caption{The studied corpus, the \code{imgs.tsv} dataframe} 
+\label{Test_table}
+\end{table}
+
 %
@@
 %
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=13, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
-# library("magick")
-pth <- system.file("extdata", package = "decorr")
-imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-                   sep="\t", stringsAsFactors = FALSE)
-lims <- list()
-for(i in 1:nrow(imgs)){
-  i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
-  lbl.txt <- paste0(imgs[i,"idf"],"\n",
-                    imgs[i,"site"],"\n",
-                    imgs[i,"decor"],"\n",
-                    imgs[i,"img"],"\n",
-                    image_info(i1)$width,"*",image_info(i1)$height," px")
-  i1 <- image_annotate(i1,lbl.txt,location = "northwest",
-                       size = 25, color = "red")
-  lims[[length(lims)+1]]<- i1
+\begin{kframe}
+\begin{verbatim}
+R> # library("magick")
+R> pth <- system.file("extdata", package = "decorr")
+R> imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
++                     sep="\t", stringsAsFactors = FALSE)
+R> lims <- list()
+R> for(i in 1:nrow(imgs)){
++    i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
++    lbl.txt <- paste0(imgs[i,"idf"],"\n",
++                      imgs[i,"site"],"\n",
++                      imgs[i,"decor"],"\n",
++                      imgs[i,"img"],"\n",
++                      image_info(i1)$width,"*",image_info(i1)$height," px")
++    i1 <- image_annotate(i1,lbl.txt,location = "northwest",
++                         size = 25, color = "red")
++    lims[[length(lims)+1]]<- i1
++  }
+R> out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
++                            image_append(c(lims[[3]],lims[[4]]))),
++                          stack = TRUE)
+R> plot(out.img)
+\end{verbatim}
+\end{kframe}\begin{figure}[H]
+
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-4-1}
+
 }
-out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
-                          image_append(c(lims[[3]],lims[[4]]))),
-                        stack = TRUE)
-plot(out.img)
-@
+
+\caption{\label{fig:figs}Decoration images of the training dataset}\label{fig:unnamed-chunk-4}
+\end{figure}
+
 %
@@
 %
-<<echo=TRUE>>=
-nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
-edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
-@
+\begin{kframe}
+\begin{verbatim}
+R> nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
++                      sep=";",stringsAsFactors = FALSE)
+R> edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
++                      sep=";",stringsAsFactors = FALSE)
+\end{verbatim}
+\end{kframe}
 
@@
 
-<<results='asis', echo=TRUE>>=
-caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
-print(xtable::xtable(head(nodes),
-                     caption = caption),
-      table.placement="H")
-@
+\begin{kframe}
+\begin{verbatim}
+R> caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
+R> print(xtable::xtable(head(nodes),
++                       caption = caption),
++        table.placement="H")
+\end{verbatim}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Sun Nov 22 13:33:54 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rl}
+  \hline
+ & V1 \\ 
+  \hline
+1 & site	decor	id	type	x	y \\ 
+  2 & 1	Cerro Muriano	Cerro Muriano 1	1	personnage	349.814824824793	-298.324408661478 \\ 
+  3 & 2	Cerro Muriano	Cerro Muriano 1	2	casque	349.814824824793	-243.98505280916 \\ 
+  4 & 3	Cerro Muriano	Cerro Muriano 1	3	lance	238.463685783159	-298.324408661478 \\ 
+  5 & 4	Cerro Muriano	Cerro Muriano 1	4	bouclier	446.022208956765	-381.169656108454 \\ 
+  6 & 5	Cerro Muriano	Cerro Muriano 1	5	peigne	283.004141399813	-358.008619187794 \\ 
+   \hline
+\end{tabular}
+\caption{Nodes (from \code{nodes.csv} dataframe)}} 
+\end{table}
 
@@
 
-<<edgesdf, results='asis', echo=TRUE>>=
-caption <- "Edges (from \\code{edges.csv} dataframe)}"
-print(xtable::xtable(head(edges),
-               caption=caption,
-               label="edgesdf"),
-      table.placement="H")
-@
+\begin{kframe}
+\begin{verbatim}
+R> caption <- "Edges (from \\code{edges.csv} dataframe)}"
+R> print(xtable::xtable(head(edges),
++                 caption=caption,
++                 label="edgesdf"),
++        table.placement="H")
+\end{verbatim}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Sun Nov 22 13:33:54 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rl}
+  \hline
+ & V1 \\ 
+  \hline
+1 & site	decor	a	b	type \\ 
+  2 & 1	Cerro Muriano	Cerro Muriano 1	1	4	= \\ 
+  3 & 2	Cerro Muriano	Cerro Muriano 1	1	5	= \\ 
+  4 & 3	Cerro Muriano	Cerro Muriano 1	3	5	= \\ 
+  5 & 4	Cerro Muriano	Cerro Muriano 1	1	2	+ \\ 
+  6 & 5	Cerro Muriano	Cerro Muriano 1	1	7	+ \\ 
+   \hline
+\end{tabular}
+\caption{Edges (from \code{edges.csv} dataframe)}} 
+\label{edgesdf}
+\end{table}
 
@@
 
-<<results='asis', echo=TRUE>>=
-cm.1 <- subset(nodes, decor == "Cerro Muriano 1" & id == 1)[,c("x","y")]
-cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
-cat(as.numeric(cm.1),";",as.numeric(cm.8))
-@
+\begin{kframe}
+\begin{verbatim}
+R> cm.1 <- subset(nodes, decor == "Cerro Muriano 1" & id == 1)[,c("x","y")]
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(e, x, parent.frame()): object 'decor' not found}}\begin{verbatim}
+R> cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(e, x, parent.frame()): object 'decor' not found}}\begin{verbatim}
+R> cat(as.numeric(cm.1),";",as.numeric(cm.8))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cat(as.numeric(cm.1), "{};"{}, as.numeric(cm.8)): object 'cm.1' not found}}\end{kframe}
 
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:figs}Plot of the first graph of the list">>=
-# par(mar=c(0.1,0.1,0.1,0.1))
-# library("decorr")
-# library("magick")
-# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-#                       sep="\t", stringsAsFactors = FALSE)
-# nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-lgrph <- list_dec(imgs,nodes,edges,var="type")
-plot(lgrph[[1]],
-     vertex.color = "orange",
-     vertex.frame.color="orange",
-     vertex.label.color = "black",
-     vertex.size = 8,
-     vertex.label.cex = 0.6,
-     edge.color = "orange",
-     vertex.label.family="Helvetica"
-)
-@
+\begin{kframe}
+\begin{verbatim}
+R> # par(mar=c(0.1,0.1,0.1,0.1))
+R> # library("decorr")
+R> # library("magick")
+R> # imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
+R> #                       sep="\t", stringsAsFactors = FALSE)
+R> # nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
+R> #                     sep="\t",stringsAsFactors = FALSE)
+R> # edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
+R> #                     sep="\t",stringsAsFactors = FALSE)
+R> lgrph <- list_dec(imgs,nodes,edges,var="type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in igraph::graph\_from\_data\_frame(g.edges, directed = FALSE, vertices = g.nodes): the data frame should contain at least two columns}}\begin{verbatim}
+R> plot(lgrph[[1]],
++       vertex.color = "orange",
++       vertex.frame.color="orange",
++       vertex.label.color = "black",
++       vertex.size = 8,
++       vertex.label.cex = 0.6,
++       edge.color = "orange",
++       vertex.label.family="Helvetica"
++  )
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot(lgrph[[1]], vertex.color = "{}orange"{}, vertex.frame.color = "{}orange"{}, : object 'lgrph' not found}}\end{kframe}
 
@@
 
-<<results='asis', echo=TRUE>>=
-# library(decorr)
-sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
-eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
-print(xtable::xtable(eds.df[1,],
-                     caption="first edge of the dataframe",
-                     label="Test_table_1",
-                     size=7),
-      table.placement="H")
-@
+\begin{kframe}
+\begin{verbatim}
+R> # library(decorr)
+R> sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
+R> nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
++                     doss = system.file("extdata", package = "decorr"))
+R> eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
++                     doss = system.file("extdata", package = "decorr"))
+R> print(xtable::xtable(eds.df[1,],
++                       caption="first edge of the dataframe",
++                       label="Test_table_1",
++                       size=7),
++        table.placement="H")
+\end{verbatim}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Sun Nov 22 13:33:54 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rllrrlrrrr}
+  \hline
+ & site & decor & a & b & type & xa & ya & xb & yb \\ 
+  \hline
+9 & Torrejon Rubio & Torrejon Rubio 1 &   6 &   5 & = & 366.70 & -563.14 & 490.12 & -513.24 \\ 
+   \hline
+\end{tabular}
+\caption{first edge of the dataframe} 
+\label{Test_table_1}
+\end{table}
+
 \end{figure}
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
-# library("decorr")
-par(mar=c(0,0,0,0))
-sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
-eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
-img.graph <- plot_dec_grph(nds.df = nds.df,
-                           eds.df = eds.df,
-                           site = sit,
-                           decor = dec,
-                           doss = system.file("extdata", package = "decorr"),
-                           lbl.txt = "type",
-                           lbl.size=1.7,
-                           shw = c("nodes","edges"))
-plot(image_trim(img.graph))
-@
+\begin{kframe}
+\begin{verbatim}
+R> # library("decorr")
+R> par(mar=c(0,0,0,0))
+R> sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
+R> nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
++                     doss = system.file("extdata", package = "decorr"))
+R> eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
++                     doss = system.file("extdata", package = "decorr"))
+R> img.graph <- plot_dec_grph(nds.df = nds.df,
++                             eds.df = eds.df,
++                             site = sit,
++                             decor = dec,
++                             doss = system.file("extdata", package = "decorr"),
++                             lbl.txt = "type",
++                             lbl.size=1.7,
++                             shw = c("nodes","edges"))
+R> plot(image_trim(img.graph))
+\end{verbatim}
+\end{kframe}\begin{figure}[H]
+
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-10-1}
+
+}
+
+\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:unnamed-chunk-10}
+\end{figure}
 
@@
 
-<<results='asis', echo=TRUE>>=
-# set wd to data folder
-# setwd(system.file("extdata", package = "decorr"))
-# library("decorr")
-g.compar <- list_eds_compar(lgrph,"type")
-df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
-                                       g.compar[[2]][[1]]$decor,
-                                       g.compar[[3]][[1]]$decor,
-                                       g.compar[[4]][[1]]$decor,
-                                       g.compar[[5]][[1]]$decor,
-                                       g.compar[[6]][[1]]$decor),
-                              decor.B=c(g.compar[[1]][[2]]$decor,
-                                       g.compar[[2]][[2]]$decor,
-                                       g.compar[[3]][[2]]$decor,
-                                       g.compar[[4]][[2]]$decor,
-                                       g.compar[[5]][[2]]$decor,
-                                       g.compar[[6]][[2]]$decor))
-print(xtable::xtable(df.edges.compar,
-                     caption="Pairwise comparisons dataframe between decor.A and decor.B",
-                     label="Test_table_1",
-                     size=7),
-      table.placement="H")
-@
+\begin{kframe}
+\begin{verbatim}
+R> # set wd to data folder
+R> # setwd(system.file("extdata", package = "decorr"))
+R> # library("decorr")
+R> g.compar <- list_eds_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
+R> df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
++                                         g.compar[[2]][[1]]$decor,
++                                         g.compar[[3]][[1]]$decor,
++                                         g.compar[[4]][[1]]$decor,
++                                         g.compar[[5]][[1]]$decor,
++                                         g.compar[[6]][[1]]$decor),
++                                decor.B=c(g.compar[[1]][[2]]$decor,
++                                         g.compar[[2]][[2]]$decor,
++                                         g.compar[[3]][[2]]$decor,
++                                         g.compar[[4]][[2]]$decor,
++                                         g.compar[[5]][[2]]$decor,
++                                         g.compar[[6]][[2]]$decor))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in data.frame(decor.A = c(g.compar[[1]][[1]]\$decor, g.compar[[2]][[1]]\$decor, : object 'g.compar' not found}}\begin{verbatim}
+R> print(xtable::xtable(df.edges.compar,
++                       caption="Pairwise comparisons dataframe between decor.A and decor.B",
++                       label="Test_table_1",
++                       size=7),
++        table.placement="H")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.edges.compar, caption = "{}Pairwise comparisons dataframe between decor.A and decor.B"{}, : object 'df.edges.compar' not found}}\end{kframe}
 
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.height=8, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
-par(mar=c(0,0,0,0))
-eds_compar <- plot_eds_compar(g.compar, c(1,4),
-                              doss = system.file("extdata", package = "decorr"))
-plot(image_trim(image_read(eds_compar)))
-@
+\begin{kframe}
+\begin{verbatim}
+R> par(mar=c(0,0,0,0))
+R> eds_compar <- plot_eds_compar(g.compar, c(1,4),
++                                doss = system.file("extdata", package = "decorr"))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unlist(listg, recursive = FALSE): object 'g.compar' not found}}\begin{verbatim}
+R> plot(image_trim(image_read(eds_compar)))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar): object 'eds\_compar' not found}}\end{kframe}
 
@@
 
-<<results='asis', echo=TRUE>>=
-df.same_edges <- same_eds(lgrph,"type")
-caption <- "Number of same edges between all decoration pairwise comparisons"
-print(xtable::xtable(df.same_edges,
-                     caption = caption,
-                     label = "Test_table_2",
-                     size = 8,
-                     digits = c(0)),
-      table.placement="H",
-      include.rownames = TRUE)
-@
+\begin{kframe}
+\begin{verbatim}
+R> df.same_edges <- same_eds(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
+R> caption <- "Number of same edges between all decoration pairwise comparisons"
+R> print(xtable::xtable(df.same_edges,
++                       caption = caption,
++                       label = "Test_table_2",
++                       size = 8,
++                       digits = c(0)),
++        table.placement="H",
++        include.rownames = TRUE)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.same\_edges, caption = caption, label = "{}Test\_table\_2"{}, : object 'df.same\_edges' not found}}\end{kframe}
 
@@
 \centering
-<<echo=FALSE,results='asis' >>=
-library("xtable")
-# library("MASS");library(decorr);library(xtable);library(magick)
-# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-#                    sep="\t",stringsAsFactors = FALSE)
-# nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# lgrph <- list_dec(imgs,nodes,edges,"type")
-df.same_edges <- same_eds(lgrph,"type")
-df.same_nodes<- same_nds(lgrph,"type")
-# library("xtable")
-print(xtable(df.same_nodes, digits=c(0)),
-      floating=FALSE,
-      hline.after=NULL)
-@
+\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_nodes, digits = c(0)): object 'df.same\_nodes' not found}}\end{kframe}
 \captionof{table}{Common nodes table}
@@
 \centering
-<<echo=FALSE,results='asis'>>=
-print(xtable(df.same_edges, digits=c(0)),
-      floating=FALSE,
-      hline.after=NULL)
-@
+\begin{kframe}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_edges, digits = c(0)): object 'df.same\_edges' not found}}\end{kframe}
 \captionof{table}{Common edges table}
@@
 \begin{figure}[H]
-<<echo=TRUE, warning=FALSE>>=
-library("matlib")
-par(mfrow=c(1,2))
-dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
-dist.edges  <- dist(inv(as.matrix(df.same_edges)))
-plot(hclust(dist.nodes), hang = -1, main = "common nodes")
-plot(hclust(dist.edges), hang = -1, main = "common edges")
-@
+\begin{kframe}
+\begin{verbatim}
+R> library("matlib")
+R> par(mfrow=c(1,2))
+R> dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_nodes): object 'df.same\_nodes' not found}}\begin{verbatim}
+R> dist.edges  <- dist(inv(as.matrix(df.same_edges)))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_edges): object 'df.same\_edges' not found}}\begin{verbatim}
+R> plot(hclust(dist.nodes), hang = -1, main = "common nodes")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.nodes): object 'dist.nodes' not found}}\begin{verbatim}
+R> plot(hclust(dist.edges), hang = -1, main = "common edges")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.edges): object 'dist.edges' not found}}\end{kframe}
 \end{figure}
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.cap="\\label{fig:figs}Decoration comparisons between 3 and 4">>=
-par(mar=c(0,0,0,0))
-par(mfrow=c(2,1))
-g.compar <- list_nds_compar(lgrph,"type")
-nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
-                                  doss = system.file("extdata", package = "decorr"))
-plot(image_read(nds_compar.3.4))
-g.compar <- list_eds_compar(lgrph,"type")
-eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
-                              doss = system.file("extdata", package = "decorr"))
-plot(image_read(eds_compar.3.4))
-@
+\begin{kframe}
+\begin{verbatim}
+R> par(mar=c(0,0,0,0))
+R> par(mfrow=c(2,1))
+R> g.compar <- list_nds_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
+R> nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
++                                    doss = system.file("extdata", package = "decorr"))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_nds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : object 'g.compar' not found}}\begin{verbatim}
+R> plot(image_read(nds_compar.3.4))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(nds\_compar.3.4): object 'nds\_compar.3.4' not found}}\begin{verbatim}
+R> g.compar <- list_eds_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
+R> eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
++                                doss = system.file("extdata", package = "decorr"))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unlist(listg, recursive = FALSE): object 'g.compar' not found}}\begin{verbatim}
+R> plot(image_read(eds_compar.3.4))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar.3.4): object 'eds\_compar.3.4' not found}}\end{kframe}
 

### 2020-11-22 article_rvTH15_6.Rnw
run

@@
 \centering
-\includegraphics[scale=0.15]{gis_1}
-\caption{\label{fig:gis1} GIS interface. The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the GU \code{personnage} (character).}
+\includegraphics[scale=0.1]{gis_1}
+\caption{\label{fig:gis1} GIS interface, Cerro Muriano 1 decoration.
+The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the GU \code{personnage} (anthropomorph).}
 \end{figure}
@@
 
+\begin{figure}[H] 
+\centering
+\includegraphics[scale=0.1]{gis_2}
+\caption{\label{fig:gis1} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
+\end{figure}
+
 \section{The R package decorr} \label{sec:models}
 
+<<results='asis', include=FALSE, echo=FALSE, warning=FALSE>>=
+library("decorr")
+library("igraph")
+library("magick")
+@
+%
 The \pkg{decorr} package can be downloaded from GitHub
@@
 %
-<<results='asis', echo=FALSE>>=
-library("decorr")
-library("igraph")
-library("magick")
-# par(mar=c(0,0,0,0))
+<<results='asis', echo=FALSE, warning=FALSE>>=
+# library("decorr")
+# library("igraph")
+# library("magick")
+# # par(mar=c(0,0,0,0))
 imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
@@
 %
-The \pkg{decorr} package training dataset is composed by four stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
+The \pkg{decorr} package training dataset is composed by five stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
 At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package.
@@
 out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
-                          image_append(c(lims[[3]],lims[[4]]))),
+                          image_append(c(lims[[3]],lims[[4]])),
+                          image_append(lims[[5]])),
                         stack = TRUE)

### 2020-11-22 article_rvTH15_6.tex
run

@@
 
+\begin{figure}[H] 
+\centering
+\includegraphics[scale=0.1]{gis_2}
+\caption{\label{fig:gis1} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
+\end{figure}
+
 \section{The R package decorr} \label{sec:models}
 
+%
 The \pkg{decorr} package can be downloaded from GitHub
@@
 %
-\begin{kframe}
-
-{\ttfamily\noindent\color{warningcolor}{\#\# Warning: package 'magick' was built under R version 3.6.3}}
-
-{\ttfamily\noindent\itshape\color{messagecolor}{\#\# Linking to ImageMagick 6.9.9.14\\\#\# Enabled features: cairo, freetype, fftw, ghostscript, lcms, pango, rsvg, webp\\\#\# Disabled features: fontconfig, x11}}\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 13:33:53 2020
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Sun Nov 22 14:29:55 2020
 \begin{table}[H]
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-4-1}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-5-1}
 
@@
 
-\caption{\label{fig:figs}Decoration images of the training dataset}\label{fig:unnamed-chunk-4}
+\caption{\label{fig:figs}Decoration images of the training dataset}\label{fig:unnamed-chunk-5}
 \end{figure}
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 13:33:54 2020
+% Sun Nov 22 14:29:56 2020
 \begin{table}[H]
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-10-1}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-11-1}
 
@@
 
-\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:unnamed-chunk-10}
+\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:unnamed-chunk-11}
 \end{figure}

### 2020-11-24 article_rvTH15_6.Rnw
article

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Modelling Prehistorical Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
-\Plaintitle{Modelling Prehistorical Iconographic Compositions. The R package 'decorr'}
-\Shorttitle{Modelling Prehistorical Iconographic Compositions}
+\title{Modelling Prehistoric Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
+\Plaintitle{Modelling Prehistoric Iconographic Compositions. The R package 'decorr'}
+\Shorttitle{Modelling Prehistoric Iconographic Compositions}
 
@@
 \citep{Renfrew91}.
-But unlike many aspects of the material culture where technological requirements and technical efficiency determine the choice of the raw material and the of the object -- a flint blade for cutting, a pottery for containing, a house for living --, the function of an iconographic composition cannot be drawn directly from itself.
+But unlike many aspects of the material culture where technological requirements and technical efficiency determine the choice of the raw material and the shape of the object -- a flint blade for cutting, a pottery for containing, a house for living --, the function of an iconographic composition cannot be drawn directly from itself.
 Whether these last decades study of ancient iconography had  undergone significative improvements at the site scale -- with GIS/database statistics, paleoenvironmental restitutions, etc.
@@
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=3, fig.cap="\\label{fig:figs}concatenate of graphical units (GUs) is art">>=
+<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=3, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 # par(mar=c(0,0,0,0))
@@
 But, as stated, in Prehistorical the writing system does not exists.
-Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise GUs can be in any order.
+Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order.
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:figs}Potential spatial relations between GUs.">>=
+<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:figs}Potential spatial relations between graphical units">>=
 par(mar=c(0.1,0.1,0.1,0.1) )
@@
 
-Because of the inherent variability of iconography, and because graphical and spatial proximities between GUs are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
+Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
 \begin{itemize}
@@
   \item consistency, proximities and relationships between these groups are often implicit and not quantified
-  \item studies develop proper descriptive vocabularies, singular relationships of grouped, idosyncratic methods in a site-dependend or period-dependend scales 
+  \item studies develop proper descriptive vocabularies, singular relationships of grouping, idosyncratic methods at site-dependend or period-dependend scales 
 \end{itemize}
@@
 In this article we present the \proglang{R} package \pkg{decorr}.
-Its purpose is to formalise a method based on relative neighborhood graphs to analyse any graphical content.
-As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities, as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}.
-The principal idea of our model is that any graphical system can be represented by features connected (or not) to each other depending on their spatial proximity.
+Its purpose is to formalise a method based on relative neighborhood graphs (RNG) to analyse any graphical content (here: decorations).
+As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities; as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}.
+And, because any graphical system is composed by features (mostly shapes) displayed spatialy, these systems can be represented by features connected (or not) to each other depending on their spatial proximity.
 To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory.
@@
 Graph theory offers a conceptual framework and indices (global at the entire graph scale, local at the vertex scale) to deal with notions of networks, relationships and neighbourhoods.
-Graphical units (GUs) can be modelled as vertices (nodes) and their spatial relations can be modelled as edges.
+Graphical units (GUs) can be modelled as vertices (nodes) separated one with another by undecorated areas.
+Their spatial relations can be modelled as edges.
 The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.
@@
 Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
-Indeed, for large series of graphical decorations, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
+Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
 These \emph{x} and \emph{y} coordinates, measured in pixels, are relative to the decoration figure which is open in the first place in a new GIS project without any projection system.
@@
 Exist a link between a pair of GUs when their area of influence share a common border.
-A RNG -- also called a geometric graph, or a planar graph -- is constructed from GUs (nodes) and their proximity links (edges).
+A neighborhood graphs (RNG) -- also called a geometric graph, or a planar graph -- is constructed from GUs (nodes) and their proximity links (edges).
 The graph itself is as a subgraph of a Delaunay triangulation.
+The entire graphical content of the decoration is mapped with the RNG, so a decoration graph is a one component graph.
 
@@
 \caption{\label{fig:gis} GIS interface.
-A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stele (drawing: \cite{DiazGuardamino10}) with its extent (\code{xmin, xmax, ymin, ymax}); B) After the polygonisation of the GUs, including the border of the stelae, the Voronoi cells, the centroid of GUs and the links between GUs having adjacent cells (ie, sharing a border) are calculated; C) For each GUs, \emph{x} and \emph{y} are calculated; D) At least one variable, like the \code{type} of the GUs is defined in order to compute composition analysis.
+A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stele (drawing: \cite{DiazGuardamino10}) with its extent (\code{xmin, xmax, ymin, ymax}); B) After the polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, the Voronoi cells, the centroid of GUs and the links between GUs having adjacent cells (ie, sharing a border) are calculated; C) For each GUs, \emph{x} and \emph{y} are calculated; D) At least one variable, like the \code{type} of the GUs (for example 'anthropomorph', 'horned helmet') is defined in order to compute composition analysis.
 A simplier solution will be to create directly centroids (\code{POINTS}) on the GUs and to draw the edges manually}
@@
 This model has a minimal of \emph{a priori} definitions.
-Those definitions only concern the intrinsic properties of GUs (type, technology, color, orientation, size, etc.) and the types of relations they share.
-Here, we will only consider one property for the nodes, its type (column \code{type}), and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
+Those definitions only concern the intrinsic variables of GUs (\code{type}, \code{technologie}, \code{color}, \code{size}, etc.) and the types of relations they share.
+Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
 
@@
 
-By convention, two different GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='} and represented with a plain line.
+By convention, two contemporaneous GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='}.
+This proximity is represented with a plain line.
 The textual notation of such an edge is \code{'-=-'}.
@@
 This type of edges is directed and, by convention, displayed with a dashed line.
-Its starts from the \emph{main unit} and ends with the \emph{attribute units}.
+Its starts from the \emph{main unit} and ends with the \emph{attribute unit}.
 The textual notation of such an edge is \code{'-+-'}.
@@
 \end{figure}
+
+\begin{table}[H]
+% \begin{center}
+  \centering
+ \begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
+ \hline
+  % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
+ node 1 & edge type & node 2 & (un)directed & \emph{birel} & description \\
+ \hline
+  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint, \code{A} and \code{B} can be contemporaneous \\
+  \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
+  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} can be more recent than \code{B} \\
+ % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
+ \hline
+\end{tabular}
+\caption{Synthesis for the different types of relations between GUs}\label{tab1}
+% \end{center}
+\end{table}
 
@@
 The result is a list of decoration graph.
-The first graph of can be plotted
+The first graph of the list can be plotted
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:figs}Plot of the first graph of the list">>=
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:figs}Plot of the first graph of the list">>=
 # par(mar=c(0.1,0.1,0.1,0.1))
@@
 The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe.
-For example, the first \emph{Torrejon Rubio 1} edge, between the nodes \code{6} and \code{5}, has the starting point (xa = 366.7001, ya = -563.1358) and the ending point (xb = 490.1195, yb = -513.2428)
+For example, the first \emph{Torrejon Rubio 1} edge, between the nodes \code{6} and \code{5}, has the starting point (\code{xa} = 366.7001, \code{ya} = -563.1358) and the ending point (\code{xb} = 490.1195, \code{yb} = -513.2428)
 
@@
 \setlength\itemsep{.1em}
-  \item \code{plot_dec_grph ()} allows to plot a RNG over a decoration image
+  \item \code{plot_dec_grph()} allows to plot a RNG over a decoration image
 \end{itemize}
@@
                            lbl.txt = "type",
-                           lbl.size=1.7,
+                           lbl.size=2,
                            shw = c("nodes","edges"))
@@
 These comparisons are performed for nodes and/or edges.
-There are four (4) decorations in the default dataset, so there is $\frac{4!}{(4-2)!2!}={6}$ pairwise comparisons
+There are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ pairwise comparisons
 
@@
                                        g.compar[[5]][[1]]$decor,
-                                       g.compar[[6]][[1]]$decor),
+                                       g.compar[[6]][[1]]$decor,
+                                       g.compar[[7]][[1]]$decor,
+                                       g.compar[[8]][[1]]$decor,
+                                       g.compar[[9]][[1]]$decor,
+                                       g.compar[[10]][[1]]$decor),
                               decor.B=c(g.compar[[1]][[2]]$decor,
@@
                                        g.compar[[5]][[2]]$decor,
-                                       g.compar[[6]][[2]]$decor))
+                                       g.compar[[6]][[2]]$decor,
+                                       g.compar[[7]][[2]]$decor,
+                                       g.compar[[8]][[2]]$decor,
+                                       g.compar[[9]][[2]]$decor,
+                                       g.compar[[10]][[2]]$decor))
 print(xtable::xtable(df.edges.compar,
-                     caption="Pairwise comparisons dataframe between decor.A and decor.B",
+                     caption="Pairwise comparisons dataframe between decor A and decor B",
                      label="Test_table_1",
@@
 
-The \code{plot_nds_compar()}and \code{plot_eds_compar()} functions create a \code{.png} image of two decorations plotted side-by-side with common nodes or edges identified.
-Functions returns also the name of the image.
+The \code{plot_nds_compar()}and \code{plot_eds_compar()} functions create a \code{.png} image of two or more decorations plotted side-by-side with common nodes or edges identified and returns the names of the images.
 The common edges or nodes are displayed in red by default.
@@
 
-\begin{table}[H]
-% \begin{center}
-  \centering
- \begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
- \hline
-  % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
- node 1 & edge type & node 2 & (un)directed & \emph{birel} & description \\
- \hline
-  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint, \code{A} and \code{B} can be contemporaneous \\
-  \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
-  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} can be more recent than \code{B} \\
- % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
- \hline
-\end{tabular}
-\caption{Synthesis for the different types of relations between GUs}\label{tab1}
-% \end{center}
-\end{table}
-
 \section*{Acknowledgments}

### 2020-11-26 article_rvTH15_6.Rnw
run

@@
 \newcommand{\fct}[1]{\code{#1()}}
-
+% \setlist[itemize]{noitemsep, topsep=0pt}
 <<preliminaries, echo=FALSE, include=FALSE, cache=FALSE >>=
@@
 
-\Keywords{Iconography, Prehistory, Graph Theory, Graph Drawing, Spatial Analysis, \proglang{R}}
-\Plainkeywords{Iconography, Prehistory, Graph Theory, Graph Drawing, Spatial Analysis, R}
+\Keywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (\emph{birel}), \proglang{R}}
+\Plainkeywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (birel), R}
 
@@
 In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat}.
-
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=3, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+\smallbreak
+<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=5, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 # par(mar=c(0,0,0,0))
@@
 And, because any graphical system is composed by features (mostly shapes) displayed spatialy, these systems can be represented by features connected (or not) to each other depending on their spatial proximity.
-To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory.
+To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory, GIS and binary topological relationships (\emph{birel}) modeling.
 This package has been grounded on the seminal work of \citet{Alexander08} and its first IT implementation by \citet{Huet18a}.
@@
 Their spatial relations can be modelled as edges.
-The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.
-
+The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.\\
+% \smallbreak
 Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
 Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
-These \emph{x} and \emph{y} coordinates, measured in pixels, are relative to the decoration figure which is open in the first place in a new GIS project without any projection system.
-The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis.
+The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis (ie, all the graphical content).
 The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive.
-The entire decoration image is tiled into different GUs' area of influence like a Voronoi diagram of the support where the Voronoi seeds are the GUs.
-Exist a link between a pair of GUs when their area of influence share a common border.
-A neighborhood graphs (RNG) -- also called a geometric graph, or a planar graph -- is constructed from GUs (nodes) and their proximity links (edges).
-The graph itself is as a subgraph of a Delaunay triangulation.
-The entire graphical content of the decoration is mapped with the RNG, so a decoration graph is a one component graph.
+The polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, allows to calculated their Voronoi seeds (ie, their centroid) and Voronoi cells (ie, their area of influence).
+So, the entire decoration image is tiled and condidered as a Voronoi diagram.
+Centroids of the GUs (\code{POINTS}) are calculated and considered as the graph nodes, their locations (\emph{x} and \emph{y} coordinates, measured in pixels) are relative to each decoration.
+Exist edges (\code{LINES}) between pairs of GUs when their Voronoi cells share a common border (\emph{birel}: 'touches').\\
+The entire graphical content of the decoration is mapped with a neighborhood graphs (RNG) -- ie geometric graph, or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
+So a decoration graph is a one component graph which is the dual graph of the Voronoi diagram (Fig.
+\ref{fig:gis}).
 
+% \begin{figure}
 \begin{figure}[H]
@@
 \caption{\label{fig:gis} GIS interface.
-A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stele (drawing: \cite{DiazGuardamino10}) with its extent (\code{xmin, xmax, ymin, ymax}); B) After the polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, the Voronoi cells, the centroid of GUs and the links between GUs having adjacent cells (ie, sharing a border) are calculated; C) For each GUs, \emph{x} and \emph{y} are calculated; D) At least one variable, like the \code{type} of the GUs (for example 'anthropomorph', 'horned helmet') is defined in order to compute composition analysis.
-A simplier solution will be to create directly centroids (\code{POINTS}) on the GUs and to draw the edges manually}
+A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonsiation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
 \end{figure}
+
+Whether this theoritical process include image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
+\smallbreak
 
@@
 Those definitions only concern the intrinsic variables of GUs (\code{type}, \code{technologie}, \code{color}, \code{size}, etc.) and the types of relations they share.
+At least one variable, like the \code{type} of the GUs ('anthropomorph', 'horned helmet', etc.), is defined in order to compute composition analysis.
 Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
@@
 The textual notation of such an edge is \code{'-+-'}.
-For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6}.
+For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6} (Fig.
+\ref{fig:gis_1}).
 
@@
 \centering
-\includegraphics[scale=0.1]{gis_1}
+\includegraphics[scale=0.05]{gis_1}
 \caption{\label{fig:gis1} GIS interface, Cerro Muriano 1 decoration.
@@
 The textual notation of such an edge is \code{'->-'}.
-For example \code{A ->- B} means that \code{A} crosses \code{B} (ie, \code{A} overlaps \code{B} in the stratigraphical sense).
+For example \code{A ->- B} means that \code{A} crosses \code{B} (ie, \code{A} overlaps \code{B} in the stratigraphical sense) (Fig.
+\ref{fig:gis_2}).
 
@@
 \centering
-\includegraphics[scale=0.1]{gis_2}
+\includegraphics[scale=0.05]{gis_2}
 \caption{\label{fig:gis1} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
 \end{figure}
+
+These different kind of relationships are resumed with their birel equivalencies in the following table (Tab. \ref{table:tedges})
 
@@
   % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
- node 1 & edge type & node 2 & (un)directed & \emph{birel} & description \\
+ GU.1 & edge type & GU.2 & (un)directed & \emph{birel} & description \\
  \hline
-  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint, \code{A} and \code{B} can be contemporaneous \\
+  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint \code{A} and \code{B} contemporaneous \\
   \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
-  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} can be more recent than \code{B} \\
+  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} is more recent than \code{B} \\
  % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
@@
 % \end{center}
+\label{table:tedges}
 \end{table}
@@
 
-The dataframe storing the inventory of decorations is \code{imgs}.
+The inventory of decorations is stored in the \code{imgs} dataframe.
 The field \code{imgs$idf} is the short name of the decoration, useful during statistical analysis.
@@
 The \pkg{decorr} package training dataset is composed by five stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
-At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package.
+\\
+At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:timg})
 %
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=13, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
+<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=18, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
 # library("magick")
@@
 }
-out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
-                          image_append(c(lims[[3]],lims[[4]])),
-                          image_append(lims[[5]])),
+out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]],lims[[3]])),
+                          image_append(c(lims[[4]],lims[[5]]))),
                         stack = TRUE)
@@
 
-It contains the required minimum variables for the analysis.
+It contains the required minimum variables for the analysis (Fig. \ref{fig:tnds}).
 
-<<results='asis', echo=TRUE>>=
+<<tnds, results='asis', echo=TRUE>>=
 caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
@@
 
-The primary key of the decoration is based on two fields: \code{nodes$site} and \code{nodes$decor}.
-The site is the current unit of analysis in Prehistory and Archaeology, but since a site can have various decorated objects, a primary key on two fields is necessary.
+Since a prehistorical site can have various decorated objects, the primary key of the decoration is based on two fields: \code{nodes$site} and \code{nodes$decor}.
 The \code{nodes$id} is the identifier of the node.
@@
 But, in a GIS, the coordinates origin (0, 0) is the bottom-left corner and exist negative values, while this origin is top-left for any \proglang{R} matrices (rasters, grids, dataframes, etc.) with only positive values.
-To recover the correct the \emph{y} value of GUs nodes and edges, that is to say the \emph{y} value on the decoration image, the \pkg{decorr} calculate the absolute \emph{y} value and used the image height as a constant offset.
+To recover the local \emph{y} value of GUs nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
 
@@
 
-The \code{edges} dataframe is quite similar to the \code{nodes} dataframe.
+The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Fig. \ref{fig:teds}).
 
-<<edgesdf, results='asis', echo=TRUE>>=
+<<teds, results='asis', echo=TRUE>>=
 caption <- "Edges (from \\code{edges.csv} dataframe)}"
@@
 The first column \code{edges$a} is the identifier of starting node or \emph{main node} or \emph{overlapping node}.
-The second column \code{edges$b} is the identifier of the ending node or \emph{attribue node} or \emph{overlapped node}.
+The second column, \code{edges$b}, is the identifier of the ending node or \emph{attribue node} or \emph{overlapped node}.
 The \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the starting node and the ending node.
-There is no need to get the coordinates of the edges, these coordinates are calculated from the \code{nodes} dataframe.
-For example, Table \ref{edgesdf} shows that the first edge of the \emph{Cerro Muriano 1} decoration connects the nodes \code{1} and \code{8} (respectively in column \code{edges$a} and \code{edges$b}).
+There is no need to record the coordinates of the edges, these coordinates are calculated from the \code{nodes} dataframe.
+For example, Table \ref{teds} shows that the Cerro Muriano 1 decoration's first edge connects the nodes \code{1} and \code{8} (respectively in column \code{edges$a} and \code{edges$b}).
 A way to retrieve these connected nodes' coordinates will be:
@@
 cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
-cat(as.numeric(cm.1),";",as.numeric(cm.8))
+paste(as.numeric(cm.1),";",as.numeric(cm.8))
 @
 
-Once these three dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.
+Once these dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.
 
@@
 The result is a list of decoration graph.
-The first graph of the list can be plotted
+The first graph of the list can be plotted (Fig.
+\ref{fig:igraph})
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:figs}Plot of the first graph of the list">>=
+<<igraph, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:figs}R interface. Plot of the first graph of the list">>=
 # par(mar=c(0.1,0.1,0.1,0.1))
@@
 #                       sep="\t", stringsAsFactors = FALSE)
-# nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
+# nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
 #                     sep="\t",stringsAsFactors = FALSE)
-# edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
+# edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
 #                     sep="\t",stringsAsFactors = FALSE)
@@
 The \code{read_nds()} function is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
+For example, for the Torrejon Rubio 1 decoration, the 6th GUs is a sword (\code{epee}) (Table \ref{nd}).
 
-The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe.
-For example, the first \emph{Torrejon Rubio 1} edge, between the nodes \code{6} and \code{5}, has the starting point (\code{xa} = 366.7001, \code{ya} = -563.1358) and the ending point (\code{xb} = 490.1195, \code{yb} = -513.2428)
-
-<<results='asis', echo=TRUE>>=
+<<nd, results='asis', echo=TRUE, fig.cap="Sixth node of the Torrejon Rubio 1 decoration, the sword">>=
 # library(decorr)
 sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
+nds.df <- read_nds(site = sit, decor = dec, dev = ".shp",
                    doss = system.file("extdata", package = "decorr"))
+print(xtable::xtable(nds.df[6,],
+                     caption="Sixth node (GU 6) of the Torrejon Rubio 1 decoration",
+                     label="nd",
+                     size=7),
+      table.placement="H")
+@
+
+The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe (Table \ref{ed}).
+
+<<ed, results='asis', echo=TRUE, fig.cap="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)">>=
+# library(decorr)
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
@@
 print(xtable::xtable(eds.df[1,],
-                     caption="first edge of the dataframe",
-                     label="Test_table_1",
+                     caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
+                     label="eds.df",
                      size=7),
@@
 @
-\end{figure}
+% \end{figure}
 
@@
 
-Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted, here for the \emph{Torrejon Rubio 1} decoration.
-The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field.
+Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted.
+The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field (Figure \ref{tr_st}).
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
+<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
 # library("decorr")
@@
 
-<<results='asis', echo=TRUE>>=
-# set wd to data folder
-# setwd(system.file("extdata", package = "decorr"))
-# library("decorr")
-g.compar <- list_eds_compar(lgrph,"type")
-df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
-                                       g.compar[[2]][[1]]$decor,
-                                       g.compar[[3]][[1]]$decor,
-                                       g.compar[[4]][[1]]$decor,
-                                       g.compar[[5]][[1]]$decor,
-                                       g.compar[[6]][[1]]$decor,
-                                       g.compar[[7]][[1]]$decor,
-                                       g.compar[[8]][[1]]$decor,
-                                       g.compar[[9]][[1]]$decor,
-                                       g.compar[[10]][[1]]$decor),
-                              decor.B=c(g.compar[[1]][[2]]$decor,
-                                       g.compar[[2]][[2]]$decor,
-                                       g.compar[[3]][[2]]$decor,
-                                       g.compar[[4]][[2]]$decor,
-                                       g.compar[[5]][[2]]$decor,
-                                       g.compar[[6]][[2]]$decor,
-                                       g.compar[[7]][[2]]$decor,
-                                       g.compar[[8]][[2]]$decor,
-                                       g.compar[[9]][[2]]$decor,
-                                       g.compar[[10]][[2]]$decor))
-print(xtable::xtable(df.edges.compar,
-                     caption="Pairwise comparisons dataframe between decor A and decor B",
-                     label="Test_table_1",
-                     size=7),
-      table.placement="H")
-@
+% <<results='asis', echo=TRUE>>=
+% # set wd to data folder
+% # setwd(system.file("extdata", package = "decorr"))
+% # library("decorr")
+% g.compar <- list_eds_compar(lgrph,"type")
+% df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
+%                                        g.compar[[2]][[1]]$decor,
+%                                        g.compar[[3]][[1]]$decor,
+%                                        g.compar[[4]][[1]]$decor,
+%                                        g.compar[[5]][[1]]$decor,
+%                                        g.compar[[6]][[1]]$decor,
+%                                        g.compar[[7]][[1]]$decor,
+%                                        g.compar[[8]][[1]]$decor,
+%                                        g.compar[[9]][[1]]$decor,
+%                                        g.compar[[10]][[1]]$decor),
+%                               decor.B=c(g.compar[[1]][[2]]$decor,
+%                                        g.compar[[2]][[2]]$decor,
+%                                        g.compar[[3]][[2]]$decor,
+%                                        g.compar[[4]][[2]]$decor,
+%                                        g.compar[[5]][[2]]$decor,
+%                                        g.compar[[6]][[2]]$decor,
+%                                        g.compar[[7]][[2]]$decor,
+%                                        g.compar[[8]][[2]]$decor,
+%                                        g.compar[[9]][[2]]$decor,
+%                                        g.compar[[10]][[2]]$decor))
+% print(xtable::xtable(df.edges.compar,
+%                      caption="Pairwise comparisons dataframe between decor A and decor B",
+%                      label="Test_table_1",
+%                      size=7),
+%       table.placement="H")
+% @
 
@@
 The common edges or nodes are displayed in red by default.
-Let us choose the decorations \code{1} (\emph{Cerro Muriano 1}) and \code{4} (\emph{Zarza de Montsanchez}) and identify common edges.
+Let us choose the decorations \code{1} (Cerro Muriano 1) and \code{4} (Zarza de Montsanchez) and identify common edges.
 
@@
                               doss = system.file("extdata", package = "decorr"))
-plot(image_trim(image_read(eds_compar)))
+plot(magick::image_trim(magick::image_read(eds_compar)))
 @
 
-The comparison shows that \code{1} (\emph{Cerro Muriano 1}) and \code{4} (\emph{Zarza de Montsanchez}) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.
+The comparison shows that \code{1} (Cerro Muriano 1) and \code{4} (Zarza de Montsanchez) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.
 

### 2020-12-01 article_rvTH15_6.tex
up

@@
 \newcommand{\fct}[1]{\code{#1()}}
+% \setlist[itemize]{noitemsep, topsep=0pt}
 
@@
 In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat}.
-
+\smallbreak
 \begin{figure}[H]
@@
 
-\caption{\label{fig:figs}concatenate of graphical units (GUs) is art}\label{fig:unnamed-chunk-1}
+\caption{\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 But, as stated, in Prehistorical the writing system does not exists.
-Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise GUs can be in any order.
+Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order.
 
@@
 
-\caption{\label{fig:figs}Potential spatial relations between GUs.}\label{fig:unnamed-chunk-2}
+\caption{\label{fig:figs}Potential spatial relations between graphical units}\label{fig:unnamed-chunk-2}
 \end{figure}
 
-Because of the inherent variability of iconography, and because graphical and spatial proximities between GUs are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
+Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
 \begin{itemize}
@@
 In this article we present the \proglang{R} package \pkg{decorr}.
-Its purpose is to formalise a method based on relative neighborhood graphs to analyse any graphical content.
-As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities, as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}.
-The principal idea of our model is that any graphical system can be represented by features connected (or not) to each other depending on their spatial proximity.
-To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory.
+Its purpose is to formalise a method based on relative neighborhood graphs (RNG) to analyse any graphical content (here: decorations).
+As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities; as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}.
+And, because any graphical system is composed by features (mostly shapes) displayed spatialy, these systems can be represented by features connected (or not) to each other depending on their spatial proximity.
+To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory, GIS and binary topological relationships (\emph{birel}) modeling.
 This package has been grounded on the seminal work of \citet{Alexander08} and its first IT implementation by \citet{Huet18a}.
@@
 Graph theory offers a conceptual framework and indices (global at the entire graph scale, local at the vertex scale) to deal with notions of networks, relationships and neighbourhoods.
-Graphical units (GUs) can be modelled as vertices (nodes) and their spatial relations can be modelled as edges.
-The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.
+Graphical units (GUs) can be modelled as vertices (nodes) separated one with another by undecorated areas.
+Their spatial relations can be modelled as edges.
+The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.\\
+% \smallbreak
+Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
+Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
+The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis (ie, all the graphical content).
+The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive.
+The polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, allows to calculated their Voronoi seeds (ie, their centroid) and Voronoi cells (ie, their area of influence).
+So, the entire decoration image is tiled and condidered as a Voronoi diagram.
+Centroids of the GUs (\code{POINTS}) are calculated and considered as the graph nodes, their locations (\emph{x} and \emph{y} coordinates, measured in pixels) are relative to each decoration.
+Exist edges (\code{LINES}) between pairs of GUs when their Voronoi cells share a common border (\emph{birel}: 'touches').\\
+The entire graphical content of the decoration is mapped with a neighborhood graphs (RNG) -- ie geometric graph, or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
+So a decoration graph is a one component graph which is the dual graph of the Voronoi diagram (Fig.
+\ref{fig:gis}).
 
-Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
-Indeed, for large series of graphical decorations, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
-These \emph{x} and \emph{y} coordinates, measured in pixels, are relative to the decoration figure which is open in the first place in a new GIS project without any projection system.
-The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis.
-The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive.
-The entire decoration image is tiled into different GUs' area of influence like a Voronoi diagram of the support where the Voronoi seeds are the GUs.
-Exist a link between a pair of GUs when their area of influence share a common border.
-A RNG -- also called a geometric graph, or a planar graph -- is constructed from GUs (nodes) and their proximity links (edges).
-The graph itself is as a subgraph of a Delaunay triangulation.
-
+% \begin{figure}
 \begin{figure}[H]
@@
 \caption{\label{fig:gis} GIS interface.
-A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stele (drawing: \cite{DiazGuardamino10}) with its extent (\code{xmin, xmax, ymin, ymax}); B) After the polygonisation of the GUs, including the border of the stelae, the Voronoi cells, the centroid of GUs and the links between GUs having adjacent cells (ie, sharing a border) are calculated; C) For each GUs, \emph{x} and \emph{y} are calculated; D) At least one variable, like the \code{type} of the GUs is defined in order to compute composition analysis.
-A simplier solution will be to create directly centroids (\code{POINTS}) on the GUs and to draw the edges manually}
+A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonsiation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
 \end{figure}
 
+Whether this theoritical process include image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
+\smallbreak
+
 This model has a minimal of \emph{a priori} definitions.
-Those definitions only concern the intrinsic properties of GUs (type, technology, color, orientation, size, etc.) and the types of relations they share.
-Here, we will only consider one property for the nodes, its type (column \code{type}), and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
+Those definitions only concern the intrinsic variables of GUs (\code{type}, \code{technologie}, \code{color}, \code{size}, etc.) and the types of relations they share.
+At least one variable, like the \code{type} of the GUs ('anthropomorph', 'horned helmet', etc.), is defined in order to compute composition analysis.
+Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
 
@@
 This type of edges is directed and, by convention, displayed with a dashed line.
-Its starts from the \emph{main unit} and ends with the \emph{attribute units}.
+Its starts from the \emph{main unit} and ends with the \emph{attribute unit}.
 The textual notation of such an edge is \code{'-+-'}.
-For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6}.
+For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6} (Fig.
+\ref{fig:gis_1}).
 
@@
 \centering
-\includegraphics[scale=0.1]{gis_2}
+\includegraphics[scale=0.05]{gis_2}
 \caption{\label{fig:gis1} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
 \end{figure}
+
+These different kind of relationships are resumed with their birel equivalencies in the following table (Tab. \ref{table:tedges})
+
+\begin{table}[H]
+% \begin{center}
+  \centering
+ \begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
+ \hline
+  % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
+ GU.1 & edge type & GU.2 & (un)directed & \emph{birel} & description \\
+ \hline
+  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint \code{A} and \code{B} contemporaneous \\
+  \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
+  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} is more recent than \code{B} \\
+ % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
+ \hline
+\end{tabular}
+\caption{Synthesis for the different types of relations between GUs}\label{tab1}
+% \end{center}
+\label{table:tedges}
+\end{table}
 
@@
 % latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 14:29:55 2020
+% Thu Nov 26 18:04:41 2020
 \begin{table}[H]
@@
 %
-The \pkg{decorr} package training dataset is composed by four stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
-At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package.
+The \pkg{decorr} package training dataset is composed by five stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
+\\
+At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:timg})
 %
@@
 +  }
-R> out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]])),
-+                            image_append(c(lims[[3]],lims[[4]]))),
+R> out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]],lims[[3]])),
++                            image_append(c(lims[[4]],lims[[5]]))),
 +                          stack = TRUE)
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-5-1}
+{\centering \includegraphics[width=\maxwidth]{figure/timg-1}
 
@@
 
-\caption{\label{fig:figs}Decoration images of the training dataset}\label{fig:unnamed-chunk-5}
+\caption{\label{fig:figs}Decoration images of the training dataset}\label{fig:timg}
 \end{figure}
@@
 
-It contains the required minimum variables for the analysis.
+It contains the required minimum variables for the analysis (Fig. \ref{fig:tnds}).
 
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 14:29:56 2020
+% Thu Nov 26 18:04:43 2020
 \begin{table}[H]
 \centering
-\begin{tabular}{rl}
+\begin{tabular}{rllrlrr}
   \hline
- & V1 \\ 
+ & site & decor & id & type & x & y \\ 
   \hline
-1 & site	decor	id	type	x	y \\ 
-  2 & 1	Cerro Muriano	Cerro Muriano 1	1	personnage	349.814824824793	-298.324408661478 \\ 
-  3 & 2	Cerro Muriano	Cerro Muriano 1	2	casque	349.814824824793	-243.98505280916 \\ 
-  4 & 3	Cerro Muriano	Cerro Muriano 1	3	lance	238.463685783159	-298.324408661478 \\ 
-  5 & 4	Cerro Muriano	Cerro Muriano 1	4	bouclier	446.022208956765	-381.169656108454 \\ 
-  6 & 5	Cerro Muriano	Cerro Muriano 1	5	peigne	283.004141399813	-358.008619187794 \\ 
+1 & Cerro Muriano & Cerro Muriano 1 &   1 & personnage & 349.81 & -298.32 \\ 
+  2 & Cerro Muriano & Cerro Muriano 1 &   2 & casque & 349.81 & -243.99 \\ 
+  3 & Cerro Muriano & Cerro Muriano 1 &   3 & lance & 238.46 & -298.32 \\ 
+  4 & Cerro Muriano & Cerro Muriano 1 &   4 & bouclier & 446.02 & -381.17 \\ 
+  5 & Cerro Muriano & Cerro Muriano 1 &   5 & peigne & 283.00 & -358.01 \\ 
+  6 & Cerro Muriano & Cerro Muriano 1 &   7 & sexe\_masculin & 342.69 & -427.49 \\ 
    \hline
@@
 
-The \code{edges} dataframe is quite similar to the \code{nodes} dataframe.
+The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Fig. \ref{fig:teds}).
 
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 14:29:56 2020
+% Thu Nov 26 18:04:43 2020
 \begin{table}[H]
 \centering
-\begin{tabular}{rl}
+\begin{tabular}{rllrrl}
   \hline
- & V1 \\ 
+ & site & decor & a & b & type \\ 
   \hline
-1 & site	decor	a	b	type \\ 
-  2 & 1	Cerro Muriano	Cerro Muriano 1	1	4	= \\ 
-  3 & 2	Cerro Muriano	Cerro Muriano 1	1	5	= \\ 
-  4 & 3	Cerro Muriano	Cerro Muriano 1	3	5	= \\ 
-  5 & 4	Cerro Muriano	Cerro Muriano 1	1	2	+ \\ 
-  6 & 5	Cerro Muriano	Cerro Muriano 1	1	7	+ \\ 
+1 & Cerro Muriano & Cerro Muriano 1 &   1 &   4 & = \\ 
+  2 & Cerro Muriano & Cerro Muriano 1 &   1 &   5 & = \\ 
+  3 & Cerro Muriano & Cerro Muriano 1 &   3 &   5 & = \\ 
+  4 & Cerro Muriano & Cerro Muriano 1 &   1 &   2 & + \\ 
+  5 & Cerro Muriano & Cerro Muriano 1 &   1 &   7 & + \\ 
+  6 & Cerro Muriano & Cerro Muriano 1 &   3 &   1 & = \\ 
    \hline
@@
 R> cm.1 <- subset(nodes, decor == "Cerro Muriano 1" & id == 1)[,c("x","y")]
+R> cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
+R> paste(as.numeric(cm.1),";",as.numeric(cm.8))
 \end{verbatim}
+\end{kframe}[1] "349.814824824793 ; 451.148926746119"  
+[2] "-298.324408661478 ; -237.478157148203"
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(e, x, parent.frame()): object 'decor' not found}}\begin{verbatim}
-R> cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in eval(e, x, parent.frame()): object 'decor' not found}}\begin{verbatim}
-R> cat(as.numeric(cm.1),";",as.numeric(cm.8))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in cat(as.numeric(cm.1), "{};"{}, as.numeric(cm.8)): object 'cm.1' not found}}\end{kframe}
-
-Once these three dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.
+Once these dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.
 
@@
 The result is a list of decoration graph.
-The first graph of can be plotted
+The first graph of the list can be plotted (Fig.
+\ref{fig:igraph})
 
@@
 R> #                       sep="\t", stringsAsFactors = FALSE)
-R> # nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
+R> # nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
 R> #                     sep="\t",stringsAsFactors = FALSE)
-R> # edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
+R> # edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
 R> #                     sep="\t",stringsAsFactors = FALSE)
 R> lgrph <- list_dec(imgs,nodes,edges,var="type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in igraph::graph\_from\_data\_frame(g.edges, directed = FALSE, vertices = g.nodes): the data frame should contain at least two columns}}\begin{verbatim}
 R> plot(lgrph[[1]],
@@
 \end{verbatim}
+\end{kframe}\begin{figure}[H]
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot(lgrph[[1]], vertex.color = "{}orange"{}, vertex.frame.color = "{}orange"{}, : object 'lgrph' not found}}\end{kframe}
+{\centering \includegraphics[width=\maxwidth]{figure/igraph-1}
+
+}
+
+\caption{\label{fig:figs}R interface. Plot of the first graph of the list}\label{fig:igraph}
+\end{figure}
 
@@
 The \code{read_nds()} function is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
-
-The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe.
-For example, the first \emph{Torrejon Rubio 1} edge, between the nodes \code{6} and \code{5}, has the starting point (xa = 366.7001, ya = -563.1358) and the ending point (xb = 490.1195, yb = -513.2428)
+For example, for the Torrejon Rubio 1 decoration, the 6th GUs is a sword (\code{epee}) (Table \ref{nd}).
 
@@
 R> sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-R> nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
+R> nds.df <- read_nds(site = sit, decor = dec, dev = ".shp",
 +                     doss = system.file("extdata", package = "decorr"))
-R> eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-+                     doss = system.file("extdata", package = "decorr"))
-R> print(xtable::xtable(eds.df[1,],
-+                       caption="first edge of the dataframe",
-+                       label="Test_table_1",
+R> print(xtable::xtable(nds.df[6,],
++                       caption="Sixth node of the Torrejon Rubio 1 decoration",
++                       label="nd",
 +                       size=7),
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Sun Nov 22 14:29:56 2020
+% Thu Nov 26 18:04:43 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rllrlrrrr}
+  \hline
+ & site & decor & id & type & x & y & coords.x1 & coords.x2 \\ 
+  \hline
+13 & Torrejon Rubio & Torrejon Rubio 1 &   6 & epee & 366.70 & -563.14 & 366.70 & -563.14 \\ 
+   \hline
+\end{tabular}
+\caption{Sixth node of the Torrejon Rubio 1 decoration} 
+\label{nd}
+\end{table}
+
+The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe (Table \ref{ed}).
+
+\begin{kframe}
+\begin{verbatim}
+R> # library(decorr)
+R> eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
++                     doss = system.file("extdata", package = "decorr"))
+R> print(xtable::xtable(eds.df[1,],
++                       caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
++                       label="eds.df",
++                       size=7),
++        table.placement="H")
+\end{verbatim}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Thu Nov 26 18:04:43 2020
 \begin{table}[H]
@@
 \end{tabular}
-\caption{first edge of the dataframe} 
-\label{Test_table_1}
+\caption{First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)} 
+\label{eds.df}
 \end{table}
 
-\end{figure}
+% \end{figure}
 
@@
 \setlength\itemsep{.1em}
-  \item \code{plot_dec_grph ()} allows to plot a RNG over a decoration image
+  \item \code{plot_dec_grph()} allows to plot a RNG over a decoration image
 \end{itemize}
 
-Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted, here for the \emph{Torrejon Rubio 1} decoration.
-The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field.
+Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted.
+The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field (Figure \ref{tr_st}).
 
@@
 +                             lbl.txt = "type",
-+                             lbl.size=1.7,
++                             lbl.size=2,
 +                             shw = c("nodes","edges"))
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-11-1}
+{\centering \includegraphics[width=\maxwidth]{figure/tr_st-1}
 
@@
 
-\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:unnamed-chunk-11}
+\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:tr_st}
 \end{figure}
@@
 These comparisons are performed for nodes and/or edges.
-There are four (4) decorations in the default dataset, so there is $\frac{4!}{(4-2)!2!}={6}$ pairwise comparisons
+There are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ pairwise comparisons
 
-\begin{kframe}
-\begin{verbatim}
-R> # set wd to data folder
-R> # setwd(system.file("extdata", package = "decorr"))
-R> # library("decorr")
-R> g.compar <- list_eds_compar(lgrph,"type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
-R> df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
-+                                         g.compar[[2]][[1]]$decor,
-+                                         g.compar[[3]][[1]]$decor,
-+                                         g.compar[[4]][[1]]$decor,
-+                                         g.compar[[5]][[1]]$decor,
-+                                         g.compar[[6]][[1]]$decor),
-+                                decor.B=c(g.compar[[1]][[2]]$decor,
-+                                         g.compar[[2]][[2]]$decor,
-+                                         g.compar[[3]][[2]]$decor,
-+                                         g.compar[[4]][[2]]$decor,
-+                                         g.compar[[5]][[2]]$decor,
-+                                         g.compar[[6]][[2]]$decor))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in data.frame(decor.A = c(g.compar[[1]][[1]]\$decor, g.compar[[2]][[1]]\$decor, : object 'g.compar' not found}}\begin{verbatim}
-R> print(xtable::xtable(df.edges.compar,
-+                       caption="Pairwise comparisons dataframe between decor.A and decor.B",
-+                       label="Test_table_1",
-+                       size=7),
-+        table.placement="H")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.edges.compar, caption = "{}Pairwise comparisons dataframe between decor.A and decor.B"{}, : object 'df.edges.compar' not found}}\end{kframe}
+% <<results='asis', echo=TRUE>>=
+% # set wd to data folder
+% # setwd(system.file("extdata", package = "decorr"))
+% # library("decorr")
+% g.compar <- list_eds_compar(lgrph,"type")
+% df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
+%                                        g.compar[[2]][[1]]$decor,
+%                                        g.compar[[3]][[1]]$decor,
+%                                        g.compar[[4]][[1]]$decor,
+%                                        g.compar[[5]][[1]]$decor,
+%                                        g.compar[[6]][[1]]$decor,
+%                                        g.compar[[7]][[1]]$decor,
+%                                        g.compar[[8]][[1]]$decor,
+%                                        g.compar[[9]][[1]]$decor,
+%                                        g.compar[[10]][[1]]$decor),
+%                               decor.B=c(g.compar[[1]][[2]]$decor,
+%                                        g.compar[[2]][[2]]$decor,
+%                                        g.compar[[3]][[2]]$decor,
+%                                        g.compar[[4]][[2]]$decor,
+%                                        g.compar[[5]][[2]]$decor,
+%                                        g.compar[[6]][[2]]$decor,
+%                                        g.compar[[7]][[2]]$decor,
+%                                        g.compar[[8]][[2]]$decor,
+%                                        g.compar[[9]][[2]]$decor,
+%                                        g.compar[[10]][[2]]$decor))
+% print(xtable::xtable(df.edges.compar,
+%                      caption="Pairwise comparisons dataframe between decor A and decor B",
+%                      label="Test_table_1",
+%                      size=7),
+%       table.placement="H")
+% @
 
@@
 R> df.same_edges <- same_eds(lgrph,"type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
 R> caption <- "Number of same edges between all decoration pairwise comparisons"
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.same\_edges, caption = caption, label = "{}Test\_table\_2"{}, : object 'df.same\_edges' not found}}\end{kframe}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Thu Nov 26 18:04:47 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & 1 & 2 & 3 & 4 & 5 \\ 
+  \hline
+1 & 0 & 0 & 1 & 2 & 0 \\ 
+  2 & 0 & 0 & 3 & 7 & 1 \\ 
+  3 & 1 & 3 & 0 & 1 & 2 \\ 
+  4 & 2 & 7 & 1 & 0 & 1 \\ 
+  5 & 0 & 1 & 2 & 1 & 0 \\ 
+   \hline
+\end{tabular}
+\caption{Number of same edges between all decoration pairwise comparisons} 
+\label{Test_table_2}
+\end{table}
 
@@
 \centering
-\begin{kframe}
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Thu Nov 26 18:04:47 2020
+\begin{tabular}{rrrrrr}
+  & 1 & 2 & 3 & 4 & 5 \\ 
+ 1 & 0 & 2 & 3 & 4 & 2 \\ 
+  2 & 2 & 0 & 5 & 7 & 3 \\ 
+  3 & 3 & 5 & 0 & 4 & 3 \\ 
+  4 & 4 & 7 & 4 & 0 & 3 \\ 
+  5 & 2 & 3 & 3 & 3 & 0 \\ 
+  \end{tabular}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_nodes, digits = c(0)): object 'df.same\_nodes' not found}}\end{kframe}
 \captionof{table}{Common nodes table}
@@
 \centering
-\begin{kframe}
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Thu Nov 26 18:04:47 2020
+\begin{tabular}{rrrrrr}
+  & 1 & 2 & 3 & 4 & 5 \\ 
+ 1 & 0 & 0 & 1 & 2 & 0 \\ 
+  2 & 0 & 0 & 3 & 7 & 1 \\ 
+  3 & 1 & 3 & 0 & 1 & 2 \\ 
+  4 & 2 & 7 & 1 & 0 & 1 \\ 
+  5 & 0 & 1 & 2 & 1 & 0 \\ 
+  \end{tabular}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_edges, digits = c(0)): object 'df.same\_edges' not found}}\end{kframe}
 \captionof{table}{Common edges table}
@@
 R> dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_nodes): object 'df.same\_nodes' not found}}\begin{verbatim}
 R> dist.edges  <- dist(inv(as.matrix(df.same_edges)))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_edges): object 'df.same\_edges' not found}}\begin{verbatim}
 R> plot(hclust(dist.nodes), hang = -1, main = "common nodes")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.nodes): object 'dist.nodes' not found}}\begin{verbatim}
 R> plot(hclust(dist.edges), hang = -1, main = "common edges")
 \end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-11-1}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.edges): object 'dist.edges' not found}}\end{kframe}
 \end{figure}
@@
 R> g.compar <- list_nds_compar(lgrph,"type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
 R> nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
 +                                    doss = system.file("extdata", package = "decorr"))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_nds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : object 'g.compar' not found}}\begin{verbatim}
 R> plot(image_read(nds_compar.3.4))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(nds\_compar.3.4): object 'nds\_compar.3.4' not found}}\begin{verbatim}
 R> g.compar <- list_eds_compar(lgrph,"type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in lapply(lgrph, function(x) x\$name): object 'lgrph' not found}}\begin{verbatim}
 R> eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
 +                                doss = system.file("extdata", package = "decorr"))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unlist(listg, recursive = FALSE): object 'g.compar' not found}}\begin{verbatim}
 R> plot(image_read(eds_compar.3.4))
 \end{verbatim}
+\end{kframe}\begin{figure}[H]
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar.3.4): object 'eds\_compar.3.4' not found}}\end{kframe}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-12-1}
+
+}
+
+\caption{\label{fig:figs}Decoration comparisons between 3 and 4}\label{fig:unnamed-chunk-12}
+\end{figure}
 
@@
 Its geographical equivalent is a Thiessen polygon
-
-\begin{table}[H]
-% \begin{center}
-  \centering
- \begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
- \hline
-  % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
- node 1 & edge type & node 2 & (un)directed & \emph{birel} & description \\
- \hline
-  \code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint, \code{A} and \code{B} can be contemporaneous \\
-  \code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
-  \code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} can be more recent than \code{B} \\
- % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
- \hline
-\end{tabular}
-\caption{Synthesis for the different types of relations between GUs}\label{tab1}
-% \end{center}
-\end{table}
 

### 2020-12-04 article_rvTH15_6_essai.Rnw
huch

\documentclass{article}
\usepackage{setspace}
\begin{document}
\doublespacing
\section*{Question Part 1.}
I am writing this document inside RStudio, using Sweave. I want everything 
in the document to be doublespaced with the exception of R code chunks, 
which I want to be singlespaced for the sake of tidiness and saving space. 
It is important that the code chunks also run in R, since they may be doing 
things like telling R to draw plots which will be part of the document.
\section*{Question Part 2.}
The problem that I seem to be having is that when I put in a command to 
begin singlespacing and a command to end single spacing around an R code 
chunk, text that is not inside the begin and end singlespacing command, for 
some reason, ends up being singlespaced as well, even though as far as I can 
tell it should be doublespaced, as can be seen with the text in this very 
paragraph (if you've compiled it into a .pdf).

\begin{minipage}{\linewidth}
\singlespacing
<<echo=TRUE>>=
k=3
mse=0.347
d=0.85
alpha=0.05
n=30
pwr.df=data.frame(row.names = 1:n)
pwr.df$n=1:n
@
\end{minipage}
Strangely, the problem doesn't occur for text which follows the "end" of the 
singlespacing commmand, but text which comes before the 'begin' 
singlespacing command is affected, as can be seen here.
\end{document}

### 2020-12-04 article_rvTH15_6_essai.tex
huch

\documentclass{article}\usepackage[]{graphicx}\usepackage[]{color}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlstr}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hlstd}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{setspace}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}
\doublespacing
\section*{Question Part 1.}
I am writing this document inside RStudio, using Sweave. I want everything 
in the document to be doublespaced with the exception of R code chunks, 
which I want to be singlespaced for the sake of tidiness and saving space. 
It is important that the code chunks also run in R, since they may be doing 
things like telling R to draw plots which will be part of the document.
\section*{Question Part 2.}
The problem that I seem to be having is that when I put in a command to 
begin singlespacing and a command to end single spacing around an R code 
chunk, text that is not inside the begin and end singlespacing command, for 
some reason, ends up being singlespaced as well, even though as far as I can 
tell it should be doublespaced, as can be seen with the text in this very 
paragraph (if you've compiled it into a .pdf).

\begin{minipage}{\linewidth}
\singlespacing
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{k}\hlkwb{=}\hlnum{3}
\hlstd{mse}\hlkwb{=}\hlnum{0.347}
\hlstd{d}\hlkwb{=}\hlnum{0.85}
\hlstd{alpha}\hlkwb{=}\hlnum{0.05}
\hlstd{n}\hlkwb{=}\hlnum{30}
\hlstd{pwr.df}\hlkwb{=}\hlkwd{data.frame}\hlstd{(}\hlkwc{row.names} \hlstd{=} \hlnum{1}\hlopt{:}\hlstd{n)}
\hlstd{pwr.df}\hlopt{$}\hlstd{n}\hlkwb{=}\hlnum{1}\hlopt{:}\hlstd{n}
\end{alltt}
\end{kframe}
\end{knitrout}
\end{minipage}
Strangely, the problem doesn't occur for text which follows the "end" of the 
singlespacing commmand, but text which comes before the 'begin' 
singlespacing command is affected, as can be seen here.
\end{document}

### 2020-12-04 article_rvTH16.Rnw
huch

--- 2020-11-26 article_rvTH15_6.Rnw
@@
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+options(SweaveHooks=list(fig=function()par(mar=c(0, 0, 0, 0))))
 # options(replace.assign=TRUE, width=90, prompt="R> ")
@@
 \smallbreak
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=10, fig.height=5, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
-# par(mar=c(0,0,0,0))
+<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 g <- graph_from_literal(a-+r, r-+t)
@@
 Whether this theoritical process include image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
-Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
-\smallbreak
-
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\
 This model has a minimal of \emph{a priori} definitions.
@@
 At least one variable, like the \code{type} of the GUs ('anthropomorph', 'horned helmet', etc.), is defined in order to compute composition analysis.
-Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.
-
-\begin{itemize}
-  \item \emph{normal} edges
-\end{itemize}
-
+Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.\\
+\\
+\textbf{\emph{normal}} edges
+\\
+% \begin{itemize}
+%   \item \emph{normal} edges
+% \end{itemize}
 By convention, two contemporaneous GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='}.
@@
 For example: \code{1 -=- 4} means that the nodes \code{1} and \code{4} have a common border.
-
-\begin{itemize}
-  \item \emph{attribute} edges
-\end{itemize}
-
+\smallbreak
+\textbf{\emph{attribute}} edges
+\smallbreak
+% \begin{itemize}
+%   \item \emph{attribute} edges
+% \end{itemize}
 It occurs frequently that a GU can be divided into a \emph{main unit} (eg, a character) and one or various \emph{attribute units} (eg, a helmet, male sex).
@@
 \centering
-\includegraphics[scale=0.05]{gis_1}
+\includegraphics[height=8cm]{gis_1}
 \caption{\label{fig:gis1} GIS interface, Cerro Muriano 1 decoration.
@@
 
-\begin{itemize}
-  \item \emph{overlapping} edges
-\end{itemize}
+\\
+\textbf{\emph{overlapping}} edges
+\\
+% \begin{itemize}
+%   \item \emph{overlapping} edges
+% \end{itemize}
 
@@
 \centering
-\includegraphics[scale=0.05]{gis_2}
-\caption{\label{fig:gis1} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
+\includegraphics[height=8cm]{gis_2}
+\caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
 \end{figure}
@@
   i1 <- image_annotate(i1,lbl.txt,location = "northwest",
-                       size = 25, color = "red")
+                       size = 30, color = "red")
   lims[[length(lims)+1]]<- i1
@@
 
-<<igraph, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:figs}R interface. Plot of the first graph of the list">>=
-# par(mar=c(0.1,0.1,0.1,0.1))
-# library("decorr")
-# library("magick")
-# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-#                       sep="\t", stringsAsFactors = FALSE)
-# nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
+<<igraph, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.cap="\\label{fig:figs}R interface. Plot of the first graph of the list">>=
 lgrph <- list_dec(imgs,nodes,edges,var="type")
@@
 print(xtable::xtable(nds.df[6,],
-                     caption="Sixth node (GU 6) of the Torrejon Rubio 1 decoration",
+                     # caption="Sixth node (GU 6) of the Torrejon Rubio 1 decoration",
                      label="nd",
@@
 <<ed, results='asis', echo=TRUE, fig.cap="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)">>=
-# library(decorr)
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
@@
 print(xtable::xtable(eds.df[1,],
-                     caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
+                     # caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
                      label="eds.df",
@@
 
-<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
+<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
 # library("decorr")
@@
                    doss = system.file("extdata", package = "decorr"))
-img.graph <- plot_dec_grph(nds.df = nds.df,
-                           eds.df = eds.df,
+img.graph <- plot_dec_grph(nodes = nds.df,
+                           edges = eds.df,
                            site = sit,
@@
                            lbl.txt = "type",
-                           lbl.size=2,
-                           shw = c("nodes","edges"))
+                           lbl.size=2)
 plot(image_trim(img.graph))
@@
 <<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.height=8, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
-par(mar=c(0,0,0,0))
-eds_compar <- plot_eds_compar(g.compar, c(1,4),
+# par(mar=c(0,0,0,0))
+g.compar <- list_compar(lgrph,"type")
+eds_compar <- plot_compar(g.compar, c(1,4),
+                          focus = "edges",
                           doss = system.file("extdata", package = "decorr"))
@@
 <<results='asis', echo=TRUE>>=
-df.same_edges <- same_eds(lgrph,"type")
+df.same_edges <- same_neds(lgrph, "type", "edges")
 caption <- "Number of same edges between all decoration pairwise comparisons"
@@
 # lgrph <- list_dec(imgs,nodes,edges,"type")
-df.same_edges <- same_eds(lgrph,"type")
-df.same_nodes<- same_nds(lgrph,"type")
+df.same_edges <- same_neds(lgrph, "type", "edges")
+df.same_nodes<- same_neds(lgrph, "type", "nodes")
 # library("xtable")
@@
 par(mfrow=c(2,1))
-g.compar <- list_nds_compar(lgrph,"type")
+g.compar <- list_compar(lgrph,"type")
 nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
@@
 plot(image_read(nds_compar.3.4))
-g.compar <- list_eds_compar(lgrph,"type")
+# g.compar <- list_eds_compar(lgrph,"type")
 eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),

### 2020-12-04 article_rvTH16.tex
huch

--- 2020-12-01 article_rvTH15_6.tex
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-1-1}
+{\centering \includegraphics[width=\maxwidth]{figure/a_r_t-1}
 
@@
 
-\caption{\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'}\label{fig:unnamed-chunk-1}
+\caption{\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'}\label{fig:a_r_t}
 \end{figure}
@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-2-1}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-1-1}
 
@@
 
-\caption{\label{fig:figs}Potential spatial relations between graphical units}\label{fig:unnamed-chunk-2}
+\caption{\label{fig:figs}Potential spatial relations between graphical units}\label{fig:unnamed-chunk-1}
 \end{figure}
@@
 % latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:41 2020
+% Fri Dec 04 10:36:29 2020
 \begin{table}[H]
@@
 +    i1 <- image_annotate(i1,lbl.txt,location = "northwest",
-+                         size = 25, color = "red")
++                         size = 30, color = "red")
 +    lims[[length(lims)+1]]<- i1
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:43 2020
+% Fri Dec 04 10:36:31 2020
 \begin{table}[H]
@@
 \begin{verbatim}
-R> # par(mar=c(0.1,0.1,0.1,0.1))
-R> # library("decorr")
-R> # library("magick")
-R> # imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-R> #                       sep="\t", stringsAsFactors = FALSE)
-R> # nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
-R> #                     sep="\t",stringsAsFactors = FALSE)
-R> # edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
-R> #                     sep="\t",stringsAsFactors = FALSE)
 R> lgrph <- list_dec(imgs,nodes,edges,var="type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in list\_dec(imgs, nodes, edges, var = "{}type"{}): unused argument (var = "{}type"{})}}\begin{verbatim}
 R> plot(lgrph[[1]],
@@
 \end{verbatim}
-\end{kframe}\begin{figure}[H]
 
-{\centering \includegraphics[width=\maxwidth]{figure/igraph-1}
-
-}
-
-\caption{\label{fig:figs}R interface. Plot of the first graph of the list}\label{fig:igraph}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot(lgrph[[1]], vertex.color = "{}orange"{}, vertex.frame.color = "{}orange"{}, : object 'lgrph' not found}}\end{kframe}
 
@@
 R> print(xtable::xtable(nds.df[6,],
-+                       caption="Sixth node of the Torrejon Rubio 1 decoration",
++                       # caption="Sixth node (GU 6) of the Torrejon Rubio 1 decoration",
 +                       label="nd",
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:43 2020
+% Fri Dec 04 10:36:35 2020
 \begin{table}[H]
 \centering
-\begin{tabular}{rllrlrrrr}
+\begin{tabular}{rllrlrr}
   \hline
- & site & decor & id & type & x & y & coords.x1 & coords.x2 \\ 
+ & site & decor & id & type & x & y \\ 
   \hline
-13 & Torrejon Rubio & Torrejon Rubio 1 &   6 & epee & 366.70 & -563.14 & 366.70 & -563.14 \\ 
+13 & Torrejon Rubio & Torrejon Rubio 1 &   6 & epee & 366.70 & -563.14 \\ 
    \hline
 \end{tabular}
-\caption{Sixth node of the Torrejon Rubio 1 decoration} 
 \label{nd}
@@
 \begin{verbatim}
-R> # library(decorr)
 R> eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
@@
 R> print(xtable::xtable(eds.df[1,],
-+                       caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
++                       # caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
 +                       label="eds.df",
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:43 2020
+% Fri Dec 04 10:36:35 2020
 \begin{table}[H]
@@
 \end{tabular}
-\caption{First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)} 
 \label{eds.df}
@@
 +                     doss = system.file("extdata", package = "decorr"))
-R> img.graph <- plot_dec_grph(nds.df = nds.df,
-+                             eds.df = eds.df,
+R> img.graph <- plot_dec_grph(nodes = nds.df,
++                             edges = eds.df,
 +                             site = sit,
@@
 +                             lbl.txt = "type",
-+                             lbl.size=2,
-+                             shw = c("nodes","edges"))
++                             lbl.size=2)
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_dec\_grph(nodes = nds.df, edges = eds.df, site = sit, decor = dec, : unused argument (lbl.txt = "{}type"{})}}\begin{verbatim}
 R> plot(image_trim(img.graph))
 \end{verbatim}
-\end{kframe}\begin{figure}[H]
 
-{\centering \includegraphics[width=\maxwidth]{figure/tr_st-1}
-
-}
-
-\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:tr_st}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in assert\_image(image): object 'img.graph' not found}}\end{kframe}
 
@@
 \begin{verbatim}
-R> par(mar=c(0,0,0,0))
-R> eds_compar <- plot_eds_compar(g.compar, c(1,4),
+R> # par(mar=c(0,0,0,0))
+R> g.compar <- list_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
+R> eds_compar <- plot_compar(g.compar, c(1,4),
++                            focus = "edges",
 +                            doss = system.file("extdata", package = "decorr"))
@@
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in unlist(listg, recursive = FALSE): object 'g.compar' not found}}\begin{verbatim}
-R> plot(image_trim(image_read(eds_compar)))
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_compar(g.compar, c(1, 4), focus = "{}edges"{}, doss = system.file("{}extdata"{}, : object 'g.compar' not found}}\begin{verbatim}
+R> plot(magick::image_trim(magick::image_read(eds_compar)))
 \end{verbatim}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar): object 'eds\_compar' not found}}\end{kframe}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in magick::image\_read(eds\_compar): path must be URL, filename or raw vector}}\end{kframe}
 
-The comparison shows that \code{1} (\emph{Cerro Muriano 1}) and \code{4} (\emph{Zarza de Montsanchez}) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.
+The comparison shows that \code{1} (Cerro Muriano 1) and \code{4} (Zarza de Montsanchez) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.
 
@@
 \begin{verbatim}
-R> df.same_edges <- same_eds(lgrph,"type")
+R> df.same_edges <- same_neds(lgrph, "type", "edges")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}edges"{}): could not find function "{}same\_neds"{}}}\begin{verbatim}
 R> caption <- "Number of same edges between all decoration pairwise comparisons"
@@
 \end{verbatim}
-\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:47 2020
-\begin{table}[H]
-\centering
-\begin{tabular}{rrrrrr}
-  \hline
- & 1 & 2 & 3 & 4 & 5 \\ 
-  \hline
-1 & 0 & 0 & 1 & 2 & 0 \\ 
-  2 & 0 & 0 & 3 & 7 & 1 \\ 
-  3 & 1 & 3 & 0 & 1 & 2 \\ 
-  4 & 2 & 7 & 1 & 0 & 1 \\ 
-  5 & 0 & 1 & 2 & 1 & 0 \\ 
-   \hline
-\end{tabular}
-\caption{Number of same edges between all decoration pairwise comparisons} 
-\label{Test_table_2}
-\end{table}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.same\_edges, caption = caption, label = "{}Test\_table\_2"{}, : object 'df.same\_edges' not found}}\end{kframe}
 
@@
 \centering
-% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:47 2020
-\begin{tabular}{rrrrrr}
-  & 1 & 2 & 3 & 4 & 5 \\ 
- 1 & 0 & 2 & 3 & 4 & 2 \\ 
-  2 & 2 & 0 & 5 & 7 & 3 \\ 
-  3 & 3 & 5 & 0 & 4 & 3 \\ 
-  4 & 4 & 7 & 4 & 0 & 3 \\ 
-  5 & 2 & 3 & 3 & 3 & 0 \\ 
-  \end{tabular}
+\begin{kframe}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}edges"{}): could not find function "{}same\_neds"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}nodes"{}): could not find function "{}same\_neds"{}}}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_nodes, digits = c(0)): object 'df.same\_nodes' not found}}\end{kframe}
 \captionof{table}{Common nodes table}
@@
 \centering
-% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Thu Nov 26 18:04:47 2020
-\begin{tabular}{rrrrrr}
-  & 1 & 2 & 3 & 4 & 5 \\ 
- 1 & 0 & 0 & 1 & 2 & 0 \\ 
-  2 & 0 & 0 & 3 & 7 & 1 \\ 
-  3 & 1 & 3 & 0 & 1 & 2 \\ 
-  4 & 2 & 7 & 1 & 0 & 1 \\ 
-  5 & 0 & 1 & 2 & 1 & 0 \\ 
-  \end{tabular}
+\begin{kframe}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_edges, digits = c(0)): object 'df.same\_edges' not found}}\end{kframe}
 \captionof{table}{Common edges table}
@@
 R> dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_nodes): object 'df.same\_nodes' not found}}\begin{verbatim}
 R> dist.edges  <- dist(inv(as.matrix(df.same_edges)))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_edges): object 'df.same\_edges' not found}}\begin{verbatim}
 R> plot(hclust(dist.nodes), hang = -1, main = "common nodes")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.nodes): object 'dist.nodes' not found}}\begin{verbatim}
 R> plot(hclust(dist.edges), hang = -1, main = "common edges")
 \end{verbatim}
-\end{kframe}
-\includegraphics[width=\maxwidth]{figure/unnamed-chunk-11-1}
 
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.edges): object 'dist.edges' not found}}\end{kframe}
 \end{figure}
@@
 R> par(mfrow=c(2,1))
-R> g.compar <- list_nds_compar(lgrph,"type")
+R> g.compar <- list_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
 R> nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
 +                                    doss = system.file("extdata", package = "decorr"))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_nds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : could not find function "{}plot\_nds\_compar"{}}}\begin{verbatim}
 R> plot(image_read(nds_compar.3.4))
-R> g.compar <- list_eds_compar(lgrph,"type")
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(nds\_compar.3.4): object 'nds\_compar.3.4' not found}}\begin{verbatim}
+R> # g.compar <- list_eds_compar(lgrph,"type")
 R> eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
 +                                doss = system.file("extdata", package = "decorr"))
+\end{verbatim}
+
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_eds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : could not find function "{}plot\_eds\_compar"{}}}\begin{verbatim}
 R> plot(image_read(eds_compar.3.4))
 \end{verbatim}
-\end{kframe}\begin{figure}[H]
 
-{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-12-1}
-
-}
-
-\caption{\label{fig:figs}Decoration comparisons between 3 and 4}\label{fig:unnamed-chunk-12}
-\end{figure}
+{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar.3.4): object 'eds\_compar.3.4' not found}}\end{kframe}
 

### 2020-12-04 article_rvTH16.Rnw
rerun 16

@@
 -- and at the sign scale with the development of archaeological sciences -- radiocarbon dating, use-wear analysis, elemental analysis, etc.
---, these improvement do not necessarly help to understand the semantic content of the iconography.
-
+--, these improvement do not necessarly help to understand the semantic content of the iconography.\\
 Semantics or semiotics can be defined as a system of conventional features -- called signs -- organised also in conventional manners.
@@
 \smallbreak
-<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=3, fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", out.height="50%", fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 g <- graph_from_literal(a-+r, r-+t)
@@
 \centering
-\includegraphics[height=8cm]{gis_1}
+\includegraphics[height=10cm, keepaspectratio]{gis_1}
 \caption{\label{fig:gis1} GIS interface, Cerro Muriano 1 decoration.
@@
 \centering
-\includegraphics[height=8cm]{gis_2}
+\includegraphics[height=10cm, keepaspectratio]{gis_2}
 \caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
@@
 <<igraph, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.cap="\\label{fig:figs}R interface. Plot of the first graph of the list">>=
-lgrph <- list_dec(imgs,nodes,edges,var="type")
+lgrph <- list_dec(imgs,nodes,edges)
 plot(lgrph[[1]],
@@
                            doss = system.file("extdata", package = "decorr"),
-                           lbl.txt = "type",
+                           nds.lbl = "type",
                            lbl.size=2)
@@
 \setlength\itemsep{.1em}
-  \item \code{same_nds()} and \code{same_eds()} functions allow to repectively count matching nodes and matching edges between decoration pairwises
+\item \code{same_elements()} function allows to count matching nodes and matching edges between decoration pairwises
 \end{itemize}
 
-\code{same_nds()} and \code{same_eds()} allow to repectively count matching nodes and matching edges between decoration pairwises.
+\code{same_elements()} allows to count matching nodes and matching edges between decoration pairwises.
 The result is a square matrix between all pairwise comparisons with the number of common nodes or edges in the cells.
@@
 <<results='asis', echo=TRUE>>=
-df.same_edges <- same_neds(lgrph, "type", "edges")
+df.same_edges <- same_elements(lgrph, "type", "edges")
 caption <- "Number of same edges between all decoration pairwise comparisons"
@@
 # lgrph <- list_dec(imgs,nodes,edges,"type")
-df.same_edges <- same_neds(lgrph, "type", "edges")
-df.same_nodes<- same_neds(lgrph, "type", "nodes")
+df.same_edges <- same_elements(lgrph, "type", "edges")
+df.same_nodes<- same_elements(lgrph, "type", "nodes")
 # library("xtable")
-print(xtable(df.same_nodes, digits=c(0)),
-      floating=FALSE,
-      hline.after=NULL)
+print(xtable::xtable(df.same_nodes,
+                     size = 8,
+                     digits = c(0)),
+      table.placement="H",
+      include.rownames = TRUE)
+# print(xtable(df.same_nodes, digits=c(0)),
+#       floating=FALSE,
+#       hline.after=NULL)
 @
@@
 <<echo=FALSE,results='asis'>>=
-print(xtable(df.same_edges, digits=c(0)),
-      floating=FALSE,
-      hline.after=NULL)
+print(xtable::xtable(df.same_edges,
+                     size = 8,
+                     digits = c(0)),
+      table.placement="H",
+      include.rownames = TRUE)
+# print(xtable(df.same_edges, digits=c(0)),
+#       floating=FALSE,
+#       hline.after=NULL)
 @
@@
 g.compar <- list_compar(lgrph,"type")
-nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
+nds_compar.3.4 <- plot_compar(g.compar, c(3,4),
+                              focus = "nodes",
                               doss = system.file("extdata", package = "decorr"))
-plot(image_read(nds_compar.3.4))
+plot(magick::image_read(nds_compar.3.4))
 # g.compar <- list_eds_compar(lgrph,"type")
-eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
+eds_compar.3.4 <- plot_compar(g.compar, c(3,4),
+                              focus = "edges",
                               doss = system.file("extdata", package = "decorr"))
-plot(image_read(eds_compar.3.4))
+plot(magick::image_read(eds_compar.3.4))
 @

### 2020-12-04 article_rvTH16.tex
rerun 16

@@
 
-{\centering \includegraphics[width=\maxwidth]{figure/a_r_t-1}
+{\centering \includegraphics[width=\maxwidth,height=0.5\textheight]{figure/a_r_t-1}
 
@@
 % latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Fri Dec 04 10:36:29 2020
+% Fri Dec 04 11:03:04 2020
 \begin{table}[H]
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Fri Dec 04 10:36:31 2020
+% Fri Dec 04 11:03:05 2020
 \begin{table}[H]
@@
 \begin{verbatim}
-R> lgrph <- list_dec(imgs,nodes,edges,var="type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in list\_dec(imgs, nodes, edges, var = "{}type"{}): unused argument (var = "{}type"{})}}\begin{verbatim}
+R> lgrph <- list_dec(imgs,nodes,edges)
 R> plot(lgrph[[1]],
@@
 \end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
-% Fri Dec 04 10:36:35 2020
+% Fri Dec 04 11:03:06 2020
 \begin{table}[H]
@@
 +                             doss = system.file("extdata", package = "decorr"),
-+                             lbl.txt = "type",
++                             nds.lbl = "type",
 +                             lbl.size=2)
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_dec\_grph(nodes = nds.df, edges = eds.df, site = sit, decor = dec, : unused argument (lbl.txt = "{}type"{})}}\begin{verbatim}
 R> plot(image_trim(img.graph))
 \end{verbatim}
+\end{kframe}\begin{figure}[H]
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in assert\_image(image): object 'img.graph' not found}}\end{kframe}
+{\centering \includegraphics[width=\maxwidth]{figure/tr_st-1}
+
+}
+
+\caption{\label{fig:figs}Torrejon Rubio 1 decoration}\label{fig:tr_st}
+\end{figure}
 
@@
 R> g.compar <- list_compar(lgrph,"type")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
 R> eds_compar <- plot_compar(g.compar, c(1,4),
@@
 +                            doss = system.file("extdata", package = "decorr"))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_compar(g.compar, c(1, 4), focus = "{}edges"{}, doss = system.file("{}extdata"{}, : object 'g.compar' not found}}\begin{verbatim}
 R> plot(magick::image_trim(magick::image_read(eds_compar)))
 \end{verbatim}
+\end{kframe}\begin{figure}[H]
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in magick::image\_read(eds\_compar): path must be URL, filename or raw vector}}\end{kframe}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-6-1}
+
+}
+
+\caption{\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations}\label{fig:unnamed-chunk-6}
+\end{figure}
 
@@
 \begin{verbatim}
-R> df.same_edges <- same_neds(lgrph, "type", "edges")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}edges"{}): could not find function "{}same\_neds"{}}}\begin{verbatim}
+R> df.same_edges <- same_elements(lgrph, "type", "edges")
 R> caption <- "Number of same edges between all decoration pairwise comparisons"
@@
 \end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable::xtable(df.same\_edges, caption = caption, label = "{}Test\_table\_2"{}, : object 'df.same\_edges' not found}}\end{kframe}
+\end{kframe}% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Fri Dec 04 11:03:10 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & 1 & 2 & 3 & 4 & 5 \\ 
+  \hline
+1 & 8 & 0 & 1 & 2 & 0 \\ 
+  2 & 0 & 15 & 3 & 7 & 1 \\ 
+  3 & 1 & 3 & 10 & 1 & 2 \\ 
+  4 & 2 & 7 & 1 & 14 & 1 \\ 
+  5 & 0 & 1 & 2 & 1 & 4 \\ 
+   \hline
+\end{tabular}
+\caption{Number of same edges between all decoration pairwise comparisons} 
+\label{Test_table_2}
+\end{table}
 
@@
 \centering
-\begin{kframe}
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Fri Dec 04 11:03:10 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & 1 & 2 & 3 & 4 & 5 \\ 
+  \hline
+1 & 7 & 2 & 3 & 4 & 2 \\ 
+  2 & 2 & 12 & 5 & 9 & 3 \\ 
+  3 & 3 & 5 & 6 & 4 & 3 \\ 
+  4 & 4 & 9 & 4 & 12 & 3 \\ 
+  5 & 2 & 3 & 3 & 3 & 4 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}edges"{}): could not find function "{}same\_neds"{}}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in same\_neds(lgrph, "{}type"{}, "{}nodes"{}): could not find function "{}same\_neds"{}}}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_nodes, digits = c(0)): object 'df.same\_nodes' not found}}\end{kframe}
 \captionof{table}{Common nodes table}
@@
 \centering
-\begin{kframe}
+% latex table generated in R 3.6.1 by xtable 1.8-4 package
+% Fri Dec 04 11:03:10 2020
+\begin{table}[H]
+\centering
+\begin{tabular}{rrrrrr}
+  \hline
+ & 1 & 2 & 3 & 4 & 5 \\ 
+  \hline
+1 & 8 & 0 & 1 & 2 & 0 \\ 
+  2 & 0 & 15 & 3 & 7 & 1 \\ 
+  3 & 1 & 3 & 10 & 1 & 2 \\ 
+  4 & 2 & 7 & 1 & 14 & 1 \\ 
+  5 & 0 & 1 & 2 & 1 & 4 \\ 
+   \hline
+\end{tabular}
+\end{table}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in xtable(df.same\_edges, digits = c(0)): object 'df.same\_edges' not found}}\end{kframe}
 \captionof{table}{Common edges table}
@@
 R> dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_nodes): object 'df.same\_nodes' not found}}\begin{verbatim}
 R> dist.edges  <- dist(inv(as.matrix(df.same_edges)))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in as.matrix(df.same\_edges): object 'df.same\_edges' not found}}\begin{verbatim}
 R> plot(hclust(dist.nodes), hang = -1, main = "common nodes")
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.nodes): object 'dist.nodes' not found}}\begin{verbatim}
 R> plot(hclust(dist.edges), hang = -1, main = "common edges")
 \end{verbatim}
+\end{kframe}
+\includegraphics[width=\maxwidth]{figure/unnamed-chunk-10-1}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in hclust(dist.edges): object 'dist.edges' not found}}\end{kframe}
 \end{figure}
@@
 R> g.compar <- list_compar(lgrph,"type")
+R> nds_compar.3.4 <- plot_compar(g.compar, c(3,4),
++                                focus = "nodes",
++                                doss = system.file("extdata", package = "decorr"))
+R> plot(magick::image_read(nds_compar.3.4))
+R> # g.compar <- list_eds_compar(lgrph,"type")
+R> eds_compar.3.4 <- plot_compar(g.compar, c(3,4),
++                                focus = "edges",
++                                doss = system.file("extdata", package = "decorr"))
+R> plot(magick::image_read(eds_compar.3.4))
 \end{verbatim}
+\end{kframe}\begin{figure}[H]
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in utils::combn(1:length(lgrph), 2): object 'lgrph' not found}}\begin{verbatim}
-R> nds_compar.3.4 <- plot_nds_compar(g.compar, c(3,4),
-+                                    doss = system.file("extdata", package = "decorr"))
-\end{verbatim}
+{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-11-1}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_nds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : could not find function "{}plot\_nds\_compar"{}}}\begin{verbatim}
-R> plot(image_read(nds_compar.3.4))
-\end{verbatim}
+}
 
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(nds\_compar.3.4): object 'nds\_compar.3.4' not found}}\begin{verbatim}
-R> # g.compar <- list_eds_compar(lgrph,"type")
-R> eds_compar.3.4 <- plot_eds_compar(g.compar, c(3,4),
-+                                doss = system.file("extdata", package = "decorr"))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in plot\_eds\_compar(g.compar, c(3, 4), doss = system.file("{}extdata"{}, : could not find function "{}plot\_eds\_compar"{}}}\begin{verbatim}
-R> plot(image_read(eds_compar.3.4))
-\end{verbatim}
-
-{\ttfamily\noindent\bfseries\color{errorcolor}{\#\# Error in image\_read(eds\_compar.3.4): object 'eds\_compar.3.4' not found}}\end{kframe}
+\caption{\label{fig:figs}Decoration comparisons between 3 and 4}\label{fig:unnamed-chunk-11}
+\end{figure}
 

### 2020-12-07 article_rvTH16.Rnw
ud

@@
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
-options(SweaveHooks=list(fig=function()par(mar=c(0, 0, 0, 0))))
+options(SweaveHooks=list(fig=function() par(mar=c(0, 0, 0, 0))))
 # options(replace.assign=TRUE, width=90, prompt="R> ")
@@
 For example, the 3-letters word "\code{art}" can be modelled with three vertices (\code{a}, \code{r}, \code{t}) and two edges (one between \code{a} and \code{r}, the other between \code{r} and \code{t}).
-In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat}.
+In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Fig.
+\ref{fig:fart}).
+
 \smallbreak
-<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", out.height="50%", fig.cap="\\label{fig:figs}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center",fig.asp=0.1, fig.width=10, fig.cap="\\label{fig:fart}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+par(mar=c(0.1,0.1,0.1,0.1) )
 g <- graph_from_literal(a-+r, r-+t)
@@
      vertex.frame.color="white",
-     vertex.size=60,
-     vertex.label.cex=1.5,
+     vertex.size=80,
+     vertex.label.cex=1,
      vertex.label.color="black",
@@
 But, as stated, in Prehistorical the writing system does not exists.
-Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order.
+Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order (Fig.
+\ref{fig:fratar}).
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:figs}Potential spatial relations between graphical units">>=
+<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:fratar}Potential spatial relations between graphical units">>=
 par(mar=c(0.1,0.1,0.1,0.1) )
 ugs <- data.frame(name=c("a","r","t"))
-lks <- data.frame(from=c("a", "r", "t"),to=c("r", "t", "a"),type=c("=","=","="))
+lks <- data.frame(from=c("a", "r", "t"), to=c("r", "t", "a"), type=c("=","=","="))
 g <- igraph::graph_from_data_frame(lks, directed=FALSE, vertices= ugs)
@@
 \\
+According to us, prehistorical composition -- like any formal system -- are spatial features (mostly shapes) related one with the other depending on rules of spatial proximities.
+Then, the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70} should be more operative than the Saussurian '\emph{linearity of the signifier}' principle \citep{Saussure89}.
+To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory and particularly relative neighborhood graphs (RNG): '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}.
 \\
 In this article we present the \proglang{R} package \pkg{decorr}.
-Its purpose is to formalise a method based on relative neighborhood graphs (RNG) to analyse any graphical content (here: decorations).
-As any formal system, iconography can be modelled as spatial features related one with the other depending on rules of spatial proximities; as state by the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70}.
-And, because any graphical system is composed by features (mostly shapes) displayed spatialy, these systems can be represented by features connected (or not) to each other depending on their spatial proximity.
-To map and analyse these proximities, the model uses concepts and methods coming from Graph Theory, GIS and binary topological relationships (\emph{birel}) modeling.
+Its purpose is to formalise a method and offer tools to analyse any graphical content (here: decorations) with methods coming from raph Theory, GIS and binary topological relationships (\emph{birel}) modeling.
 This package has been grounded on the seminal work of \citet{Alexander08} and its first IT implementation by \citet{Huet18a}.
@@
 \\
-% \begin{itemize}
-%   \item \emph{normal} edges
-% \end{itemize}
 By convention, two contemporaneous GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='}.
@@
 The textual notation of such an edge is \code{'-=-'}.
-For example: \code{1 -=- 4} means that the nodes \code{1} and \code{4} have a common border.
+For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Fig.
+\ref{fig:gis_1})
 \smallbreak
 \textbf{\emph{attribute}} edges
-\smallbreak
-% \begin{itemize}
-%   \item \emph{attribute} edges
-% \end{itemize}
-It occurs frequently that a GU can be divided into a \emph{main unit} (eg, a character) and one or various \emph{attribute units} (eg, a helmet, male sex).
+\\
+It occurs frequently that a GU can be divided into a \emph{main unit} (eg, an anthropomorph) and one or various \emph{attribute units} (eg, a helmet, a male sex).
 Broadly speaking, for further statistical analysis, it is better to use this \emph{attribute} method than to multiply the categories of GUs.
@@
 The textual notation of such an edge is \code{'-+-'}.
-For example \code{4 -+- 6} means that the main node \code{4} has the attribute node \code{6} (Fig.
+For example \code{personnage -+- casque} means that the main node \code{personnage} has the attribute node \code{casque} (Fig.
 \ref{fig:gis_1}).
@@
 \includegraphics[height=10cm, keepaspectratio]{gis_1}
-\caption{\label{fig:gis1} GIS interface, Cerro Muriano 1 decoration.
+\caption{\label{fig:gis_1} GIS interface, Cerro Muriano 1 decoration.
 The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the GU \code{personnage} (anthropomorph).}
@@
 
-\\
+\smallbreak
 \textbf{\emph{overlapping}} edges
 \\
-% \begin{itemize}
-%   \item \emph{overlapping} edges
-% \end{itemize}
-
-Finally, it is quite common that a graphical composition shows superimpositions between different UGs permit to distinguish different decoration phases for a single support.
+Finally, it is quite common that a graphical composition shows superimpositions between different GUs permitting to distinguish different decoration phases, or iconographic layers, for a single support.
 So, at first, the analyse must be performed on each different phases of decoration separately.
@@
 The textual notation of such an edge is \code{'->-'}.
-For example \code{A ->- B} means that \code{A} crosses \code{B} (ie, \code{A} overlaps \code{B} in the stratigraphical sense) (Fig.
+For example \code{ecriture ->- bouclier} means that \code{ecriture} crosses \code{bouclier} (ie, \code{ecriture} overlaps \code{bouclier} in the stratigraphical sense) (Fig.
 \ref{fig:gis_2}).
@@
 \centering
-\begin{tabular}{|p{.5cm} p{.5cm} p{.5cm} p{2cm} p{2cm} p{5cm}|}
+\begin{tabular}{|p{.6cm} p{.5cm} p{.6cm} p{2cm} p{2cm} p{5cm}|}
 \hline
@@
 A training dataset (nodes and edges coordinates, decoration images) is stored in the \code{extdata} folder of the \pkg{decorr} package
-
-\begin{itemize}
-\item The \code{imgs} dataframe
-\end{itemize}
-
-The inventory of decorations is stored in the \code{imgs} dataframe.
+\smallbreak
+% \begin{itemize}
+% \item The \code{imgs} dataframe
+% \end{itemize}
+\textbf{\code{imgs} dataframe}
+\\
+The inventory of decorations is stored in the \code{imgs} dataframe (Tab.
+\ref{tab:timg}).
 The field \code{imgs$idf} is the short name of the decoration, useful during statistical analysis.
 The primary key of each decoration is the concatenate of \code{imgs$site} and \code{imgs$decor}.
-These keys will allow joints with the other dataframes (\code{nodes} and \code{edges})\\
+These keys will allow joints with the other dataframes (\code{nodes} and \code{edges})
 %
@@
                      caption="The studied corpus, the \\code{imgs.tsv} dataframe",
-                     label="Test_table",
+                     label="tab:timg",
                      size=6,
@@
 \\
-At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:timg})
+At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:dataset})
 %
-<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=18, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
+<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=18, fig.height=13, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
 # library("magick")
@@
 @
+% 
+% \begin{itemize}
+% \item The \code{nodes} dataframe
+% \end{itemize}
+\smallbreak
+\textbf{\code{nodes} dataframe}
+\\
+It contains the required minimum variables for the analysis (Tab. \ref{tab:tnds}).
 
-\begin{itemize}
-\item The \code{nodes} dataframe
-\end{itemize}
-
-It contains the required minimum variables for the analysis (Fig. \ref{fig:tnds}).
-
-<<tnds, results='asis', echo=TRUE>>=
+<<tndss, results='asis', echo=TRUE>>=
 caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
 print(xtable::xtable(head(nodes),
-                     caption = caption),
+                     caption = caption,
+                     label="tab:tnds"),
       table.placement="H")
@@
 To recover the local \emph{y} value of GUs nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
+\smallbreak
+% \begin{itemize}
+% \item The \code{edges} dataframe
+% \end{itemize}
+\textbf{\code{edges} dataframe}
+\\
+The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Tab. \ref{tab:teds}).
 
-\begin{itemize}
-\item The \code{edges} dataframe
-\end{itemize}
-
-The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Fig. \ref{fig:teds}).
-
-<<teds, results='asis', echo=TRUE>>=
+<<tedss, results='asis', echo=TRUE>>=
 caption <- "Edges (from \\code{edges.csv} dataframe)}"
@@
                      caption=caption,
-                     label="edgesdf"),
+                     label="tab:teds"),
       table.placement="H")
@@
 
-<<igraph, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.cap="\\label{fig:figs}R interface. Plot of the first graph of the list">>=
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.cap="\\label{fig:igraph}R interface. Plot of the first graph of the list">>=
 lgrph <- list_dec(imgs,nodes,edges)
@@
                            doss = system.file("extdata", package = "decorr"),
-                           nds.lbl = "type",
+                           nd.var = "type",
                            lbl.size=2)

### 2020-12-07 article_rvTH15_6_essai.Rnw
rm

@@
-\documentclass{article}
-\usepackage{setspace}
+\documentclass[article]{jss}
+
+%% -- LaTeX packages and custom commands ---------------------------------------
+
+%% recommended packages
+\usepackage{thumbpdf,lmodern}
+\usepackage{float,tikz}
+%% another package (only for this demo article)
+\usepackage{framed}
+%% author pakckage
+\usepackage{xcolor,caption, float}
+
+\graphicspath{ {./imgs/} }
+
+%% new custom commands
+\newcommand{\class}[1]{`\code{#1}'}
+\newcommand{\fct}[1]{\code{#1()}}
+% \setlist[itemize]{noitemsep, topsep=0pt}
+\SweaveOpts{engine=R, eps=FALSE, keep.source = TRUE}
+
+<<preliminaries, echo=FALSE, include=FALSE, cache=FALSE >>=
+# options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+library("MASS")
+library("extrafont") #;library("ggplot")
+library("igraph")
+opts_chunk$set(prompt=TRUE, fig.pos = 'H', results='asis', highlight=FALSE)
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+options(SweaveHooks=list(fig=function() par(mar=c(0, 0, 0, 0))))
+# options(replace.assign=TRUE, width=90, prompt="R> ")
+opts_knit$set(concordance=TRUE, self.contained=FALSE, background = "#FFFFFF")
+# knitr::opts_chunk$set(fig.pos = 'H')
+loadfonts(device = "win")
+@
+
+%% -- Article metainformation (author, title, ...) -----------------------------
+
+%% - \author{} with primary affiliation
+%% - \Plainauthor{} without affiliations
+%% - Separate authors by \And or \AND (in \author) or by comma (in \Plainauthor).
+%% - \AND starts a new line, \And does not.
+\author{Thomas Huet\\UMR 5140 \And Craig Alexander \And Jose Pozo}
+%   \And Second Author\\Plus Affiliation}
+\Plainauthor{Thomas Huet, Craig Alexander, Jose Pozo}
+
+%% - \title{} in title case
+%% - \Plaintitle{} without LaTeX markup (if any)
+%% - \Shorttitle{} with LaTeX markup (if any), used as running title
+\title{Modelling Prehistoric Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
+\Plaintitle{Modelling Prehistoric Iconographic Compositions. The R package 'decorr'}
+\Shorttitle{Modelling Prehistoric Iconographic Compositions}
+
+%% - \Abstract{} almost as usual
+\Abstract{
+By definition, Prehistorical societies are characterised by the absence of a writing system.
+During, the largest part of human history -- from far -- symbolic expressions belong to illiterate societies which express themself with rock-art paintings, pottery decorations, figurines and statuary, etc., and a lot of now disappeared carved woods, textil design, etc.
+At the composition level, recognition of meaningful associations of signs and reccurent patterns indicate clearly the existence of social conventions in the way to display and to read these expressions.
+We present the \pkg{decorr} \proglang{R} package which grounds concepts, methods and tools to analyse any ancient graphical systems.
+Our asumption is that i) any graphical system is a spatial distribution of features, and ii) these features have possibly any meaningful relationships one with another depending on their pairwise spatial proximities.
+To model the graphical compositions we employ concepts coming from the Graph Theory.
+To ensure the feasability of this type of analysis, we propose a GIS-based method for inputs and a serie of functions for data management.
+}
+
+\Keywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (\emph{birel}), \proglang{R}}
+\Plainkeywords{Semiotics, Iconography, Prehistory, Graphs, Spatial Analysis, Binary Topological Relationships (birel), R}
+
+\Address{
+Thomas Huet\\
+CNRS-UMR 5140 \\
+Archeologie des Societes Mediterraneennes\\
+Universite Paul Valery\\
+route de Mende\\
+Montpellier 34199, France\\
+E-mail: \email{thomashuet7@gmail.com}\\
+%  URL: \url{https://eeecon.uibk.ac.at/~zeileis/}
+}
+
 \begin{document}
-\doublespacing
-\section*{Question Part 1.}
-I am writing this document inside RStudio, using Sweave. I want everything 
-in the document to be doublespaced with the exception of R code chunks, 
-which I want to be singlespaced for the sake of tidiness and saving space. 
-It is important that the code chunks also run in R, since they may be doing 
-things like telling R to draw plots which will be part of the document.
-\section*{Question Part 2.}
-The problem that I seem to be having is that when I put in a command to 
-begin singlespacing and a command to end single spacing around an R code 
-chunk, text that is not inside the begin and end singlespacing command, for 
-some reason, ends up being singlespaced as well, even though as far as I can 
-tell it should be doublespaced, as can be seen with the text in this very 
-paragraph (if you've compiled it into a .pdf).
+% \SweaveOpts{concordance=TRUE}
+% \Sweave2knitr("article_rvTH14_1.Rnw")
+% \SweaveOpts{concordance=TRUE}
 
-\begin{minipage}{\linewidth}
-\singlespacing
-<<echo=TRUE>>=
-k=3
-mse=0.347
-d=0.85
-alpha=0.05
-n=30
-pwr.df=data.frame(row.names = 1:n)
-pwr.df$n=1:n
+\section[Introduction]{Introduction} \label{sec:intro}
+
+The \pkg{decorr} package training dataset is composed by five stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
+\\
+At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:timg})
+%
+<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=18, fig.height=13, fig.cap="\\label{fig:figs}Decoration images of the training dataset">>=
+# library("magick")
+pth <- system.file("extdata", package = "decorr")
+imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
+                   sep="\t", stringsAsFactors = FALSE)
+lims <- list()
+for(i in 1:nrow(imgs)){
+  i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
+  lbl.txt <- paste0(imgs[i,"idf"],"\n",
+                    imgs[i,"site"],"\n",
+                    imgs[i,"decor"],"\n",
+                    imgs[i,"img"],"\n",
+                    image_info(i1)$width,"*",image_info(i1)$height," px")
+  i1 <- image_annotate(i1,lbl.txt,location = "northwest",
+                       size = 30, color = "red")
+  lims[[length(lims)+1]]<- i1
+}
+out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]],lims[[3]])),
+                          image_append(c(lims[[4]],lims[[5]]))),
+                        stack = TRUE)
+plot(out.img)
 @
-\end{minipage}
-Strangely, the problem doesn't occur for text which follows the "end" of the 
-singlespacing commmand, but text which comes before the 'begin' 
-singlespacing command is affected, as can be seen here.
+%
+To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load \code{nodes} and \code{edges} dataframes.
+
 \end{document}

### 2020-12-10 article_rvTH16.Rnw

@@
 \usepackage{xcolor,caption, float}
+\usepackage{array}
 
@@
 \newcommand{\fct}[1]{\code{#1()}}
+
+%% This modifies which proportion of the page must be occupied by the figure
+%% or table to take the full page (with no additional text).
+%% Check if the journal does not have any rule about not changing this.
+\renewcommand{\floatpagefraction}{.8}
+
 % \setlist[itemize]{noitemsep, topsep=0pt}
@@
 library("igraph")
-opts_chunk$set(prompt=TRUE, fig.pos = 'H', results='asis', highlight=FALSE)
+opts_chunk$set(prompt=TRUE, fig.pos = '!htbp', results='asis', highlight=FALSE,
+               class.source = "bg-success")
 options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
@@
 opts_knit$set(concordance=TRUE, self.contained=FALSE, background = "#FFFFFF")
-# knitr::opts_chunk$set(fig.pos = 'H')
+# knitr::opts_chunk$set(fig.pos = '!htbp')
 loadfonts(device = "win")
@@
 \begin{document}
+%\SweaveOpts{concordance=TRUE}
 % \SweaveOpts{concordance=TRUE}
@@
 
-\smallbreak
-<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center",fig.asp=0.1, fig.width=10, fig.cap="\\label{fig:fart}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.25, fig.height=0.25, fig.cap="\\label{fig:fart}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 par(mar=c(0.1,0.1,0.1,0.1) )
@@
      vertex.frame.color="white",
-     vertex.size=80,
+     vertex.size=40,
      vertex.label.cex=1,
      vertex.label.color="black",
-     vertex.label.family="Courier New",
-     edge.arrow.size=.5)
+     vertex.label.family="Courier",
+     edge.arrow.size=0.5,
+     xlim=c(-0.94,0.94), 
+     ylim=c(-1,-1))
 @
@@
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=2, fig.height=2, fig.cap="\\label{fig:fratar}Potential spatial relations between graphical units">>=
-par(mar=c(0.1,0.1,0.1,0.1) )
+<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.4, fig.height=1.4, fig.cap="\\label{fig:fratar}Potential spatial relations between graphical units">>=
+par(mar=c(0,0,0,0))
 ugs <- data.frame(name=c("a","r","t"))
@@
      vertex.label.color="black",
-     vertex.label.family="Courier New"
+     vertex.label.family="Courier",
+     xlim=c(-0.65,0.65), 
+     ylim=c(-0.65,0.65)
 )
@@
 \setlength\itemsep{.05em}
-\item unexplicit groupings of graphical units -- like graphical units grouped into \emp{figures},  \emp{figures} grouped into \emp{patterns}, \emp{patterns} grouped into \emp{motives}, etc.
+\item unexplicit groupings of graphical units -- like graphical units grouped into \emph{figures},  \emph{figures} grouped into \emph{patterns}, \emph{patterns} grouped into \emph{motives}, etc.
 -- with tedious number of groups
 \item consistency, proximities and relationships between these groups are often implicit and not quantified
-\item studies develop proper descriptive vocabularies, singular relationships of grouping, idosyncratic methods at site-dependend or period-dependend scales 
+\item studies develop proper descriptive vocabularies, singular relationships of grouping, idosyncratic methods at site-dependent or period-dependent scales
 \end{itemize}
@@
 These issues limit drastically the possibility to conduct cross-cultural comparisons and to draw a synthesis of humankind's symbolism at a large scale and over the long-term.
-\\
+
 According to us, prehistorical composition -- like any formal system -- are spatial features (mostly shapes) related one with the other depending on rules of spatial proximities.
@@
 To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory and particularly relative neighborhood graphs (RNG): '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}.
-\\
+
 In this article we present the \proglang{R} package \pkg{decorr}.
@@
 Their spatial relations can be modelled as edges.
-The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.\\
-% \smallbreak
+The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.
+
 Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
@@
 Centroids of the GUs (\code{POINTS}) are calculated and considered as the graph nodes, their locations (\emph{x} and \emph{y} coordinates, measured in pixels) are relative to each decoration.
-Exist edges (\code{LINES}) between pairs of GUs when their Voronoi cells share a common border (\emph{birel}: 'touches').\\
+Exist edges (\code{LINES}) between pairs of GUs when their Voronoi cells share a common border (\emph{birel}: 'touches').
+
 The entire graphical content of the decoration is mapped with a neighborhood graphs (RNG) -- ie geometric graph, or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
@@
 
-% \begin{figure}
-\begin{figure}[H]
+\begin{figure}[!htbp]
 \centering
@@
 \caption{\label{fig:gis} GIS interface.
-A) Original decoration of the Late Bronze Age \emp{Cerro Muriano 1} stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonsiation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
+A) Original decoration of the Late Bronze Age \emph{Cerro Muriano 1} stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonsiation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
 \end{figure}
@@
 At least one variable, like the \code{type} of the GUs ('anthropomorph', 'horned helmet', etc.), is defined in order to compute composition analysis.
-Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs: \emph{normal}, \emph{attribute} and \emph{overlapping} edges.\\
-\\
-\textbf{\emph{normal}} edges
-\\
+Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs:
+
+\begin{description}
+\item[Normal edges]
 By convention, two contemporaneous GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='}.
@@
 For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Fig.
-\ref{fig:gis_1})
-\smallbreak
-\textbf{\emph{attribute}} edges
-\\
+\ref{fig:gis_1}).
+
+\item[Attribute edges]
 It occurs frequently that a GU can be divided into a \emph{main unit} (eg, an anthropomorph) and one or various \emph{attribute units} (eg, a helmet, a male sex).
@@
 
-\begin{figure}[H] 
+\begin{figure}[!htbp]
 \centering
-\includegraphics[height=10cm, keepaspectratio]{gis_1}
+\includegraphics[height=10cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
 \caption{\label{fig:gis_1} GIS interface, Cerro Muriano 1 decoration.
@@
 
-\smallbreak
-\textbf{\emph{overlapping}} edges
-\\
+\item[Overlapping edges]
 Finally, it is quite common that a graphical composition shows superimpositions between different GUs permitting to distinguish different decoration phases, or iconographic layers, for a single support.
@@
 
-\begin{figure}[H] 
+\end{description}
+
+\begin{figure}[!htbp]
 \centering
-\includegraphics[height=10cm, keepaspectratio]{gis_2}
+\includegraphics[height=10cm, trim=30mm 16mm 30mm 13mm, clip, keepaspectratio]{gis_2}
 \caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
@@
 
-These different kind of relationships are resumed with their birel equivalencies in the following table (Tab. \ref{table:tedges})
+These different relationship types are resumed with their birel equivalencies in Table \ref{table:tedges}.
 
-\begin{table}[H]
-% \begin{center}
+\begin{table}[!htbp]
 \centering
-\begin{tabular}{|p{.6cm} p{.5cm} p{.6cm} p{2cm} p{2cm} p{5cm}|}
+\begin{tabular}{cccccp{4.9cm}}
 \hline
 % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
-GU.1 & edge type & GU.2 & (un)directed & \emph{birel} & description \\
-\hline
-\code{A} & \code{=} & \code{B} & undirected & $ A \cap B = \emptyset $ & \code{A} and \code{B} are disjoint \code{A} and \code{B} contemporaneous \\
-\code{A} & \code{+} & \code{B} & directed & $ A \cap B = A $ & \code{A} and \code{B} are contemporaneous, \code{B} is an attribute of \code{A} \\
-\code{A} & \code{>} & \code{B} & directed & $ A \cap B = \exists $ & \code{A} overlaps \code{B}, \code{A} is more recent than \code{B} \\
+GU.1 & Edge type & GU.2 & Edge directness & \emph{Birel} & Description \rule{0pt}{1em} \\[1pt]
+\hline \rule{0pt}{1em}%
+\code{A} & \code{=} & \code{B} & Undirected & $ A \cap B = \emptyset$ & 
+  \code{A} and \code{B} are contemporaneous and disjoint. \\
+\code{A} & \code{+} & \code{B} & Directed & $ A \cap B = A $ & 
+  \code{A} and \code{B} are contemporaneous. \code{B} is an attribute of \code{A}. \\
+\code{A} & \code{>} & \code{B} & Directed & $ A \cap B = \exists $ & 
+  \code{A} is more recent than \code{B}.\newline \code{A} overlaps \code{B}. \\
 % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
@@
 \end{tabular}
-\caption{Synthesis for the different types of relations between GUs}\label{tab1}
-% \end{center}
+\caption{Synthesis for the different types of relations between GUs.}\label{tab1}
 \label{table:tedges}
@@
     align=c(
-                       "|p{0.01cm}|",
-                       "|p{0.5cm}|",
-                       "p{2.5cm}|",
-                       "p{2.5cm}|",
-                       "p{8.5cm}|")),
-      table.placement="H",
+      "p{0.01cm}",
+      "c",
+      ">{\\raggedright\\arraybackslash\\rule{0pt}{1em}\\addtolength{\\baselineskip}{-1.5pt}}p{2.1cm}",
+      ">{\\raggedright\\arraybackslash\\addtolength{\\baselineskip}{-1.5pt}}p{2.1cm}",
+      "p{8.5cm}")),
+    table.placement="!htbp",
     include.rownames=FALSE)
@@
 %
-<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=18, fig.height=13, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
+<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.height=11.2, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
 # library("magick")
@@
   i1 <- image_annotate(i1,lbl.txt,location = "northwest",
-                       size = 30, color = "red")
+                       size = 28, color = "red")
   lims[[length(lims)+1]]<- i1
@@
                         stack = TRUE)
+par(mar=c(0,0,0,0))
 plot(out.img)
@@
 <<tndss, results='asis', echo=TRUE>>=
-caption <- "Nodes (from \\code{nodes.csv} dataframe)}"
+caption <- "Nodes (from \\code{nodes.csv} dataframe)"
 print(xtable::xtable(head(nodes),
@@
                      label="tab:tnds"),
-      table.placement="H")
+      table.placement="!htbp")
 @
@@
 <<tedss, results='asis', echo=TRUE>>=
-caption <- "Edges (from \\code{edges.csv} dataframe)}"
+caption <- "Edges (from \\code{edges.csv} dataframe)"
 print(xtable::xtable(head(edges),
@@
                      label="tab:teds"),
-      table.placement="H")
+      table.placement="!htbp")
 @
@@
 
-<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.cap="\\label{fig:igraph}R interface. Plot of the first graph of the list">>=
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
+Plot of the first graph of the list">>=
 lgrph <- list_dec(imgs,nodes,edges)
+par(mar=c(0,0,0,0))
 plot(lgrph[[1]],
@@
      vertex.label.color = "black",
-     vertex.size = 8,
-     vertex.label.cex = 0.6,
+     vertex.size = 16,
+     vertex.label.cex = 1,
      edge.color = "orange",
@@
                      size=7),
-      table.placement="H")
+      table.placement="!htbp")
 @
@@
 
-<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
+<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
 # library("decorr")
@@
 %                      size=7),
-%       table.placement="H")
+%       table.placement="!htbp")
 % @
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.height=8, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
-# par(mar=c(0,0,0,0))
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
+par(mar=c(0,0,0,0))
 g.compar <- list_compar(lgrph,"type")
@@
                      digits = c(0)),
-      table.placement="H",
+      table.placement="!htbp",
       include.rownames = TRUE)
@@
 
-\begin{table}[H]
-\begin{minipage}{.45\textwidth}
+\begin{table}[!htbp]
 \centering
+\begin{minipage}{.3\textwidth}
+\centering
+\textbf{Common nodes}\raisebox{-1ex}{}\\
 <<echo=FALSE,results='asis' >>=
@@
                      digits = c(0)),
-      table.placement="H",
+      floating=FALSE,
       include.rownames = TRUE)
@@
 @
-\captionof{table}{Common nodes table}
 \end{minipage}
-\begin{minipage}{.45\textwidth}
+\begin{minipage}{.3\textwidth}
 \centering
+\textbf{Common edges}\raisebox{-1ex}{}\\
 <<echo=FALSE,results='asis'>>=
@@
 @
-\captionof{table}{Common edges table}
 \end{minipage}
+\caption{Tables showing the number of common nodes and common edges between each pair of decorations}
 \end{table}
@@
 
-\begin{figure}[H]
+\begin{figure}[!htbp]
 <<echo=TRUE, warning=FALSE>>=
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = 'H', fig.align="center", fig.cap="\\label{fig:figs}Decoration comparisons between 3 and 4">>=
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.cap="\\label{fig:figs}Decoration comparisons between 3 and 4">>=
 par(mar=c(0,0,0,0))

### 2020-12-11 article_rvTH16.Rnw
article

@@
                            lbl.size=2)
-plot(image_trim(img.graph))
+image_read(img.graph)
 @

### 2020-12-12 article_rvTH16.Rnw
up

@@
 But unlike many aspects of the material culture where technological requirements and technical efficiency determine the choice of the raw material and the shape of the object -- a flint blade for cutting, a pottery for containing, a house for living --, the function of an iconographic composition cannot be drawn directly from itself.
+Furthermore, there is no specific and controled vocabularies to describe hand-made images.
+There is always a difficulty to decribe an iconographical content with alphanumerical signs.
 Whether these last decades study of ancient iconography had  undergone significative improvements at the site scale -- with GIS/database statistics, paleoenvironmental restitutions, etc.
@@
 \setlength\itemsep{.05em}
-\item unexplicit groupings of graphical units -- like graphical units grouped into \emph{figures},  \emph{figures} grouped into \emph{patterns}, \emph{patterns} grouped into \emph{motives}, etc.
+\item unexplicit spatial groupings of graphical units -- like graphical units grouped into \emph{figures},  \emph{figures} grouped into \emph{patterns}, \emph{patterns} grouped into \emph{motives}, etc.
 -- with tedious number of groups
@@
 
+Furthermore, at the graphical units level, top-down hierachical typologies (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) lead also to tedious inventories and difficult the comparisons \emph{within} and \emph{between} the categories.
+
 These issues limit drastically the possibility to conduct cross-cultural comparisons and to draw a synthesis of humankind's symbolism at a large scale and over the long-term.
@@
 Then, the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70} should be more operative than the Saussurian '\emph{linearity of the signifier}' principle \citep{Saussure89}.
-To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory and particularly relative neighborhood graphs (RNG): '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}.
+To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory, which offers a general framework to manage neighbourhoods, and particularly relative neighborhood graphs (RNG) which is '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}.
+Furthermore, we address the questions of managing graphical units attributes (for example an anthropomorph with a sword \emph{vs} an anthropomorph with a spear) and superimpositions of graphical units.
 
@@
 
-Nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
+Graph elements, nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.
 Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
@@
 The entire graphical content of the decoration is mapped with a neighborhood graphs (RNG) -- ie geometric graph, or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
-So a decoration graph is a one component graph which is the dual graph of the Voronoi diagram (Fig.
+So, a decoration graph is a one component graph.
+This graph is the dual graph of the decoration Voronoi diagram (Fig.
 \ref{fig:gis}).
 
-\begin{figure}[!htbp]
+\begin{figure}[H]
+% \begin{figure}[!htbp]
 \centering
@@
 \caption{\label{fig:gis} GIS interface.
-A) Original decoration of the Late Bronze Age \emph{Cerro Muriano 1} stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonsiation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
+A) Original decoration of the Late Bronze Age Cerro Muriano 1 stelae (drawing: \cite{DiazGuardamino10}) opened in new GIS project without any projection system with its extent (\code{xmin, xmax, ymin, ymax} in pixels); B) Polygonisation of the GUs, calculation of their centroids (nodes), calculation of the Voronoi cells, calculation of the proximity links (edges); C) plot of the relative \emph{x} and \emph{y} coordinates of the GUs centroids; D) plot of the values in the \code{type} of the GUs}
 \end{figure}
@@
 Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\
-This model has a minimal of \emph{a priori} definitions.
-Those definitions only concern the intrinsic variables of GUs (\code{type}, \code{technologie}, \code{color}, \code{size}, etc.) and the types of relations they share.
-At least one variable, like the \code{type} of the GUs ('anthropomorph', 'horned helmet', etc.), is defined in order to compute composition analysis.
-Here, we will only consider one property for the nodes, its \code{type}, and the most common types of relations between GUs:
+
+\subsection[elements]{Graph elements} \label{sec:elements}
+
+This model has a minimal of \emph{a priori} definitions. Those definitions only concern the graph elements: nodes variables (ie, GUs variables) and types of relations they share (edges).
+
+\smallbreak
+\textbf{Nodes types}
+\\
+At least one variable, like the \code{type} of the nodes, is defined in order to analyse and compare iconographical compositions.
+In this article, we will only consider nodes \code{type} values ('anthropomorph', 'sword', etc.) but it the package user can add other variables (\code{technique}, \code{color}, etc.)
+
+\smallbreak
+\textbf{Egdes types}
+\\
+The package considers the most common types of relations between nodes:
 
 \begin{description}
-\item[Normal edges]
-By convention, two contemporaneous GUs having a Voronoi cell sharing a border, have a common edge tagged \code{'='}.
-This proximity is represented with a plain line.
-The textual notation of such an edge is \code{'-=-'}.
+\item[\emph{normal edges}]
+Two contemporaneous nodes having a Voronoi cell sharing a border have a common edge.
+By convention, this edge is tagged \code{'='} (\code{type = '='}), represented with an orange plain line and its textual notation is \code{'-=-'}.
 For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Fig.
 \ref{fig:gis_1}).
-
-\item[Attribute edges]
-It occurs frequently that a GU can be divided into a \emph{main unit} (eg, an anthropomorph) and one or various \emph{attribute units} (eg, a helmet, a male sex).
-Broadly speaking, for further statistical analysis, it is better to use this \emph{attribute} method than to multiply the categories of GUs.
-To record this information, a new type of edge, tagged with \code{'+'}, is be introduced.
-This type of edges is directed and, by convention, displayed with a dashed line.
-Its starts from the \emph{main unit} and ends with the \emph{attribute unit}.
-The textual notation of such an edge is \code{'-+-'}.
-For example \code{personnage -+- casque} means that the main node \code{personnage} has the attribute node \code{casque} (Fig.
+As said, in Prehistorical iconography, the read direction is not known, so this edge is undirected, ie \code{personnage -=- bouclier} is equal to \code{bouclier -=- personnage} (Fig.
 \ref{fig:gis_1}).
 
-\begin{figure}[!htbp]
-\centering
+\item[\emph{attribute edges}]
+It occurs frequently that a node can be divided into a main unit (eg, an anthropomorph) and one or various attribute units (eg, a helmet, a male sex).
+To record this ordered set, the edge between a main and an attribute node is tagged \code{'+'} (\code{type = '+'}), represented with an orange dashed line and its textual notation is \code{'-+-'}.
+For example \code{personnage -+- casque} means that the main node \code{personnage} has the attribute node \code{casque}.
+The edge is directed since \code{personnage -+- casque} (the anthropomorph has an helmet) is not equal to \code{casque -+- personnage} (the helmet has an anthropomorph) (Fig.
+\ref{fig:gis_1}).
+
+\begin{figure}[H]
+% \begin{figure}[!htbp]
+\centering\emph{}
 \includegraphics[height=10cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
 \caption{\label{fig:gis_1} GIS interface, Cerro Muriano 1 decoration.
-The GUs \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the GU \code{personnage} (anthropomorph).}
+The nodes \code{casque} (helmet) and \code{sexe\_masculin} (male sex) are two nodes attributes of the node \code{personnage} (anthropomorph).}
 \end{figure}
 
-\item[Overlapping edges]
-Finally, it is quite common that a graphical composition shows superimpositions between different GUs permitting to distinguish different decoration phases, or iconographic layers, for a single support.
-So, at first, the analyse must be performed on each different phases of decoration separately.
-This stratigraphical information (\code{A} \emph{over} \code{B}, or \code{B} \emph{under} \code{A}) helps to understand the relative chronology between GUs and must be recorded.
-A simple way to achieve this is to introduce the new tag \code{'>'} for the for the type of edge.
-This type of edges is directed.
-The textual notation of such an edge is \code{'->-'}.
-For example \code{ecriture ->- bouclier} means that \code{ecriture} crosses \code{bouclier} (ie, \code{ecriture} overlaps \code{bouclier} in the stratigraphical sense) (Fig.
+Broadly speaking, for further statistical analysis, it is better to use this \emph{attribute} method than to multiply the categories of nodes.
+
+\item[\emph{overlapping edges}]
+Finally, it occurs that a graphical composition shows decoration phases.
+These decoration phases can be identified by observing the superimpositions between different nodes or by other means (the technique or the patina for example).
+In the first case, two nodes belonging to two different phase will have a common part (the overlapped intersection) while in the second case the two nodes will not necessarly share a part of their surface.
+To record this relative chronology, an edge between two non-contemporaneous nodes is tagged \code{'>'} (\code{type = '>'}), represented with a blue plain line and its textual notation is \code{'->-'}.
+For example \code{ecriture ->- bouclier} means that \code{ecriture} is more recent than the \code{bouclier}, or \emph{overlaps} the\code{bouclier} in the stratigraphical sense.
+The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier} ) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Fig.
 \ref{fig:gis_2}).
@@
 
-\begin{figure}[!htbp]
+\begin{figure}[H]
+% \begin{figure}[!htbp]
 \centering
 \includegraphics[height=10cm, trim=30mm 16mm 30mm 13mm, clip, keepaspectratio]{gis_2}
-\caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The GUs \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
+\caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The node \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
 \end{figure}
+
+This stratigraphical information (node 1 \emph{over} node 2, or node 2 \emph{under} node 1) helps to understand the relative chronology between node and must be recorded.
+The analysis of the iconographical content can be stratified (stratified analysis) and performed on each different phases of decoration separately.
 
@@
 \centering
-\begin{tabular}{cccccp{4.9cm}}
+\begin{tabular}{ccccp{2cm}p{4.9cm}}
 \hline
 % node A & type of edge & node B & undirected & birel & stratigraphical meaning \\
-GU.1 & Edge type & GU.2 & Edge directness & \emph{Birel} & Description \rule{0pt}{1em} \\[1pt]
+Node 1 & Edge type & Node 2 & Edge directness & \emph{Birel} & Description \rule{0pt}{1em} \\[1pt]
 \hline \rule{0pt}{1em}%
-\code{A} & \code{=} & \code{B} & Undirected & $ A \cap B = \emptyset$ & 
+\code{A} & \code{=} & \code{B} & Undirected & $ A \cap B = \emptyset $ & 
   \code{A} and \code{B} are contemporaneous and disjoint. \\
@@
   \code{A} and \code{B} are contemporaneous. \code{B} is an attribute of \code{A}. \\
-\code{A} & \code{>} & \code{B} & Directed & $ A \cap B = \exists $ & 
-  \code{A} is more recent than \code{B}.\newline \code{A} overlaps \code{B}. \\
+\code{A} & \code{>} & \code{B} & Directed & $ A \cap B = \exists $ \newline or \newline $ A \cap B = \emptyset $ & 
+  \code{A} is more recent than \code{B}. \\
 % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
@@
 <<results='asis', echo=FALSE, warning=FALSE>>=
-# library("decorr")
-# library("igraph")
-# library("magick")
-# # par(mar=c(0,0,0,0))
 imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
@@
 <<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.height=11.2, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
-# library("magick")
 pth <- system.file("extdata", package = "decorr")
@@
 To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load \code{nodes} and \code{edges} dataframes.
-%
-<<echo=TRUE>>=
-nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
-edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
-@
-%
 % \begin{itemize}
@@
 \smallbreak
-\textbf{\code{nodes} dataframe}
+\textbf{Nodes dataframe}
 \\
@@
 caption <- "Nodes (from \\code{nodes.csv} dataframe)"
+nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
+                    sep=";",stringsAsFactors = FALSE)
 print(xtable::xtable(head(nodes),
@@
 The \code{nodes$type} field is the default variable for further statistical analysis.
-Here, \code{nodes$type} refers to the typology of the GUs (anthropomorph, weapons, etc.).
+Here, \code{nodes$type} refers to the typology of the nodes (anthropomorph, weapons, etc.).
 The \code{nodes$x} and \code{nodes$y} columns refer to the \emph{x} and \emph{}y coordinates of the nodes.
@@
 But, in a GIS, the coordinates origin (0, 0) is the bottom-left corner and exist negative values, while this origin is top-left for any \proglang{R} matrices (rasters, grids, dataframes, etc.) with only positive values.
-To recover the local \emph{y} value of GUs nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
+To recover the local \emph{y} value of nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
 \smallbreak
@@
 \\
-The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Tab. \ref{tab:teds}).
+The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Table \ref{tab:teds}).
 
 <<tedss, results='asis', echo=TRUE>>=
-caption <- "Edges (from \\code{edges.csv} dataframe)"
-print(xtable::xtable(head(edges),
+caption <- "Edges of Cerro Muriano 1 (from \\code{edges.csv} dataframe)"
+edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
+                    sep=";",stringsAsFactors = FALSE)
+edges.cm1 <- subset(edges, decor == "Cerro Muriano 1")
+print(xtable::xtable(head(edges.cm1),
                      caption=caption,
@@
 The fields \code{edges$a} and \code{edges$b} are the equivalent to columns \emph{from} and \emph{to} in Graph theory.
-Even if undirected graphs will the most common models in further studies, this direction helps to distinguish between nodes.
-The first column \code{edges$a} is the identifier of starting node or \emph{main node} or \emph{overlapping node}.
-The second column, \code{edges$b}, is the identifier of the ending node or \emph{attribue node} or \emph{overlapped node}.
-The \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the starting node and the ending node.
+Even if undirected graphs will the most common in further studies, this direction helps to distinguish between nodes.
+The column \code{edges$a} is the identifier of \emph{starting node} or \emph{main node} or \emph{overlapping node}.
+The column \code{edges$b} is the identifier of the \emph{ending node} or \emph{attribue node} or \emph{overlapped node}.
+The column \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the node a and the node b.
 There is no need to record the coordinates of the edges, these coordinates are calculated from the \code{nodes} dataframe.
-For example, Table \ref{teds} shows that the Cerro Muriano 1 decoration's first edge connects the nodes \code{1} and \code{8} (respectively in column \code{edges$a} and \code{edges$b}).
-A way to retrieve these connected nodes' coordinates will be:
 
-<<results='asis', echo=TRUE>>=
-cm.1 <- subset(nodes, decor == "Cerro Muriano 1" & id == 1)[,c("x","y")]
-cm.8 <- subset(nodes, decor == "Cerro Muriano 1" & id == 8)[,c("x","y")]
-paste(as.numeric(cm.1),";",as.numeric(cm.8))
-@
-
-Once these dataframes loaded, the list of decoration graphs can be calculated with the \code{list_dec()} function.
-
-\subsection{\code{list\_dec()} function} \label{sec:funlistdec}
-
-The \code{list\_dec()} function allows to calculate graphs for all decorations stored into \code{nodes}, \code{edges} and \code{images}.
-The result is a list of decoration graph.
-The first graph of the list can be plotted (Fig.
-\ref{fig:igraph})
-
-<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
-Plot of the first graph of the list">>=
-lgrph <- list_dec(imgs,nodes,edges)
-par(mar=c(0,0,0,0))
-plot(lgrph[[1]],
-     vertex.color = "orange",
-     vertex.frame.color="orange",
-     vertex.label.color = "black",
-     vertex.size = 16,
-     vertex.label.cex = 1,
-     edge.color = "orange",
-     vertex.label.family="Helvetica"
-)
-@
-
-The others \pkg{decorr} package functions can be divided into:
+The \pkg{decorr} package functions can be divided into:
 \begin{enumerate}
+\item management functions
 \item graphical functions
@@
 
+Once these dataframes loaded, the list of decoration graphs can be calculated with \code{list_dec()}.
+
+\subsection{Management functions} \label{sec:manage}
+
+\end{Begin}
+list_dec(imgs,
+         nodes,
+         edges)
+\end{Code}
+
+Once \code{imgs}, \code{nodes} and \code{edges} dataframes loaded, the \code{list\_dec()} function allows to create graphs for all decorations within these dataframes.
+The result is a list of \code{igraph} objects.
+As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function (Fig.
+\ref{fig:igraph})
+
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
+Plot of the first graph of the list (Cerro Muriano 1 decoration)">>=
+lgrph <- list_dec(imgs, nodes, edges)
+par(mar=c(0, 0, 0, 0))
+plot(lgrph[[1]],
+     vertex.color = "orange",
+     vertex.frame.color = "orange",
+     vertex.label.color = "black",
+     vertex.size = 16,
+     vertex.label.cex = 1,
+     edge.color = "orange",
+     vertex.label.family = "Helvetica"
+)
+@
+
+\begin{Code}
+named_elements(grph,
+               focus = "edges",
+               nd.var = "type")
+\end{Code}
+
+The \code{named\_elements()} function allows to calculate the textual expression of graph elements (nodes and edges) with a disambiguation of these elements -- adding '#' suffixes -- when several nodes or edges have the same values.
+
+<<named, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
+Plot of the first graph of the list (Cerro Muriano 1 decoration)">>=
+lgrph <- list_dec(imgs, nodes, edges)
+as.character(named_elements(lgrph[[2]], focus = "edges", nd.var="type"))
+as.character(named_elements(lgrph[[2]], focus = "nodes", nd.var="type"))
+@
+
 \subsection{Graphical functions} \label{sec:functions_gr}
 
-The \pkg{decorr} package has three graphical functions
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{labels_shadow()} function is a re-use of the \code{shadowtext()} function from the \pkg{TeachingDemos} package \citep{Snow20}.
-\item \code{side_plot_nds()} and \code{side_plot_eds()} allow to plot figures side-by-side for nodes or edges comparisons
-\end{itemize}
+The \pkg{decorr} package has two graphical functions.
+
+\begin{Code}
+labels_shadow(x, y = NULL, labels,
+              col = "black", bg = "white",
+              theta = seq(0, 2 * pi, length.out = 50),
+              r = 0.1,
+              cex = 1, ...)
+\end{Code}
+
+\code{labels_shadow()} is a re-use of the \code{shadowtext()} function from the \pkg{TeachingDemos} package \citep{Snow20}.
+It plots labels (\code{labels}) at specific coordinates (\code{x, y}) with a contrasting buffer to make them more visible.
+
+\begin{Code}
+side_plot(grp, doss, nd.var, focus = "nodes",
+          nd.color = c("orange", "red"),
+          nd.size = c(0.5, 1),
+          ed.color = c("orange", "red"),
+          ed.width = c(1, 2),
+          lbl.size = 0.5)
+\end{Code}
+
+\code{side_plot()} allows to plot two decoration graphs side-by-side for elements (nodes or edges) comparisons.
+It take a list of two graph (\code{grp}), the path of the working directory (\code{doss}) and the name of the studied node variable (\code{nd.var}) and return an image.
 
@@
 
-Functions allowing to create a RNG for a single decoration are:
+Functions allowing to create a RNG for a single decoration are \code{read\_nds()} and \code{read\_eds()}.
+These functions allow to read respectively a file of nodes and a file of edges (\code{.csv}, \code{.tsv} or \code{.shp} files).
 
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{read_nds()} and \code{read_eds()} functions allow to read respectively a file of nodes and a file of edges (\code{.tsv} or \code{.shp} files)
-\end{itemize}
+\begin{Code}
+read_nds(site,
+         decor,
+         doss = getwd(),
+         nodes = "nodes",
+         dev = ".tsv")
+\end{Code}
 
-The \code{read_nds()} function is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
-For example, for the Torrejon Rubio 1 decoration, the 6th GUs is a sword (\code{epee}) (Table \ref{nd}).
+\code{read_nds()} is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
+For example, for the Torrejon Rubio 1 decoration, the node 6 is a sword (\code{epee}) (Table \ref{tab:nd}).
 
-<<nd, results='asis', echo=TRUE, fig.cap="Sixth node of the Torrejon Rubio 1 decoration, the sword">>=
-# library(decorr)
+<<nd, results='asis', echo=TRUE>>=
+caption <- "Torrejon Rubio 1 decoration node 6: the sword"
 sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
@@
                    doss = system.file("extdata", package = "decorr"))
-print(xtable::xtable(nds.df[6,],
-                     # caption="Sixth node (GU 6) of the Torrejon Rubio 1 decoration",
-                     label="nd",
-                     size=7),
-      table.placement="!htbp")
+print(xtable::xtable(nds.df[nds.df$id == 6, ],
+                     caption = caption,
+                     label = "tab:nd",
+                     size = 7),
+      table.placement = "!htbp")
 @
 
-The \code{read_eds()} permits to read a \emph{shapefiles} of edges or to retrieve the coordinates of the the edges from the \code{nodes} dataframe (Table \ref{ed}).
+\begin{Code}
+read_eds(site,
+         decor,
+         doss = getwd(),
+         edges = "edges",
+         nodes = "nodes",
+         dev = ".tsv")
+\end{Code}
+
+\code{read_eds()} permits to read a \emph{shapefiles} of edges (\code{dev = ".shp"}) or, when the data source is a \code{.csv} or \code{.tsv} dataframe, to retrieve the coordinates of the edges from the \code{nodes} dataframe (Table \ref{tab:ed}).
 
 <<ed, results='asis', echo=TRUE, fig.cap="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)">>=
+caption <- paste0("First edge of the Torrejon Rubio 1 decoration, ",
+                  "the proximity between the sword (node 6) and the fibula (node 5)")
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
@@
 print(xtable::xtable(eds.df[1,],
-                     # caption="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)",
-                     label="eds.df",
-                     size=7),
-      table.placement="!htbp")
+                     caption = caption,
+                     label = "tab:ed",
+                     size = 7),
+      table.placement = "!htbp")
 @
@@
 
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{plot_dec_grph()} allows to plot a RNG over a decoration image
-\end{itemize}
+% \begin{itemize}
+% \setlength\itemsep{.1em}
+% \item \code{plot_dec_grph()} allows to plot a RNG over a decoration image
+% \end{itemize}
 
-Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the decoration graph is build and can be plotted.
-The \code{lbl.txt} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field, but here it is the \code{nodes$type} field (Figure \ref{tr_st}).
+\begin{Code}
+plot_dec_grph(nodes = NULL,
+              edges = NULL,
+              site,
+              decor,
+              doss = getwd(),
+              nd.var = 'id',
+              nd.color = 'orange', nd.size = 1.7,
+              lbl.color = 'black', lbl.size = 1.2,
+              ed.color = c("orange", "blue"), ed.lwd = 4,
+              img.format = "png")
+\end{Code}
 
-<<tr_st, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:figs}Torrejon Rubio 1 decoration">>=
+\code{plot_dec_grph()} allows to plot a RNG over a decoration image.
+Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the graph of a specific decoration (\code{site} and \code{decor} parameters) is build and plotted.
+The \code{nd.var} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field.
+For example, we can plot the graph of the Torrejon Rubio 1 decoration with the type of the nodes (Figure \ref{fig:trst}).
+
+<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
 # library("decorr")
@@
                            decor = dec,
-                           doss = system.file("extdata", package = "decorr"),
+                           doss = system.file("extdata", 
+                                              package = "decorr"),
                            nd.var = "type",
                            lbl.size=2)
-image_read(img.graph)
+magick::image_read(img.graph)
 @
@@
 
-Functions allowing to compare different decorations with RNG are
+These functions allow to compare different graph decorations.
+The function \code{list_compar()} is used to compare common nodes and common edges depending on the node variable \code{nd.var}.
+This function calls \code{nds_compar()} and \code{eds_compar()}
 
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{list_nds_compar()} and \code{list_eds_compar()} functions allow to compare respectively the common nodes and the common edges between two decorations
-\end{itemize}
+\begin{Code}
+list_compar(lgrph,
+            nd.var = "type",
+            verbose = FALSE)
+\end{Code}
 
-Comparisons between pairwise of decorations are first stored into a list.
-These comparisons are performed for nodes and/or edges.
-There are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ pairwise comparisons
+Whereas the \code{nd.var} (by default, the column \code{type}) refers to the nodes and can be any other column create by the user (e.g., \code{technique}, \code{color}, etc.), the comparisons always take into account the field \code{type} (\code{=}, \code{+}, \code{>}, ...) of the edges.
+Comparisons are done between one or more pairwise of decorations.
+In the training dataset, there are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ possible pairwise comparisons
 
-% <<results='asis', echo=TRUE>>=
-% # set wd to data folder
-% # setwd(system.file("extdata", package = "decorr"))
-% # library("decorr")
-% g.compar <- list_eds_compar(lgrph,"type")
-% df.edges.compar <- data.frame(decor.A=c(g.compar[[1]][[1]]$decor,
-%                                        g.compar[[2]][[1]]$decor,
-%                                        g.compar[[3]][[1]]$decor,
-%                                        g.compar[[4]][[1]]$decor,
-%                                        g.compar[[5]][[1]]$decor,
-%                                        g.compar[[6]][[1]]$decor,
-%                                        g.compar[[7]][[1]]$decor,
-%                                        g.compar[[8]][[1]]$decor,
-%                                        g.compar[[9]][[1]]$decor,
-%                                        g.compar[[10]][[1]]$decor),
-%                               decor.B=c(g.compar[[1]][[2]]$decor,
-%                                        g.compar[[2]][[2]]$decor,
-%                                        g.compar[[3]][[2]]$decor,
-%                                        g.compar[[4]][[2]]$decor,
-%                                        g.compar[[5]][[2]]$decor,
-%                                        g.compar[[6]][[2]]$decor,
-%                                        g.compar[[7]][[2]]$decor,
-%                                        g.compar[[8]][[2]]$decor,
-%                                        g.compar[[9]][[2]]$decor,
-%                                        g.compar[[10]][[2]]$decor))
-% print(xtable::xtable(df.edges.compar,
-%                      caption="Pairwise comparisons dataframe between decor A and decor B",
-%                      label="Test_table_1",
-%                      size=7),
-%       table.placement="!htbp")
-% @
+\code{plot_compar()} allows to plot and save two figures side-by-side for one or more pairwise of decorations for common elements.
 
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{plot_nds_compar()} and \code{plot_eds_compar()} functions allow to plot and save two figures side-by-side for a decorations pairwise with, respectively, common nodes and common edges identified
-\end{itemize}
+\begin{Code}
+  plot_compar(listg, graph2 = NULL, focus = "nodes",
+              doss = getwd(),
+              nd.color = c("orange", "red"), nd.size = c(0.5, 1),
+              ed.color = c("orange", "red"), ed.width = c(1, 2),
+              lbl.size = 0.5, img.format = "png", res = 300)
+\end{Code}
+% \begin{itemize}
+% \setlength\itemsep{.1em}
+% \item \code{plot_nds_compar()} and \code{plot_eds_compar()} functions allow to plot and save two figures side-by-side for a decorations pairwise with, respectively, common nodes and common edges identified
+% \end{itemize}
+\code{plot_compar()} takes a list of graphs \code{listg}, created previously by \code{list_compar()}, and a vector with the ids of the graphs to compare (\code{graph2}).
+The function creates an image for each pairwise comparison with the decorations plotted side-by-side with elements: common nodes (\code{focus = "nodes"}, by default) or common edges (\code{focus = "edges"}).
+For each pairwise, if an element exists only in one decoration, it is displayed in orange by default (\code{nd.color[1]} or \code{ed.color[1]}).
+When an element is present on the two decorations, it is displayed in red by default (\code{nd.color[2]} or \code{ed.color[2]}).
+The function returns the paths to the created images.
 
-The \code{plot_nds_compar()}and \code{plot_eds_compar()} functions create a \code{.png} image of two or more decorations plotted side-by-side with common nodes or edges identified and returns the names of the images.
-The common edges or nodes are displayed in red by default.
-Let us choose the decorations \code{1} (Cerro Muriano 1) and \code{4} (Zarza de Montsanchez) and identify common edges.
+Let us choose the decorations 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) and identify common edges (Figure \ref{fig:ceds}).
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:figs}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:ceds}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
 par(mar=c(0,0,0,0))
@@
 
-\begin{itemize}
-\setlength\itemsep{.1em}
-\item \code{same_elements()} function allows to count matching nodes and matching edges between decoration pairwises
-\end{itemize}
+\code{same_elements()} function allows to count matching nodes and matching edges between decoration pairwises.
 
-\code{same_elements()} allows to count matching nodes and matching edges between decoration pairwises.
-The result is a square matrix between all pairwise comparisons with the number of common nodes or edges in the cells.
-For example, we can compute the matrix of common edges.
+\begin{Code}
+same_elements(lgrph,
+              nd.var = "type",
+              focus = "nodes")
+\end{Code}
+
+The \code{same_elements()} function take the list of graphs (\code{lgrph}) produced by \code{list_dec()} and the node variable on which the comparison will be done, by default \code{type} (\code{nd.var = "type"}).
+The result is a square matrix between all pairwise comparisons with the number of common elements (nodes or edges in the cells).
+The diagonal of the square matrix show the total number of element for a given decoration.
+For example, we can compute the matrix of common edges (Table \ref{tab:mat.ed}).
 
@@
 df.same_edges <- same_elements(lgrph, "type", "edges")
-caption <- "Number of same edges between all decoration pairwise comparisons"
+caption <- "Number of common edges between all graph decoration pairwises"
 print(xtable::xtable(df.same_edges,
                      caption = caption,
-                     label = "Test_table_2",
+                     label = "tab:mat.ed",
                      size = 8,
@@
 
-For these two last exemples, the edges comparisons between the decoration \code{1} and the decoration \code{4} show that they have two (2) common edges.
+The edges comparisons between the decoration 1 and the decoration 4 show that they have two (2) common edges, see also the Figure \ref{fig:ceds}.
 
@@
 
-As said, since the precise location of the GUs is usualy not registred, the most commonly used method in statistical analysis on prehistorical iconography is the presence of common nodes.
+As said, since the precise location of the nodes is usualy not registred, the most commonly used method in statistical analysis on prehistorical iconography is the presence of common nodes.
 In order to demonstrate the first insight of a graph-based analysis of the decorations, we will compare two classifications, the first one based on the "classic" presence of common nodes, the second one based on the presence of common edges.
@@
 
-Decorations \code{3} and \code{4} share four (4) common GUs (\code{bouclier, epee, lance, miroir}) but  these GUs have different spatial organisations with only one common edge (\code{bouclier -=- lance}).
+Decorations \code{3} and \code{4} share four (4) common nodes (\code{bouclier, epee, lance, miroir}) but  these nodes have different spatial organisations with only one common edge (\code{bouclier -=- lance}).
 At the opposite, decorations \code{2} and \code{4} show more properly the same compositions.
@@
 
-In this example we propose the iconographical \code{nodes$type} (character, weapon, etc.) GUs as the studied variable, but the user of the package can create and choose any other study variable: color for a painting, technique of realisation, size, etc.
+In this example we propose the iconographical \code{nodes$type} (character, weapon, etc.) nodes as the studied variable, but the user of the package can create and choose any other study variable: color for a painting, technique of realisation, size, etc.
 \code{edges$type} can also be extended to other types than normal, attribute, overlapping.

### 2020-12-12 article_rvTH17.Rnw
new version

--- 2020-12-12 article_rvTH16.Rnw
@@
 
-<<ed, results='asis', echo=TRUE, fig.cap="First edge of the Torrejon Rubio 1 decoration, the proximity between the sword (GU 6) and the fibula (GU 5)">>=
+<<ed, results='asis', echo=TRUE >>=
 caption <- paste0("First edge of the Torrejon Rubio 1 decoration, ",
-                  "the proximity between the sword (node 6) and the fibula (node 5)")
+                  "the proximity between the sword (node 6) ",
+                  "and the fibula (node 5)")
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:ceds}comparisons between 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) decorations">>=
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:ceds}Comparisons between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
 par(mar=c(0,0,0,0))
-g.compar <- list_compar(lgrph,"type")
-eds_compar <- plot_compar(g.compar, c(1,4),
+g.compar <- list_compar(lgrph, "type")
+eds_compar <- plot_compar(g.compar, c(1, 4),
                           focus = "edges",
@@
 
-As said, since the precise location of the nodes is usualy not registred, the most commonly used method in statistical analysis on prehistorical iconography is the presence of common nodes.
-In order to demonstrate the first insight of a graph-based analysis of the decorations, we will compare two classifications, the first one based on the "classic" presence of common nodes, the second one based on the presence of common edges.
+In order to demonstrate the main insight of a graph-based analysis of the decorations, we will compare two classifications, the first one based on the "classic" presence of common nodes -- since the precise location of the nodes is usualy not registred in the studies --, the second one based on the presence of common edges.
+
+At first, the Late Bronze Age phase of Ibahernando stelae must be selected (ie, the Roman latin writing must be removed).
+After the selection of the graph elements of Ibahernando, the connected component of the node 2, the lance (spear), is selected with \code{contemp_nds()} and the list of graphs (\code{lgrph}) is updated
+
+<<results='asis', echo=TRUE, fig.height= 5>>=
+dataDir <- system.file("extdata", package = "decorr")
+# Read a decoration
+nds.df <- read_nds(site = "Ibahernando",
+                   decor = "Ibahernando",
+                   doss = dataDir)
+
+eds.df <- read_eds(site = "Ibahernando",
+                   decor = "Ibahernando",
+                   doss = dataDir)
+# Extract the graph contemporaneous to the node 2
+l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 2)
+# remove first co
+l_dec_df[[1]]$site <- l_dec_df[[1]]$decor <- NULL
+l_dec_df[[2]]$site <- l_dec_df[[2]]$decor <- NULL
+g <- graph_from_data_frame(l_dec_df[[2]], vertices = l_dec_df[[1]],
+                           directed = FALSE)
+# re-add the site and decor columns 
+# with Ibahernando values and the graph name
+g$site <- g$decor <- "Ibahernando"
+g$name <- 5
+# overwrite the decor in the list
+lgrph[[5]] <- g
+@
+
+Then, \code{same_elements()} is done on nodes and edges (Table \ref{tab:common.elts})
+
+<<results='asis', echo=TRUE>>=
+df.same_edges <- same_elements(lgrph, "type", "edges")
+df.same_nodes<- same_elements(lgrph, "type", "nodes")
+@
 
 \begin{table}[!htbp]
+% \begin{table}[H]
 \centering
@@
 <<echo=FALSE,results='asis' >>=
-library("xtable")
-# library("MASS");library(decorr);library(xtable);library(magick)
-# imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-#                    sep="\t",stringsAsFactors = FALSE)
-# nodes <- read.table(system.file("extdata", "nodes.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# edges <- read.table(system.file("extdata", "edges.tsv", package = "decorr"),
-#                     sep="\t",stringsAsFactors = FALSE)
-# lgrph <- list_dec(imgs,nodes,edges,"type")
-df.same_edges <- same_elements(lgrph, "type", "edges")
-df.same_nodes<- same_elements(lgrph, "type", "nodes")
-# library("xtable")
 print(xtable::xtable(df.same_nodes,
@@
       include.rownames = TRUE)
-# print(xtable(df.same_nodes, digits=c(0)),
-#       floating=FALSE,
-#       hline.after=NULL)
 @
@@
       include.rownames = TRUE)
-# print(xtable(df.same_edges, digits=c(0)),
-#       floating=FALSE,
-#       hline.after=NULL)
 @
@@
 \caption{Tables showing the number of common nodes and common edges between each pair of decorations}
+\label{tab:common.elts}
 \end{table}
 
-Once the heatmap matrices calculated, the native \code{dist()} and \code{hclust()} functions \citep{R19} are calculated from the inverse matrices with the function \code{inv()} of the \pkg{matlib} package \citep{Friendly20}
+% the inverse values of common nodes and common edges are calculated with the function \code{inv()} from the package \pkg{matlib} \citep{Friendly20}, and then 
+Once the heatmap matrices calculated, the distance matrices are calculated with the native \code{dist()} function \citep{R19} (Table \ref{tab:dist.elts}).
 
-\begin{figure}[!htbp]
-<<echo=TRUE, warning=FALSE>>=
+\begin{table}[!htbp]
+% \begin{table}[H]
+\centering
+\begin{minipage}{.4\textwidth}
+\centering
+\textbf{Distance matrix on nodes}\raisebox{-1ex}{}\\
+<<echo=FALSE,results='asis', warning=F >>=
 library("matlib")
+dist.nodes  <- dist(as.matrix(df.same_nodes))
+mat.nodes <- as.matrix(dist.nodes)
+mat.nodes[lower.tri(mat.nodes, diag = TRUE)] <- NA
+print(xtable::xtable(mat.nodes,
+                     size = 8,
+                     digits = c(2)),
+      floating=FALSE,
+      include.rownames = TRUE)
+@
+\end{minipage}
+\begin{minipage}{.4\textwidth}
+\centering
+\textbf{Distance matrix on edges}\raisebox{-1ex}{}\\
+<<echo=FALSE,results='asis'>>=
+dist.edges <- dist(as.matrix(df.same_edges))
+mat.edges <- as.matrix(dist.edges)
+mat.edges[lower.tri(mat.edges, diag = TRUE)] <- NA
+print(xtable::xtable(mat.edges,
+                     size = 8,
+                     digits = c(2)),
+      floating=FALSE,
+      include.rownames = TRUE)
+@
+\end{minipage}
+\caption{Distance matrices on nodes and common edges between each pair of decorations}
+\label{tab:dist.elts}
+\end{table}
+
+For both nodes and edges, the results show that the most important differences are between the decoration 1 and the decoration 2: for nodes, dist = \Sexpr{round(max(mat.nodes, na.rm = T), 2)} ; for edges, dist =  \Sexpr{round(max(mat.edges, na.rm = T), 2)}.
+At the opposite, the decorations showing the more similar compositions are decoration 3 and decoration 5: for nodes, dist = \Sexpr{round(min(mat.nodes, na.rm = T), 2)} ; for edges, dist =  \Sexpr{round(min(mat.edges, na.rm = T), 2)}.
+A hierachical clustering (HC) of the distance matrices, with the native function \code{hclust()} \citep{R19}, summarize these similarites/dissimilarities (Figure \ref{fig:hclust})
+
+\begin{figure}[H]
+% \begin{figure}[!htbp]
+<<echo=TRUE, warning=FALSE, fig.height=5>>=
 par(mfrow=c(1,2))
-dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
-dist.edges  <- dist(inv(as.matrix(df.same_edges)))
-plot(hclust(dist.nodes), hang = -1, main = "common nodes")
-plot(hclust(dist.edges), hang = -1, main = "common edges")
+# dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
+# dist.edges  <- dist(inv(as.matrix(df.same_edges)))
+# plot(hclust(dist.nodes), hang = -1, main = "common nodes")
+# plot(hclust(dist.edges), hang = -1, main = "common edges")
+plot(hclust(dist.nodes), main = "common nodes")
+plot(hclust(dist.edges), main = "common edges")
 @
+\label{fig:hclust}
 \end{figure}
 
-Results of classifications show that for both common nodes and common edges, the most different decorations are \code{1} and \code{4}.
-These two decorations share four (4) common nodes and, as previously seen, only two (2) common edges.
-In any cases decorations \code{2} and \code{3} are closer to decoration \code{4} than to decoration \code{1}, but their classifications changes depending on counting of common nodes or common edges.
-While decorations \code{2} and \code{4} have 7 common edges and 7 common nodes, plotting the comparisons for \code{3} and \code{4}, helps to understand the differences between the two classifications.
+These latter decorations (decoration 1 and decoration 2) can be plotted with \code{plot_compar()} to vizualise the common elements (Figure \ref{fig:dec1.dec2}).
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.cap="\\label{fig:figs}Decoration comparisons between 3 and 4">>=
-par(mar=c(0,0,0,0))
-par(mfrow=c(2,1))
-g.compar <- list_compar(lgrph,"type")
-nds_compar.3.4 <- plot_compar(g.compar, c(3,4),
+\begin{figure}[H]
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.cap="\\label{fig:dec1.dec2}Decoration comparisons between decoration 1 and decoration 2">>=
+par(mar = c(0, 0, 0, 0))
+par(mfrow = c(2, 1))
+lgrph <- list_dec(imgs, nodes, edges)
+g.compar <- list_compar(lgrph, "type")
+nds_compar <- plot_compar(g.compar, c(3, 5),
                           focus = "nodes",
-                              doss = system.file("extdata", package = "decorr"))
-plot(magick::image_read(nds_compar.3.4))
-# g.compar <- list_eds_compar(lgrph,"type")
-eds_compar.3.4 <- plot_compar(g.compar, c(3,4),
+                          nd.size = c(0.5, 1.5),
+                          ed.width = c(1, 2.5),
+                          lbl.size = .7,
+                          doss = system.file("extdata", 
+                                             package = "decorr"))
+plot(magick::image_read(nds_compar))
+eds_compar <- plot_compar(g.compar, c(3, 5),
                           focus = "edges",
-                              doss = system.file("extdata", package = "decorr"))
-plot(magick::image_read(eds_compar.3.4))
+                          nd.size = c(0.5, 1.5),
+                          ed.width = c(1, 2.5),
+                          lbl.size = .7,
+                          doss = system.file("extdata", 
+                                             package = "decorr"))
+plot(magick::image_read(eds_compar))
 @
+\label{fig:dec1.dec2}
+\end{figure}
 
-Decorations \code{3} and \code{4} share four (4) common nodes (\code{bouclier, epee, lance, miroir}) but  these nodes have different spatial organisations with only one common edge (\code{bouclier -=- lance}).
-At the opposite, decorations \code{2} and \code{4} show more properly the same compositions.
+Even the classifications on nodes and edges give the same results,
 

### 2020-12-13 article_rvTH17.Rnw
version17

@@
 %% - \AND starts a new line, \And does not.
-\author{Thomas Huet\\UMR 5140 \And Craig Alexander \And Jose Pozo}
+\author{Thomas Huet\\UMR 5140 \And Jose Pozo \And Craig Alexander}
 %   \And Second Author\\Plus Affiliation}
-\Plainauthor{Thomas Huet, Craig Alexander, Jose Pozo}
+\Plainauthor{Thomas Huet, Jose Pozo, Craig Alexander}
 
@@
 
-Furthermore, at the graphical units level, top-down hierachical typologies (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) lead also to tedious inventories and difficult the comparisons \emph{within} and \emph{between} the categories.
-
 These issues limit drastically the possibility to conduct cross-cultural comparisons and to draw a synthesis of humankind's symbolism at a large scale and over the long-term.
@@
 Then, the First Law of Geography: "\emph{everything is related to everything else, but near things are more related than distant things}" \citep{Tobler70} should be more operative than the Saussurian '\emph{linearity of the signifier}' principle \citep{Saussure89}.
-To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory, which offers a general framework to manage neighbourhoods, and particularly relative neighborhood graphs (RNG) which is '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}.
+To get these '\emph{near things}', or features proximities, we use concepts coming from the Graph Theory, which offers a general framework to manage neighbourhoods.
 Furthermore, we address the questions of managing graphical units attributes (for example an anthropomorph with a sword \emph{vs} an anthropomorph with a spear) and superimpositions of graphical units.
+% and particularly relative neighborhood graphs (RNG) which is '\emph{a powerful model of low-level visual process}' \citep{Toussaint80}
 
@@
 Their spatial relations can be modelled as edges.
-The different levels of GUs spatial organisation can be retrieve by a relative neighborhood graphs analysis (Graph Theory) and a spatial (GIS) analysis.
+The different levels of GUs spatial organisation can be retrieve by graphs analysis (Graph Theory) and a spatial (GIS) analysis.
 
 Graph elements, nodes and edges -- repectively GUs and connexions between GUs -- are created on a GIS interface.  
-Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates.
+Indeed, for large series of decorated supports, GIS offers the most suitable and flexible interface to register all GUs and to get their coordinates, makes easy the updating of GUs and GUs relation and opens to develop further spatial analysis.
 The decoration image is considerated as the basemap of the project and will cover the region of interest of the analysis (ie, all the graphical content).
 The decoration image can be binarized: GUs are considerated active, the undecorated parts of the support -- the background -- are considerated inactive.
-The polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, allows to calculated their Voronoi seeds (ie, their centroid) and Voronoi cells (ie, their area of influence).
+The polygonisation of the GUs (\code{POLYGONS}), including the border of the stelae, allows to calculated their Voronoi seeds (ie, their centroid) and Voronoi cells (ie, Thiessen polygons, their area of influence).
 So, the entire decoration image is tiled and condidered as a Voronoi diagram.
@@
 
-The entire graphical content of the decoration is mapped with a neighborhood graphs (RNG) -- ie geometric graph, or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
+The entire graphical content of the decoration is mapped with a geometric graph -- or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
 So, a decoration graph is a one component graph.
@@
 \subsection[elements]{Graph elements} \label{sec:elements}
-
-This model has a minimal of \emph{a priori} definitions. Those definitions only concern the graph elements: nodes variables (ie, GUs variables) and types of relations they share (edges).
 
@@
 
-\begin{figure}[H]
-% \begin{figure}[!htbp]
+% \begin{figure}[H]
+\begin{figure}[!htbp]
 \centering\emph{}
-\includegraphics[height=10cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
+\includegraphics[height=7cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
+% \includegraphics[height=10cm, trim=15mm 20mm 15mm 20mm, clip, keepaspectratio]{gis_1}
 \caption{\label{fig:gis_1} GIS interface, Cerro Muriano 1 decoration.
@@
 
-Broadly speaking, for further statistical analysis, it is better to use this \emph{attribute} method than to multiply the categories of nodes.
-
 \item[\emph{overlapping edges}]
 Finally, it occurs that a graphical composition shows decoration phases.
-These decoration phases can be identified by observing the superimpositions between different nodes or by other means (the technique or the patina for example).
+These decoration phases can be identified by observing the superimpositions between different nodes or by other means (differences in the techniques or in the patina for example).
 In the first case, two nodes belonging to two different phase will have a common part (the overlapped intersection) while in the second case the two nodes will not necessarly share a part of their surface.
@@
 For example \code{ecriture ->- bouclier} means that \code{ecriture} is more recent than the \code{bouclier}, or \emph{overlaps} the\code{bouclier} in the stratigraphical sense.
-The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier} ) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Fig.
+The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier}) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Fig.
 \ref{fig:gis_2}).
@@
 \centering
-\includegraphics[height=10cm, trim=30mm 16mm 30mm 13mm, clip, keepaspectratio]{gis_2}
+\includegraphics[height=8cm, trim=30mm 16mm 30mm 13mm, clip, keepaspectratio]{gis_2}
 \caption{\label{fig:gis_2} GIS interface, Ibahernando decoration. The node \code{ecriture} (writing) overlaps \code{lance} (spear) and the upper part of the \code{bouclier} (shield).}
@@
 
-These different relationship types are resumed with their birel equivalencies in Table \ref{table:tedges}.
+The nodes and edges equalities are resumed in the Table \ref{tab:ex.comm}.
 
-\begin{table}[!htbp]
+<<equalities, results='asis', echo=FALSE, fig.pos = '!htbp'>>=
+caption <- paste0("Example of equalities and differences for nodes and edges ")
+df.ex <- data.frame(type = c("node", "node", "edge", 
+                             "edge", "edge", "edge", "edge"),
+                    part.1 = c("epee", "epee", "epee --=-- lance",
+                               "epee --=-- lance", "personnage --=-- bouclier", 
+                               "personnage --+-- bouclier", "ecriture -->-- bouclier"),
+                    part.2 = c("epee", "lance", "epee --=-- lance",
+                               "lance --=-- epee", "personnage --+-- bouclier",
+                               "bouclier --+-- personnage", "bouclier -->-- ecriture"),
+                    is.equal = c("True", "False", "True", 
+                                 "False", "False", "False", "False"),
+                    stringsAsFactors = F)
+print(xtable::xtable(df.ex,
+                     caption = caption,
+                     label = "tab:ex.comm",
+                     size = 7),
+      include.rownames = FALSE,
+      table.placement = "!htbp")
+@
+
+These different relationship types are resumed with their \emph{birel} equivalencies and chronological assumptions in Table \ref{table:tedges}.
+
+% \begin{table}[!htbp]
+\begin{table}[H]
 \centering
@@
 
-A training dataset (nodes and edges coordinates, decoration images) is stored in the \code{extdata} folder of the \pkg{decorr} package
+A training dataset is stored in the \code{extdata} folder of the \pkg{decorr} package.
+It contains dataframes of nodes' coordinates, edges' coordinates, decorations' listing in different formats (\code{.tsv}, \code{.csv} or \emph{shapefiles}) and the decoration drawings (\code{.jpg}) of five Late Bronze Age stelae belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- mostly concentrated in the Southwest Iberian peninsula \citep{Celestino01}.
+\\
+
 \smallbreak
-% \begin{itemize}
-% \item The \code{imgs} dataframe
-% \end{itemize}
-\textbf{\code{imgs} dataframe}
+\textbf{Imgs dataframe}
 \\
@@
 The primary key of each decoration is the concatenate of \code{imgs$site} and \code{imgs$decor}.
-These keys will allow joints with the other dataframes (\code{nodes} and \code{edges})
+These keys will allow joints with the other dataframes (nodes and edges)
 %
@@
 print(xtable::xtable(imgs,
-                     caption="The studied corpus, the \\code{imgs.tsv} dataframe",
-                     label="tab:timg",
+                     caption = "The studied corpus, the \\code{imgs.tsv} dataframe",
+                     label = "tab:timg",
                      size=6,
@@
 %
-The \pkg{decorr} package training dataset is composed by five stelaes decorations drawings \citep{DiazGuardamino10} belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- dated to the Late Bronze Age of SW Iberian peninsula \citep{Celestino01}.
-\\
-At first the drawings dataset can be checked by using the \code{imgs} dataframe and the \pkg{magick} package (Fig. \ref{fig:dataset})
+
+At first the drawings dataset \citep{DiazGuardamino10} can be checked by using the imgs dataframe and the \pkg{magick} package (Fig. \ref{fig:dataset})
 %
@@
   i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
-  lbl.txt <- paste0(imgs[i,"idf"],"\n",
-                    imgs[i,"site"],"\n",
-                    imgs[i,"decor"],"\n",
-                    imgs[i,"img"],"\n",
-                    image_info(i1)$width,"*",image_info(i1)$height," px")
-  i1 <- image_annotate(i1,lbl.txt,location = "northwest",
+  lbl.txt <- paste0(imgs[i,"idf"], "\n",
+                    imgs[i,"site"], "\n",
+                    imgs[i,"decor"], "\n",
+                    imgs[i,"img"], "\n",
+                    image_info(i1)$width, "*", image_info(i1)$height, " px")
+  i1 <- image_annotate(i1, lbl.txt, location = "northwest",
                        size = 28, color = "red")
@@
 }
-out.img <- image_append(c(image_append(c(lims[[1]],lims[[2]],lims[[3]])),
-                          image_append(c(lims[[4]],lims[[5]]))),
+out.img <- image_append(c(image_append(c(lims[[1]], lims[[2]], lims[[3]])),
+                          image_append(c(lims[[4]], lims[[5]]))),
                         stack = TRUE)
-par(mar=c(0,0,0,0))
+par(mar=c(0, 0, 0, 0))
 plot(out.img)
@@
 %
-To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load \code{nodes} and \code{edges} dataframes.
+To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load nodes and edges dataframes.
 % \begin{itemize}
@@
 nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
+                    sep=";", stringsAsFactors = FALSE)
 print(xtable::xtable(head(nodes),
                      caption = caption,
-                     label="tab:tnds"),
-      table.placement="!htbp")
+                     label = "tab:tnds"),
+      include.rownames = FALSE,
+      table.placement = "!htbp")
 @
@@
 % \end{itemize}
-\textbf{\code{edges} dataframe}
+\textbf{Edges dataframe}
 \\
-The \code{edges} dataframe is quite similar to the \code{nodes} dataframe (Table \ref{tab:teds}).
+The edges dataframe is quite similar to the nodes dataframe (Table \ref{tab:teds}).
 
@@
                      label="tab:teds"),
+      include.rownames = FALSE,
       table.placement="!htbp")
@@
 The column \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the node a and the node b.
-There is no need to record the coordinates of the edges, these coordinates are calculated from the \code{nodes} dataframe.
+There is no need to record the coordinates of the edges, these coordinates are calculated from the nodes dataframe.
 
@@
 \item single decoration functions
-\item comparisons between different decorations functions
+\item comparisons functions
 \end{enumerate}
@@
 
-\end{Begin}
+The management functions are \code{list\_dec()}, \code{contemp\_nds()} and \code{named\_elements()}
+
+\begin{Code}
 list_dec(imgs,
@@
 
-Once \code{imgs}, \code{nodes} and \code{edges} dataframes loaded, the \code{list\_dec()} function allows to create graphs for all decorations within these dataframes.
+Once imgs, nodes and edges dataframes loaded, the \code{list\_dec()} function allows to create graphs for all decorations within these dataframes.
 The result is a list of \code{igraph} objects.
@@
      vertex.label.color = "black",
-     vertex.size = 16,
-     vertex.label.cex = 1,
+     vertex.size = 14,
+     vertex.label.cex = .8,
      edge.color = "orange",
@@
 )
+@
+
+It occurs that some nodes are non-contemporaneous one with the others, like for the Ibahernando stelae.
+This stelae have been found reused as a funerary steale during Roman times \citep{Almagro66b}.
+A latin inscription (\code{ecriture}, node 1) have been carved over a spear (\code{lance}, node 2) and overlaps partially a V-notched shield (\code{bouclier}, node 3).
+The edges between node 1 and node 2, and the edge between node 1 and node 3, are \emph{overlap} edges (Figure \ref{fig:ibah}).
+
+<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:ibah}Ibahernando decoration, node 1 overlaps node 2 and node 3">>=
+par(mar=c(0,0,0,0))
+nds.df <- read_nds(site = "Ibahernando", 
+                   decor = "Ibahernando",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Ibahernando",
+                   decor = "Ibahernando",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+img.graph <- plot_dec_grph(nodes = nds.df,
+                           edges = eds.df,
+                           site = "Ibahernando",
+                           decor = "Ibahernando",
+                           lbl.size = 2,
+                           doss = system.file("extdata", 
+                                              package = "decorr"))
+plot(magick::image_read(img.graph))
+@
+
+In this case, the non-contemporaneous layers of decoration should be removed before the comparison process.
+To do so, the original graph (1-component) will be splitted into different sub-graphs (\emph{n}-component) by selecting \code{>} edges.
+The studied graph component will be retrieved with the component membership of a selected node in the \code{contemp_nds()} function parameters.
+
+To study only the Late Bronze Age iconographic layer of the Ibahernando steale, we can choose the Late Bronze Age node 4, the figure of sword (\code{epee}) dated to the middle and final stages of Late Bronze Age (ca 1250-950 BC).
+This node is supposed to be contemporaneous to node 2 and node 3 (Figure \ref{fig:ibah2}).
+
+<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=1, fig.cap="\\label{fig:ibah2}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
+par(mar=c(0,0,0,0))
+selected.nd <- 4
+nds.df <- read_nds(site = "Ibahernando", 
+                   decor = "Ibahernando",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Ibahernando",
+                   decor = "Ibahernando",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd)
+Ibahernando <- plot_dec_grph(nodes = nds.df,
+                             edges = eds.df,
+                             site = "Ibahernando",
+                             decor = "Ibahernando",
+                             nd.var = "type",
+                             lbl.color = "brown",
+                             lbl.size=2.2,
+                             doss = system.file("extdata", 
+                                                package = "decorr"))
+Ibahernando.img <- magick::image_read(Ibahernando)
+Ibahernando.contemp <- plot_dec_grph(nodes = l_dec_df[[1]],
+                                     edges = l_dec_df[[2]],
+                                     site = "Ibahernando",
+                                     decor = "Ibahernando",
+                                     nd.var = "type",
+                                     lbl.color = "brown",
+                                     lbl.size=2.2,
+                                     doss = system.file("extdata", 
+                                                        package = "decorr"))
+Ibahernando.contemp.img <- magick::image_read(Ibahernando.contemp)
+out <- magick::image_append(c(Ibahernando.img, Ibahernando.contemp.img))
+plot(out)
 @
@@
 
-<<named, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
-Plot of the first graph of the list (Cerro Muriano 1 decoration)">>=
-lgrph <- list_dec(imgs, nodes, edges)
-as.character(named_elements(lgrph[[2]], focus = "edges", nd.var="type"))
-as.character(named_elements(lgrph[[2]], focus = "nodes", nd.var="type"))
-@
+\begin{CodeChunk}
+\begin{CodeInput}
+R> lgrph <- list_dec(imgs, nodes, edges)
+R> as.character(named_elements(lgrph[[2]], focus = "edges", nd.var="type"))
+## [1] "epee-=-fibule" "bouclier-=-miroir"
+## [3] "epee-=-lance" "bouclier-=-chariot_char"
+## [5] "chariot_char-+-cheval" "chariot_char-+-cheval#"
+## [7] "arc-=-bouclier" "arc-=-chariot_char"
+## [9] "arc-=-miroir" "bouclier-=-fibule"
+## [11] "bouclier-=-epee" "epee-=-miroir"
+## [13] "arc-+-fleche" "chariot_char-+-roue"
+## [15] "chariot_char-+-roue#"
+R> as.character(named_elements(lgrph[[2]], focus = "nodes", nd.var="type"))
+## [1] "chariot_char" "bouclier" "arc" "miroir"
+## [5] "fibule" "epee" "lance" "cheval"
+## [9] "cheval#" "roue" "roue#" "fleche"
+\end{CodeInput}
+\end{CodeChunk}
 
@@
 
-The \pkg{decorr} package has two graphical functions.
+The \pkg{decorr} package has two graphical functions: \code{labels_shadow()} and \code{side_plot()}.
 
@@
 
-Functions allowing to create a RNG for a single decoration are \code{read\_nds()} and \code{read\_eds()}.
+Functions allowing to create a geometric graph for a single decoration are \code{read\_nds()} and \code{read\_eds()}.
 These functions allow to read respectively a file of nodes and a file of edges (\code{.csv}, \code{.tsv} or \code{.shp} files).
@@
 \code{read_nds()} is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
-For example, for the Torrejon Rubio 1 decoration, the node 6 is a sword (\code{epee}) (Table \ref{tab:nd}).
+For example, the nodes of the Torrejon Rubio 1 decoration can be read from the the \emph{shapefile}, and the 6\textsuperscript{th} node -- a sword (\code{epee}) -- displayed (Table \ref{tab:nd}).
 
 <<nd, results='asis', echo=TRUE>>=
-caption <- "Torrejon Rubio 1 decoration node 6: the sword"
+caption <- "Torrejon Rubio 1 decoration node 6: the sword."
 sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
@@
                      size = 7),
+      include.rownames = FALSE,
       table.placement = "!htbp")
@@
 
-\code{read_eds()} permits to read a \emph{shapefiles} of edges (\code{dev = ".shp"}) or, when the data source is a \code{.csv} or \code{.tsv} dataframe, to retrieve the coordinates of the edges from the \code{nodes} dataframe (Table \ref{tab:ed}).
+\code{read_eds()} permits to read a \code{.csv} or \code{.tsv} dataframe, or a \emph{shapefiles}, of edges.
+When the data source is a dataframe, the coordinates of the edges are retrieved from the nodes dataframe (Table \ref{tab:ed}).
 
@@
                      size = 7),
+      include.rownames = FALSE,
       table.placement = "!htbp")
 @
-% \end{figure}
 
-% \begin{itemize}
-% \setlength\itemsep{.1em}
-% \item \code{plot_dec_grph()} allows to plot a RNG over a decoration image
-% \end{itemize}
+Once, the imgs, nodes and edges dataframes have been read, a geometric graph can be build and plotted over a specific decoration image with the \code{plot_dec_grph()} function.
 
@@
 
-\code{plot_dec_grph()} allows to plot a RNG over a decoration image.
-Once, the \code{imgs}, \code{nodes} and \code{edges} dataframes have been read, the graph of a specific decoration (\code{site} and \code{decor} parameters) is build and plotted.
+The parameters \code{nodes} and \code{edges} concern respectively the nodes and edges dataframes.
+The parameters \code{site} and \code{decor} allow to choose a specific decoration.
 The \code{nd.var} parameter allows to decide which field of the nodes will be displayed as the label, by default this is the \code{nodes$id} field.
@@
 <<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
-# library("decorr")
-par(mar=c(0,0,0,0))
+par(mar=c(0, 0, 0, 0))
 sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
@@
                                               package = "decorr"),
-                           nd.var = "type",
-                           lbl.size=2)
-magick::image_read(img.graph)
+                           nd.var = "type", lbl.size = 2)
+out <- magick::image_read(img.graph)
+plot(out)
 @
@@
 
-These functions allow to compare different graph decorations.
-The function \code{list_compar()} is used to compare common nodes and common edges depending on the node variable \code{nd.var}.
-This function calls \code{nds_compar()} and \code{eds_compar()}
+These functions allow to compare different graph decorations. The function \code{list_compar()} is used to compare common nodes and common edges depending on the node variable \code{nd.var}.
 
@@
 Comparisons are done between one or more pairwise of decorations.
-In the training dataset, there are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ possible pairwise comparisons
+In the training dataset, there are five (5) decorations in the default dataset, so there is $\frac{5!}{(5-2)!2!}={10}$ possible pairwise comparisons.
+This function calls \code{nds_compar()} and \code{eds_compar()}.
 
-\code{plot_compar()} allows to plot and save two figures side-by-side for one or more pairwise of decorations for common elements.
+\begin{Code}
+nds_compar(grphs,
+           nd.var = "type")
+
+eds_compar(grphs,
+           nd.var = "type")
+\end{Code}
+
+The functions \code{nds_compar()} and \code{eds_compar()} take a pair of graphs \code{grphs} and return respectively the common nodes and the common edges based on the \code{nd.var} field of nodes (by default, the column \code{type}).
+
+\smallbreak
+The function \code{plot_compar()} allows to plot and save two figures side-by-side for one pairwise, or more, of decorations identifying common elements.
 
@@
 \end{Code}
-% \begin{itemize}
-% \setlength\itemsep{.1em}
-% \item \code{plot_nds_compar()} and \code{plot_eds_compar()} functions allow to plot and save two figures side-by-side for a decorations pairwise with, respectively, common nodes and common edges identified
-% \end{itemize}
+
 \code{plot_compar()} takes a list of graphs \code{listg}, created previously by \code{list_compar()}, and a vector with the ids of the graphs to compare (\code{graph2}).
@@
 
-Let us choose the decorations 1 (Cerro Muriano 1) and 4 (Zarza de Montsanchez) and identify common edges (Figure \ref{fig:ceds}).
+Let us compare the decoration 1 and the decoration 4 and identify common edges (Figure \ref{fig:ceds}).
 
@@
 
-The comparison shows that \code{1} (Cerro Muriano 1) and \code{4} (Zarza de Montsanchez) decorations have two (2) common edges: \code{lance -=- personnage} and \code{personnage -=- bouclier}.
+The comparison shows that the decoration 1 (Cerro Muriano 1) and the decoration 4 (Zarza de Montsanchez) have two (2) common edges: \code{lance -=- personnage} and \code{bouclier -=- personnage}.
 
-\code{same_elements()} function allows to count matching nodes and matching edges between decoration pairwises.
+The function \code{same_elements()} allows to count matching nodes and matching edges between decoration pairwises.
 
@@
 The result is a square matrix between all pairwise comparisons with the number of common elements (nodes or edges in the cells).
-The diagonal of the square matrix show the total number of element for a given decoration.
+The diagonal of the square matrix shows the total number of element for a given decoration.
 For example, we can compute the matrix of common edges (Table \ref{tab:mat.ed}).
@@
 
-In order to demonstrate the main insight of a graph-based analysis of the decorations, we will compare two classifications, the first one based on the "classic" presence of common nodes -- since the precise location of the nodes is usualy not registred in the studies --, the second one based on the presence of common edges.
+In order to demonstrate the main insight of a graph-based analysis of the decorations, we will compare two classifications.
+The first classification is based on the counts of common nodes.
+Since the precise location of the nodes is usualy not registred, his is the most commonly analysis applied on large series.
+The second classification is based on the presence of common edges.
 
 At first, the Late Bronze Age phase of Ibahernando stelae must be selected (ie, the Roman latin writing must be removed).
-After the selection of the graph elements of Ibahernando, the connected component of the node 2, the lance (spear), is selected with \code{contemp_nds()} and the list of graphs (\code{lgrph}) is updated
+After the selection of the graph elements of Ibahernando stelae (decoration 5), nodes and edges of the node 2 connected component, the lance (spear), is selected with \code{contemp_nds()}, a new graph object is created (\code{grph}) and used to update the list of graphs (\code{lgrph}).
 
@@
 dataDir <- system.file("extdata", package = "decorr")
-# Read a decoration
 nds.df <- read_nds(site = "Ibahernando",
@@
                    doss = dataDir)
-
 eds.df <- read_eds(site = "Ibahernando",
@@
                    doss = dataDir)
-# Extract the graph contemporaneous to the node 2
 l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 2)
-# remove first co
 l_dec_df[[1]]$site <- l_dec_df[[1]]$decor <- NULL
 l_dec_df[[2]]$site <- l_dec_df[[2]]$decor <- NULL
-g <- graph_from_data_frame(l_dec_df[[2]], vertices = l_dec_df[[1]],
+grph <- graph_from_data_frame(l_dec_df[[2]], vertices = l_dec_df[[1]],
                            directed = FALSE)
-# re-add the site and decor columns 
-# with Ibahernando values and the graph name
-g$site <- g$decor <- "Ibahernando"
-g$name <- 5
-# overwrite the decor in the list
-lgrph[[5]] <- g
+grph$site <- grph$decor <- "Ibahernando"
+grph$name <- 5
+lgrph[[5]] <- grph
 @
@@
 At the opposite, the decorations showing the more similar compositions are decoration 3 and decoration 5: for nodes, dist = \Sexpr{round(min(mat.nodes, na.rm = T), 2)} ; for edges, dist =  \Sexpr{round(min(mat.edges, na.rm = T), 2)}.
+The compared decorations, with their common elements, can be plotted with \code{plot_compar()}.
 A hierachical clustering (HC) of the distance matrices, with the native function \code{hclust()} \citep{R19}, summarize these similarites/dissimilarities (Figure \ref{fig:hclust})
 
-\begin{figure}[H]
-% \begin{figure}[!htbp]
-<<echo=TRUE, warning=FALSE, fig.height=5>>=
-par(mfrow=c(1,2))
-# dist.nodes  <- dist(inv(as.matrix(df.same_nodes)))
-# dist.edges  <- dist(inv(as.matrix(df.same_edges)))
-# plot(hclust(dist.nodes), hang = -1, main = "common nodes")
-# plot(hclust(dist.edges), hang = -1, main = "common edges")
-plot(hclust(dist.nodes), main = "common nodes")
-plot(hclust(dist.edges), main = "common edges")
+<<hclust, echo=TRUE, warning=FALSE, fig.height=5, fig.pos = '!htbp', fig.cap="\\label{fig:hclust}Hierachical clusterings on decoration common elements">>=
+par(mfrow=c(1, 2))
+hc1 <- hclust(dist.nodes, method = "ward.D")
+hc2 <- hclust(dist.edges, method = "ward.D") 
+plot(hc1, main = "common nodes")
+plot(hc2, main = "common edges")
 @
-\label{fig:hclust}
-\end{figure}
 
-These latter decorations (decoration 1 and decoration 2) can be plotted with \code{plot_compar()} to vizualise the common elements (Figure \ref{fig:dec1.dec2}).
+Even if in this example the clustering of decorations on common nodes and the clustering of decorations on common edges give the same results, another distribution of the same nodes, could give two different clusterings.
+The clustering on common nodes and the clustering on common edges can be compared with the \code{tanglegram()} function from the \pkg{dendextend} package \citep{Galili15} (Figure \ref{fig:hclust2})
 
-\begin{figure}[H]
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.cap="\\label{fig:dec1.dec2}Decoration comparisons between decoration 1 and decoration 2">>=
-par(mar = c(0, 0, 0, 0))
-par(mfrow = c(2, 1))
-lgrph <- list_dec(imgs, nodes, edges)
-g.compar <- list_compar(lgrph, "type")
-nds_compar <- plot_compar(g.compar, c(3, 5),
-                          focus = "nodes",
-                          nd.size = c(0.5, 1.5),
-                          ed.width = c(1, 2.5),
-                          lbl.size = .7,
-                          doss = system.file("extdata", 
-                                             package = "decorr"))
-plot(magick::image_read(nds_compar))
-eds_compar <- plot_compar(g.compar, c(3, 5),
-                          focus = "edges",
-                          nd.size = c(0.5, 1.5),
-                          ed.width = c(1, 2.5),
-                          lbl.size = .7,
-                          doss = system.file("extdata", 
-                                             package = "decorr"))
-plot(magick::image_read(eds_compar))
+<<hclust2, echo=TRUE, warning=FALSE, fig.height=3, fig.pos = '!htbp', fig.cap="\\label{fig:hclust2}Comparison of hierachical clusterings on decoration common elements, nodes (left) and edges (right)">>=
+suppressPackageStartupMessages(library(dendextend))
+suppressPackageStartupMessages(library(dplyr))
+dend1 <- as.dendrogram (hc1)
+dend2 <- as.dendrogram (hc2)
+dendlist(dend1, dend2) %>%
+  untangle(method = "step1side") %>% 
+  tanglegram()                       
 @
-\label{fig:dec1.dec2}
-\end{figure}
-
-Even the classifications on nodes and edges give the same results,
 
@@
 
-In this example we propose the iconographical \code{nodes$type} (character, weapon, etc.) nodes as the studied variable, but the user of the package can create and choose any other study variable: color for a painting, technique of realisation, size, etc.
-\code{edges$type} can also be extended to other types than normal, attribute, overlapping.
-The background is considered as homogeneous but a crack, a pit, a something can also be considered
-The plasticity of Graph Theory allows to develop conventions in order to quote the different types of relations s.
-Its geographical equivalent is a Thiessen polygon
+The \pkg{decorr} framework is grounded on a parsimonious model, where Tobler's \emph{near things} \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions.
+Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges).
+In this example we propose the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) nodes as the studied variable (\code{nd.var}), but the user of the package can create and choose any other study variable: \code{technique}, \code{color}, etc.
+The current edge notation with \emph{normal} edge, \emph{attribute} edge and \emph{overlapping} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
+Currently, the undecorated parts of the support -- the background -- is inactive and considered as homogeneous it is possible to consider but a crack on a rock or a carene on a pottery \citep{Huet18a}, like relevant topologic entities to understand the iconographical composition.
+Then a new class of nodes, beside the graphical units, can be registred to create a multipartite graph (\emph{n}-mode graph).
+The plasticity of graph theory will allow to filter these different classes of nodes.
+
+Beside giving a parsimonious framework to register, manage and analyse compositions, for further statistical analysis, the \pkg{decorr} package allows also to avoid, empirical and idiosynchratic top-down hierachical typologies (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious inventories and difficult possible comparisons \emph{within} and \emph{between} these categories.
+Indeed, to quote the GUs it is more parsimonious to use the edge notation rather than multiplying hierarchical categories.
+For example, with the two types of GUs \code{casque} (helmet) and \code{personnage} (anthropomorphe) and the \emph{attribute} edge it is possible to describe: (1) a case where an anthropomorph wears a helmet (like for the Cerro Muriano steale, Figure \ref{fig:dec.1.4} left), (2) a case where an anthropomorph is separated from a helmet (like for the Zarza de Montanchez, Figure \ref{fig:dec.1.4} right).
+With a top-down hierachical typology, three types of GUs would have been required: (1) anthropomorph wearing a helmet (Cerro Muriano); (2) anthropomorph and (3) helmet (Zarza de Montanchez).
+
+<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=1, fig.cap="\\label{fig:dec.1.4}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
+par(mar=c(0, 0, 0, 0))
+nds.df <- read_nds(site = "Cerro Muriano", 
+                   decor = "Cerro Muriano 1",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Cerro Muriano",
+                   decor = "Cerro Muriano 1",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+cm1 <- plot_dec_grph(nodes = nds.df,
+                     edges = eds.df,
+                     site = "Cerro Muriano",
+                     decor = "Cerro Muriano 1",
+                     nd.var = "type",
+                     lbl.size = 1.7,
+                     doss = system.file("extdata", 
+                                        package = "decorr"))
+cm1.img <- magick::image_read(cm1)
+nds.df <- read_nds(site = "Zarza de Montanchez", 
+                   decor = "Zarza De Montanchez",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Zarza de Montanchez",
+                   decor = "Zarza De Montanchez",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+zm <- plot_dec_grph(nodes = nds.df,
+                            edges = eds.df,
+                            site = "Zarza de Montanchez",
+                            decor = "Zarza De Montanchez",
+                            nd.var = "type",
+                            lbl.size = 1.7,
+                            doss = system.file("extdata", 
+                                               package = "decorr"))
+zm.img <- magick::image_read(zm)
+out.img <- magick::image_append(c(cm1.img, zm.img))
+plot(out.img)
+@
+
+Archaeologists use quite commonly the Harris matrix to model the relative chronology of archaeological layers (in a case of an excavation) or the relative chronology of iconographical layers (in case of a decorated support).
+The \emph{overlap} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.
+
+In Archaeology, the iconography has been considered as a privileged elements to the identification of cultural facies \citep{Renfrew91}.
+By integrating the recording, the management and the analysis process in a simple manner, within the framework of graph Theory and binary topological telationships (\emph{birel}), the \pkg{decorr} framework opens possibilities of studying graphical systems with normalized indexes over large series (Later Paleolithic cave paintings, \emph{cups-and-rings} art, Valcamonica Iron Age \emph{pittoti}, etc.) a long period of time, and on heterogeneous source data.
 
@@
 
-\begin{leftbar}
-All acknowledgments (note the AE spelling) should be collected in this
-unnumbered section before the references. It may contain the usual information
-about funding and feedback from colleagues/reviewers/etc. Furthermore,
-information such as relative contributions of the authors may be added here
-(if any).
-\end{leftbar}
+A part of the IT development of the \pkg{decorr} package has been done in the frame of a postdoctoral position (2015-16) supported by LabEx ARCHIMEDE from "Investissement d'Avenir" program ANR-11-LABX-0032-01.
 
@@
 
-%% -- Appendix (if any) --------------------------------------------------------
-%% - After the bibliography with page break.
-%% - With proper section titles and _not_ just "Appendix".
-
-\newpage
-
-\begin{appendix}
-
-\section{More technical details} \label{app:technical}
-
-\begin{leftbar}
-Appendices can be included after the bibliography (with a page break). Each
-section within the appendix should have a proper section title (rather than
-just \emph{Appendix}).
-
-For more technical style details, please check out JSS's style FAQ at
-\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
-which includes the following topics:
-\begin{itemize}
-\item Title vs.\ sentence case.
-\item Graphics formatting.
-\item Naming conventions.
-\item Turning JSS manuscripts into \proglang{R} package vignettes.
-\item Trouble shooting.
-\item Many other potentially helpful details\dots
-\end{itemize}
-\end{leftbar}
-
-\section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
-
-\begin{leftbar}
-References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
-references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
-\verb|\citealp| etc.\ (and never hard-coded). This commands yield different
-formats of author-year citations and allow to include additional details (e.g.,
-pages, chapters, \dots) in brackets. In case you are not familiar with these
-commands see the JSS style FAQ for details.
-
-Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
-when acquiring the entries automatically from mixed online sources. However,
-it is important that informations are complete and presented in a consistent
-style to avoid confusions. JSS requires the following format.
-\begin{itemize}
-\item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
-be used in the references.
-\item Titles should be in title case.
-\item Journal titles should not be abbreviated and in title case.
-\item DOIs should be included where available.
-\item Software should be properly cited as well. For \proglang{R} packages
-\code{citation("pkgname")} typically provides a good starting point.
-\end{itemize}
-\end{leftbar}
-
-\end{appendix}
+% 
+% %% -- Appendix (if any) --------------------------------------------------------
+% %% - After the bibliography with page break.
+% %% - With proper section titles and _not_ just "Appendix".
+% 
+% \newpage
+% 
+% \begin{appendix}
+% 
+% \section{More technical details} \label{app:technical}
+% 
+% \begin{leftbar}
+% Appendices can be included after the bibliography (with a page break). Each
+% section within the appendix should have a proper section title (rather than
+% just \emph{Appendix}).
+% 
+% For more technical style details, please check out JSS's style FAQ at
+% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
+% which includes the following topics:
+% \begin{itemize}
+% \item Title vs.\ sentence case.
+% \item Graphics formatting.
+% \item Naming conventions.
+% \item Turning JSS manuscripts into \proglang{R} package vignettes.
+% \item Trouble shooting.
+% \item Many other potentially helpful details\dots
+% \end{itemize}
+% \end{leftbar}
+% 
+% 
+% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}
+% 
+% \begin{leftbar}
+% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
+% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
+% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
+% formats of author-year citations and allow to include additional details (e.g.,
+% pages, chapters, \dots) in brackets. In case you are not familiar with these
+% commands see the JSS style FAQ for details.
+% 
+% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
+% when acquiring the entries automatically from mixed online sources. However,
+% it is important that informations are complete and presented in a consistent
+% style to avoid confusions. JSS requires the following format.
+% \begin{itemize}
+% \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
+% be used in the references.
+% \item Titles should be in title case.
+% \item Journal titles should not be abbreviated and in title case.
+% \item DOIs should be included where available.
+% \item Software should be properly cited as well. For \proglang{R} packages
+% \code{citation("pkgname")} typically provides a good starting point.
+% \end{itemize}
+% \end{leftbar}
+% 
+% \end{appendix}
 

### 2020-12-17 article_rvTH17.Rnw

@@
 # knitr::opts_chunk$set(fig.pos = '!htbp')
-loadfonts(device = "win")
+#loadfonts(device = "win")
+loadfonts(device = "pdf")
 @
@@
 \begin{document}
+%\SweaveOpts{concordance=TRUE}
 %\SweaveOpts{concordance=TRUE}
@@
 For example, the 3-letters word "\code{art}" can be modelled with three vertices (\code{a}, \code{r}, \code{t}) and two edges (one between \code{a} and \code{r}, the other between \code{r} and \code{t}).
-In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Fig.
-\ref{fig:fart}).
+In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Figure \ref{fig:fart}).
 
@@
 But, as stated, in Prehistorical the writing system does not exists.
-Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order (Fig.
-\ref{fig:fratar}).
+Spatial relationships between graphical features, or graphical units, are not necessarly linear and directed but could most probably be more multi-directional and undirected: the direction of the interactions of pairwise graphical units can be in any order (Figure \ref{fig:fratar}).
 
@@
 
-Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generaly not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
+Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generally not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
 \begin{itemize}
-\setlength\itemsep{.05em}
 \item unexplicit spatial groupings of graphical units -- like graphical units grouped into \emph{figures},  \emph{figures} grouped into \emph{patterns}, \emph{patterns} grouped into \emph{motives}, etc.
@@
 The entire graphical content of the decoration is mapped with a geometric graph -- or planar graph -- of GUs centroids (nodes) and GUs proximity links (edges).
-So, a decoration graph is a one component graph.
-This graph is the dual graph of the decoration Voronoi diagram (Fig.
-\ref{fig:gis}).
-
-\begin{figure}[H]
-% \begin{figure}[!htbp]
+Thus, a decoration graph is a one component graph.
+This graph is the dual graph of the decoration Voronoi diagram (Figure \ref{fig:gis}).
+
+\begin{figure}[!htbp]
 \centering
@@
 
-Whether this theoritical process include image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
-Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\
-
-\subsection[elements]{Graph elements} \label{sec:elements}
-
-\smallbreak
-\textbf{Nodes types}
-\\
-At least one variable, like the \code{type} of the nodes, is defined in order to analyse and compare iconographical compositions.
-In this article, we will only consider nodes \code{type} values ('anthropomorph', 'sword', etc.) but it the package user can add other variables (\code{technique}, \code{color}, etc.)
-
-\smallbreak
-\textbf{Egdes types}
-\\
-The package considers the most common types of relations between nodes:
-
+Whether this theoretical process includes image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
+
+\subsection{Node types} \label{sec:nodes}
+
+Nodes are uniquely identified by a number.
+But this does not identify or classify the nodes by the corresponding GU meaning or type.
+For this, name nodes could be assigned.
+However this presents two problems: nodes with repeated GU type in the same graph could not be handled and comparisons by different GU classifications would be hindered.
+For this reason, the R package decorr considers that the GU type is assigned in a node attribute or variable.
+Thus, at least one node variable must be defined in order to analyse and compare iconographical compositions.
+By default this node variable is named \code{type}.
+In this article examples, we consider node \code{type} values ('anthropomorph', 'sword', etc.), but the package user can add other variables (\code{technique}, \code{color}, etc.), which would produce complementary analyses.
+
+\subsection{Egdes types} \label{sec:egdes}
+
+The graph edges can represent different types of relationships between GUs.
+This relationship type is encoded in an edge variable named \code{type}.
+The current package version considers the most common types:
+%
 \begin{description}
-\item[\emph{normal edges}]
+\item[\emph{Normal edges.}]
 Two contemporaneous nodes having a Voronoi cell sharing a border have a common edge.
 By convention, this edge is tagged \code{'='} (\code{type = '='}), represented with an orange plain line and its textual notation is \code{'-=-'}.
-For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Fig.
-\ref{fig:gis_1}).
-As said, in Prehistorical iconography, the read direction is not known, so this edge is undirected, ie \code{personnage -=- bouclier} is equal to \code{bouclier -=- personnage} (Fig.
-\ref{fig:gis_1}).
-
-\item[\emph{attribute edges}]
-It occurs frequently that a node can be divided into a main unit (eg, an anthropomorph) and one or various attribute units (eg, a helmet, a male sex).
+For example: \code{personnage -=- bouclier} means that the nodes \code{personnage} and \code{bouclier} have a common border (Figure \ref{fig:gis_1}).
+As said, in Prehistorical iconography, the read direction is not known.
+Thus this edge type is undirected, ie \code{personnage -=- bouclier} is equal to \code{bouclier -=- personnage}.
+
+\item[\emph{Attribute edges.}]
+Frequently a node can be divided into a main unit (eg, an anthropomorph) and one or various attribute units (eg, a helmet, a male sex).
 To record this ordered set, the edge between a main and an attribute node is tagged \code{'+'} (\code{type = '+'}), represented with an orange dashed line and its textual notation is \code{'-+-'}.
 For example \code{personnage -+- casque} means that the main node \code{personnage} has the attribute node \code{casque}.
-The edge is directed since \code{personnage -+- casque} (the anthropomorph has an helmet) is not equal to \code{casque -+- personnage} (the helmet has an anthropomorph) (Fig.
-\ref{fig:gis_1}).
-
-% \begin{figure}[H]
+The edge is directed since \code{personnage -+- casque} (the anthropomorph has an helmet) is not equal to \code{casque -+- personnage} (the helmet has an anthropomorph) (Figure \ref{fig:gis_1}).
+
+\item[\emph{Overlap edges.}]
+Sometimes a graphical composition shows decoration phases.
+These decoration phases can be identified by observing the superimpositions between different nodes or by other means (differences in the techniques or in the patina for example).
+In the first case, two nodes belonging to two different phase will have a common part (the overlapped intersection) while in the second case the two nodes will not necessarly share a part of their surface.
+To record this relative chronology, an edge between two non-contemporaneous nodes is tagged \code{'>'} (\code{type = '>'}), represented with a blue plain line and its textual notation is \code{'->-'}.
+For example \code{ecriture ->- bouclier} means that \code{ecriture} is more recent than the \code{bouclier}, or \emph{overlaps} the\code{bouclier} in the stratigraphical sense.
+The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier}) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Figure \ref{fig:gis_2}).
+
+This stratigraphical information (node 1 \emph{over} node 2, or node 2 \emph{under} node 1) helps to understand the relative chronology between nodes and must be recorded.
+The analysis of the iconographical content can be stratified (stratified analysis) and performed on each different phases of decoration separately.
+
+\end{description}
+
 \begin{figure}[!htbp]
@@
 
-\item[\emph{overlapping edges}]
-Finally, it occurs that a graphical composition shows decoration phases.
-These decoration phases can be identified by observing the superimpositions between different nodes or by other means (differences in the techniques or in the patina for example).
-In the first case, two nodes belonging to two different phase will have a common part (the overlapped intersection) while in the second case the two nodes will not necessarly share a part of their surface.
-To record this relative chronology, an edge between two non-contemporaneous nodes is tagged \code{'>'} (\code{type = '>'}), represented with a blue plain line and its textual notation is \code{'->-'}.
-For example \code{ecriture ->- bouclier} means that \code{ecriture} is more recent than the \code{bouclier}, or \emph{overlaps} the\code{bouclier} in the stratigraphical sense.
-The edge is directed since \code{ecriture ->- bouclier} (ie, \code{ecriture} \emph{overlaps} \code{bouclier}) is not equal to \code{bouclier ->- ecriture} (ie, \code{bouclier} \emph{overlaps} \code{ecriture}) (Fig.
-\ref{fig:gis_2}).
-
-\end{description}
-
-\begin{figure}[H]
-% \begin{figure}[!htbp]
+\begin{figure}[!htbp]
 \centering
@@
 
-This stratigraphical information (node 1 \emph{over} node 2, or node 2 \emph{under} node 1) helps to understand the relative chronology between node and must be recorded.
-The analysis of the iconographical content can be stratified (stratified analysis) and performed on each different phases of decoration separately.
-
-The nodes and edges equalities are resumed in the Table \ref{tab:ex.comm}.
-
-<<equalities, results='asis', echo=FALSE, fig.pos = '!htbp'>>=
-caption <- paste0("Example of equalities and differences for nodes and edges ")
-df.ex <- data.frame(type = c("node", "node", "edge", 
-                             "edge", "edge", "edge", "edge"),
-                    part.1 = c("epee", "epee", "epee --=-- lance",
-                               "epee --=-- lance", "personnage --=-- bouclier", 
-                               "personnage --+-- bouclier", "ecriture -->-- bouclier"),
-                    part.2 = c("epee", "lance", "epee --=-- lance",
-                               "lance --=-- epee", "personnage --+-- bouclier",
-                               "bouclier --+-- personnage", "bouclier -->-- ecriture"),
-                    is.equal = c("True", "False", "True", 
-                                 "False", "False", "False", "False"),
-                    stringsAsFactors = F)
-print(xtable::xtable(df.ex,
-                     caption = caption,
-                     label = "tab:ex.comm",
-                     size = 7),
-      include.rownames = FALSE,
-      table.placement = "!htbp")
-@
-
-These different relationship types are resumed with their \emph{birel} equivalencies and chronological assumptions in Table \ref{table:tedges}.
-
-% \begin{table}[!htbp]
-\begin{table}[H]
+These three relationship types are resumed with their \emph{birel} equivalencies and chronological assumptions in Table \ref{table:tedges}.
+The equality conditions between nodes and between edges are resumed in Table \ref{tab:ex.comm}, taking into account the directness of each edge type.
+
+\begin{table}[!htbp]
 \centering
-\begin{tabular}{ccccp{2cm}p{4.9cm}}
+\begin{tabular}{cccccp{4.4cm}}
 \hline
@@
 \code{A} & \code{=} & \code{B} & Undirected & $ A \cap B = \emptyset $ & 
-\code{A} and \code{B} are contemporaneous and disjoint. \\
+\code{A} and \code{B} contemporaneous \newline and disjoint. \\
 \code{A} & \code{+} & \code{B} & Directed & $ A \cap B = A $ & 
-\code{A} and \code{B} are contemporaneous. \code{B} is an attribute of \code{A}. \\
-\code{A} & \code{>} & \code{B} & Directed & $ A \cap B = \exists $ \newline or \newline $ A \cap B = \emptyset $ & 
-\code{A} is more recent than \code{B}. \\
+\code{A} and \code{B} contemporaneous. \code{B} is an attribute of \code{A}. \\
+\code{A} & \code{>} & \code{B} & Directed & Any $A \cap B$ & 
+\code{A} more recent than \code{B}. \newline Overlapping or disjoint. \\
 % \code{1} & = & 2 & undirected & A \cap B = \varnothing & 1 and 2 can be contemporaneous \\
@@
 
+<<equalities, results='asis', echo=FALSE, fig.pos = '!htbp'>>=
+caption <- paste0("Example of equalities and differences for nodes and edges ")
+df.ex <- as.data.frame(t(array(c(
+  "node", "epee", "epee", "True", 
+  "node", "epee", "lance", "False", 
+  "edge", "epee --=-- lance", "epee --=-- lance", "True", 
+  "edge", "epee --=-- lance", "lance --=-- epee", "True", 
+  "edge", "personnage --+-- bouclier", "personnage --+-- bouclier", "True",
+  "edge", "personnage --+-- bouclier", "bouclier --+-- personnage", "False",
+  "edge", "ecriture -->-- bouclier", "ecriture -->-- bouclier", "True",
+  "edge", "ecriture -->-- bouclier", "bouclier -->-- ecriture", "False",
+  "edge", "epee --=-- lance", "epee --=-- bouclier", "False", 
+  "edge", "personnage --=-- bouclier", "personnage --+-- bouclier", "False"),
+      dim = c(4,10))))
+names(df.ex) <- c("Element class", "Element 1", "Element 2", "is.equal?")
+print(xtable::xtable(df.ex,
+                     caption = caption,
+                     label = "tab:ex.comm",
+                     size = 7,
+                     align = c("c", "c", "c", "c", "c")),
+      include.rownames = FALSE,
+      table.placement = "!htbp")
+@
+
 \section{The R package decorr} \label{sec:models}
@@
 
-\subsection{External package} \label{sec:ext_pck}
+\subsection{External package dependencies} \label{sec:ext_pck}
 
@@
 
-A training dataset is stored in the \code{extdata} folder of the \pkg{decorr} package.
-It contains dataframes of nodes' coordinates, edges' coordinates, decorations' listing in different formats (\code{.tsv}, \code{.csv} or \emph{shapefiles}) and the decoration drawings (\code{.jpg}) of five Late Bronze Age stelae belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- mostly concentrated in the Southwest Iberian peninsula \citep{Celestino01}.
-\\
-
-\smallbreak
-\textbf{Imgs dataframe}
-\\
-The inventory of decorations is stored in the \code{imgs} dataframe (Tab.
-\ref{tab:timg}).
+A small testing dataset is provided with the \pkg{decorr} package, stored in its \code{extdata} folder.
+It contains dataframes of nodes' coordinates, edges' coordinates, decorations' listing in different formats (\code{.tsv}, \code{.csv} or \emph{shapefiles}), and the decoration drawings (\code{.jpg}) of five Late Bronze Age stelae belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- mostly concentrated in the Southwest Iberian peninsula \citep{Celestino01}.
+
+\subsubsection{Imgs dataframe}
+The inventory of decorations is stored in the comma-separated values file \code{imgs.csv} and the tab-separated values files \code{imgs.tsv}.
+Any of them can be read as a dataframe (Table \ref{tab:timg}),
+%
+<<results='asis', echo=TRUE, warning=FALSE>>=
+imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
+                   sep="\t", stringsAsFactors = FALSE)
+@
 The field \code{imgs$idf} is the short name of the decoration, useful during statistical analysis.
 The primary key of each decoration is the concatenate of \code{imgs$site} and \code{imgs$decor}.
-These keys will allow joints with the other dataframes (nodes and edges)
-%
+These keys will allow joints with the other dataframes (nodes and edges).
 <<results='asis', echo=FALSE, warning=FALSE>>=
-imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-                   sep="\t", stringsAsFactors = FALSE)
 print(xtable::xtable(imgs,
@@
 @
-%
-
-At first the drawings dataset \citep{DiazGuardamino10} can be checked by using the imgs dataframe and the \pkg{magick} package (Fig. \ref{fig:dataset})
-%
-<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.height=11.2, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
+
+At first the drawings dataset \citep{DiazGuardamino10} can be checked by using the information in the \code{imgs} dataframe and the \pkg{magick} package (Figure \ref{fig:dataset})
+%
+<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.7507508, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
 pth <- system.file("extdata", package = "decorr")
-imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
-                   sep="\t", stringsAsFactors = FALSE)
 lims <- list()
 for(i in 1:nrow(imgs)){
-  i1 <- image_read(paste0(pth,"\\",imgs[i,"img"]))
-  lbl.txt <- paste0(imgs[i,"idf"], "\n",
-                    imgs[i,"site"], "\n",
-                    imgs[i,"decor"], "\n",
-                    imgs[i,"img"], "\n",
+  i1 <- image_read(paste0(pth,"/",imgs$img[i]))
+  lbl.txt <- paste0(paste(imgs[i,], collapse = "\n"), "\n",
                     image_info(i1)$width, "*", image_info(i1)$height, " px")
-  i1 <- image_annotate(i1, lbl.txt, location = "northwest",
+  lims[[i]] <- image_annotate(i1, lbl.txt, location = "northwest",
                               size = 28, color = "red")
-  lims[[length(lims)+1]]<- i1
 }
@@
 %
-To construct a graph overlapping the decoration images listed in the \code{images} dataframe, the first step is to load nodes and edges dataframes.
+To construct a graph overlapping the decoration images listed in the \code{imgs} dataframe, the first step is to load nodes and edges dataframes.
 % \begin{itemize}
@@
 % \end{itemize}
-\smallbreak
-\textbf{Nodes dataframe}
-\\
-It contains the required minimum variables for the analysis (Tab. \ref{tab:tnds}).
-
-<<tndss, results='asis', echo=TRUE>>=
-caption <- "Nodes (from \\code{nodes.csv} dataframe)"
+
+\subsubsection{Nodes dataframe}
+The nodes for the five decorations are stored in different formats in \code{nodes.csv}, \code{nodes.tsv}, and \code{nodes.shp}. This can be read by
+
+<<results='asis', echo=TRUE>>=
 nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
                     sep=";", stringsAsFactors = FALSE)
-print(xtable::xtable(head(nodes),
+@
+
+generating a dataframe containing the required minimum variables for the analysis (Table \ref{tab:tnds}).
+
+<<tndss, results='asis', echo=FALSE>>=
+caption <- "Nodes (from \\code{nodes.csv} dataframe)"
+print(xtable::xtable(nodes,
                      caption = caption,
@@
 To recover the local \emph{y} value of nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
-\smallbreak
-% \begin{itemize}
-% \item The \code{edges} dataframe
-% \end{itemize}
-\textbf{Edges dataframe}
-\\
-The edges dataframe is quite similar to the nodes dataframe (Table \ref{tab:teds}).
-
-<<tedss, results='asis', echo=TRUE>>=
+
+\subsubsection{Edges dataframe}
+Analogously, the edges for the five decorations are stored in different formats in \code{edges.csv}, \code{edges.tsv}, and \code{edges.shp}. This can be read by 
+<<results='asis', echo=TRUE>>=
+edges <- read.table(system.file("extdata", "edges.csv", 
+                                package = "decorr"),
+                    sep=";",stringsAsFactors = FALSE)
+@
+
+generating the corresponding edges dataframe (Table \ref{tab:teds}).
+
+<<tedss, results='asis', echo=FALSE>>=
 caption <- "Edges of Cerro Muriano 1 (from \\code{edges.csv} dataframe)"
-edges <- read.table(system.file("extdata", "edges.csv", package = "decorr"),
-                    sep=";",stringsAsFactors = FALSE)
-edges.cm1 <- subset(edges, decor == "Cerro Muriano 1")
-print(xtable::xtable(head(edges.cm1),
+print(xtable::xtable(edges,
+                     align = "rllrrc",
                      caption=caption,
@@
       include.rownames = FALSE,
+      scalebox = 0.95,
+      size = "\\renewcommand{\\arraystretch}{0.92}",
+#      size = "\\small\\renewcommand{\\arraystretch}{0.989}",
       table.placement="!htbp")
@@
 
-Fields \code{edges$site} and \code{edges$decor} are the primary key of decoration.
+The fields \code{edges$site} and \code{edges$decor} are the primary key of decoration.
 The fields \code{edges$a} and \code{edges$b} are the equivalent to columns \emph{from} and \emph{to} in Graph theory.
@@
 The column \code{edges$b} is the identifier of the \emph{ending node} or \emph{attribue node} or \emph{overlapped node}.
-The column \code{edges$type} is the type of relation (normal, attribute, overlapping, etc.) between the node a and the node b.
-There is no need to record the coordinates of the edges, these coordinates are calculated from the nodes dataframe.
-
+The column \code{edges$type} is the type of relation (normal, attribute, overlap, etc.) between the node a and the node b.
+There is no need to record the coordinates of the edges, these coordinates are extracted from the nodes dataframe.
+
+\bigskip
 The \pkg{decorr} package functions can be divided into:
@@
 The result is a list of \code{igraph} objects.
-As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function (Fig.
-\ref{fig:igraph})
+As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function (Figure \ref{fig:igraph})
 
@@
 
-<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:ibah}Ibahernando decoration, node 1 overlaps node 2 and node 3">>=
+<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.5, fig.asp=1.126126, fig.cap="\\label{fig:ibah}Ibahernando decoration, node 1 overlaps node 2 and node 3">>=
 par(mar=c(0,0,0,0))
@@
 
-<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=1, fig.cap="\\label{fig:ibah2}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
+<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=0.5630631, fig.cap="\\label{fig:ibah2}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
 par(mar=c(0,0,0,0))
@@
 
-The \code{named\_elements()} function allows to calculate the textual expression of graph elements (nodes and edges) with a disambiguation of these elements -- adding '#' suffixes -- when several nodes or edges have the same values.
+The \code{named\_elements()} function allows to calculate the textual expression of graph elements (nodes and edges) with a disambiguation of these elements -- adding '\#' suffixes -- when several nodes or edges have the same values.
 
@@
 
-<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.asp=1, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
+<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.5, fig.asp=1.126126, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
 par(mar=c(0, 0, 0, 0))
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.height=4.2, fig.cap="\\label{fig:ceds}Comparisons between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
+<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.asp=0.5630631, fig.cap="\\label{fig:ceds}Comparison between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
 par(mar=c(0,0,0,0))
@@
 
-The edges comparisons between the decoration 1 and the decoration 4 show that they have two (2) common edges, see also the Figure \ref{fig:ceds}.
+The edges comparisons between the decoration 1 and the decoration 4 show that they have two (2) common edges, see also Figure \ref{fig:ceds}.
 
@@
 df.same_edges <- same_elements(lgrph, "type", "edges")
-df.same_nodes<- same_elements(lgrph, "type", "nodes")
+df.same_nodes <- same_elements(lgrph, "type", "nodes")
 @
@@
                      size = 8,
-                     digits = c(0)),
+                     digits = 0),
       floating=FALSE,
@@
 \centering
-\begin{minipage}{.4\textwidth}
+\begin{minipage}{.44\textwidth}
 \centering
@@
 library("matlib")
-dist.nodes  <- dist(as.matrix(df.same_nodes))
+dist.nodes  <- dist(as.matrix(df.same_nodes), upper=TRUE)
 mat.nodes <- as.matrix(dist.nodes)
@@
                      size = 8,
-                     digits = c(2)),
+                     digits = 2),
       floating=FALSE,
@@
 \end{minipage}
-\begin{minipage}{.4\textwidth}
+\begin{minipage}{.44\textwidth}
 \centering
@@
 
-<<hclust, echo=TRUE, warning=FALSE, fig.height=5, fig.pos = '!htbp', fig.cap="\\label{fig:hclust}Hierachical clusterings on decoration common elements">>=
-par(mfrow=c(1, 2))
+<<hclust, echo=TRUE, warning=FALSE, fig.height=4, fig.width=8, fig.pos = '!htbp', fig.cap="\\label{fig:hclust}Hierachical clusterings on decoration common elements">>=
+par(mfrow=c(1, 2), mar=c(0.2,4,1.2,0.5))
 hc1 <- hclust(dist.nodes, method = "ward.D")
 hc2 <- hclust(dist.edges, method = "ward.D") 
-plot(hc1, main = "common nodes")
-plot(hc2, main = "common edges")
+plot(hc1, main = "Common nodes", cex.main = 1.1)
+plot(hc2, main = "Common edges", cex.main = 1.1)
 @
@@
   untangle(method = "step1side") %>% 
-  tanglegram()                       
+  tanglegram(main_left="Common nodes", main_right="Common edges", cex.main=1.5,
+             margin_top = 2)                       
 @
@@
 In this example we propose the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) nodes as the studied variable (\code{nd.var}), but the user of the package can create and choose any other study variable: \code{technique}, \code{color}, etc.
-The current edge notation with \emph{normal} edge, \emph{attribute} edge and \emph{overlapping} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
+The current edge notation with \emph{normal} edge, \emph{attribute} edge, and \emph{overlap} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
 Currently, the undecorated parts of the support -- the background -- is inactive and considered as homogeneous it is possible to consider but a crack on a rock or a carene on a pottery \citep{Huet18a}, like relevant topologic entities to understand the iconographical composition.
@@
 
-<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=1, fig.cap="\\label{fig:dec.1.4}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
+<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.5630631, fig.cap="\\label{fig:dec.1.4}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
 par(mar=c(0, 0, 0, 0))

### 2020-12-17 article_rvTH17.Rnw
article

@@
 
-Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generally not quantified, applying the Saussurian model to any prehistorical graphical content had led to considerable problems:
+Because of the inherent variability of iconography, and because graphical and spatial proximities between graphical units are generally not quantified, applying the Saussurian model to any structured prehistorical graphical content had led to considerable problems:
 \begin{itemize}
@@
 Whether this theoretical process includes image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
-Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\
+
+The notation of with nodes allows also to avoid, empirical and  top-down hierachical typologies .
+According to us, it is more convenient to record GUs (eg, an anthropomorph) and GUs varietes (an anthropomorph wearing an helmet) with different spatialized nodes (firstly within a GIS) rather than multiplying hierarchical levels of categories (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious idiosynchratic inventories and difficult possible comparisons \emph{within} and \emph{between} these categories.
+For example, with the two types of GUs \code{casque} (helmet) and \code{personnage} (anthropomorphe) and the \emph{attribute} edge it is possible to describe: (1) a case where an anthropomorph wears a helmet (like for the Cerro Muriano steale, Figure \ref{fig:dec.1.4} left), (2) a case where an anthropomorph is separated from a helmet (like for the Zarza de Montanchez, Figure \ref{fig:dec.1.4} right).
+With a top-down hierachical typology, three types of GUs would have been required: (1) anthropomorph wearing a helmet (Cerro Muriano); (2) anthropomorph and (3) helmet (Zarza de Montanchez).
+
+<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.5630631, fig.cap="\\label{fig:dec.1.4} Cerro Muriano 1 and Zarza de Montanchez decorations.">>=
+library(decorr)
+par(mar=c(0, 0, 0, 0))
+nds.df <- read_nds(site = "Cerro Muriano", 
+                   decor = "Cerro Muriano 1",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Cerro Muriano",
+                   decor = "Cerro Muriano 1",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+cm1 <- plot_dec_grph(nodes = nds.df,
+                     edges = eds.df,
+                     site = "Cerro Muriano",
+                     decor = "Cerro Muriano 1",
+                     nd.var = "type",
+                     lbl.size = 1.7,
+                     doss = system.file("extdata", 
+                                        package = "decorr"))
+cm1.img <- magick::image_read(cm1)
+nds.df <- read_nds(site = "Zarza de Montanchez", 
+                   decor = "Zarza De Montanchez",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+eds.df <- read_eds(site = "Zarza de Montanchez",
+                   decor = "Zarza De Montanchez",
+                   doss = system.file("extdata", 
+                                      package = "decorr"))
+zm <- plot_dec_grph(nodes = nds.df,
+                            edges = eds.df,
+                            site = "Zarza de Montanchez",
+                            decor = "Zarza De Montanchez",
+                            nd.var = "type",
+                            lbl.size = 1.7,
+                            doss = system.file("extdata", 
+                                               package = "decorr"))
+zm.img <- magick::image_read(zm)
+out.img <- magick::image_append(c(cm1.img, zm.img))
+plot(out.img)
+@
 
@@
 The \code{nodes$x} and \code{nodes$y} columns refer to the \emph{x} and \emph{}y coordinates of the nodes.
-As said, in the first place theses coordinates come from the GIS.
-But, in a GIS, the coordinates origin (0, 0) is the bottom-left corner and exist negative values, while this origin is top-left for any \proglang{R} matrices (rasters, grids, dataframes, etc.) with only positive values.
+As said, in the first place theses coordinates come from the GIS where the origin (0, 0) is the bottom-left corner, while for any \proglang{R} matrices (rasters, grids, dataframes, etc.) this origin is top-left (see Figure \ref{fig:gis}.
+A).
 To recover the local \emph{y} value of nodes and edges on the decoration image, the absolute value \emph{y} value and the image height, as a constant offset, are computed.
@@
 caption <- "Edges of Cerro Muriano 1 (from \\code{edges.csv} dataframe)"
-print(xtable::xtable(edges,
+print(xtable::xtable(edges[edges$decor == "Cerro Muriano 1", ],
                      align = "rllrrc",
@@
 The result is a list of \code{igraph} objects.
-As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function (Figure \ref{fig:igraph})
-
-<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=2.5, fig.height=2.5, fig.cap="\\label{fig:igraph}R interface.
-Plot of the first graph of the list (Cerro Muriano 1 decoration)">>=
+As an example, the first graph of the list can be plotted with the \code{igraph::plot()} function.
+By default, the graphs spazialisation (\code{layout}) is based on \code{x} and \code{y} columns, when exist.
+This is our case since we work with geometric graphs.
+For a same graph, very different layouts (\emph{réalisations}) exist \citep{Mathis03}(\emp{graph drawing}) (Figure \ref{fig:igraph})
+
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=6, fig.height=4, fig.cap="\\label{fig:igraph}R interface.
+Plot of the first graph of the list (Cerro Muriano 1 decoration) with graph drawing on x, y coordinates (left), and force-directed graph drawing with nodes sized on their degree centralities (right)">>=
 lgrph <- list_dec(imgs, nodes, edges)
-par(mar=c(0, 0, 0, 0))
-plot(lgrph[[1]],
-     vertex.color = "orange",
-     vertex.frame.color = "orange",
-     vertex.label.color = "black",
+g <- lgrph[[1]]
+par(mar = c(0, 0, 0, 0), mfrow = c(1, 2))
+coords <- layout.fruchterman.reingold(g)
+plot(g,
      vertex.size = 14,
-     vertex.label.cex = .8,
+     vertex.label.family = "Helvetica",
      edge.color = "orange",
-     vertex.label.family = "Helvetica"
+     vertex.label.cex = .8
+)
+plot(g,
+     layout = coords,
+     vertex.size = 4 + (degree(g)*10),
+     vertex.label.family = "Helvetica",
+     edge.color = "orange",
+     vertex.label.cex = .8
 )
@@
 R> lgrph <- list_dec(imgs, nodes, edges)
-R> as.character(named_elements(lgrph[[2]], focus = "edges", nd.var="type"))
+R> named_elements(lgrph[[2]], focus = "edges", nd.var="type")
 ## [1] "epee-=-fibule" "bouclier-=-miroir"
@@
 par(mfrow=c(1, 2), mar=c(0.2,4,1.2,0.5))
-hc1 <- hclust(dist.nodes, method = "ward.D")
-hc2 <- hclust(dist.edges, method = "ward.D") 
-plot(hc1, main = "Common nodes", cex.main = 1.1)
-plot(hc2, main = "Common edges", cex.main = 1.1)
+hc.nds <- hclust(dist.nodes, method = "ward.D")
+hc.eds <- hclust(dist.edges, method = "ward.D") 
+plot(hc.nds, main = "common nodes", cex = 1.1)
+plot(hc.eds, main = "common edges", cex = 1.1)
 @
@@
 suppressPackageStartupMessages(library(dplyr))
-dend1 <- as.dendrogram (hc1)
-dend2 <- as.dendrogram (hc2)
-dendlist(dend1, dend2) %>%
+par(mfrow=c(1, 2))
+dend.nds <- as.dendrogram (hc.nds)
+dend.eds <- as.dendrogram (hc.eds)
+dendlist(dend.nds, dend.eds) %>%
   untangle(method = "step1side") %>% 
-  tanglegram(main_left="Common nodes", main_right="Common edges", cex.main=1.5,
-             margin_top = 2)                       
+  tanglegram(columns_width = c(6, 1, 6),
+             main_left = "common nodes",
+             main_right = "common edges",
+             lab.cex = 1.5,
+             cex_main = 1.5,
+             highlight_branches_lwd = F)                    
 @
@@
 \section{Summary and discussion} \label{sec:summary}
-
-The \pkg{decorr} framework is grounded on a parsimonious model, where Tobler's \emph{near things} \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions.
-Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges).
-In this example we propose the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) nodes as the studied variable (\code{nd.var}), but the user of the package can create and choose any other study variable: \code{technique}, \code{color}, etc.
-The current edge notation with \emph{normal} edge, \emph{attribute} edge, and \emph{overlap} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
-Currently, the undecorated parts of the support -- the background -- is inactive and considered as homogeneous it is possible to consider but a crack on a rock or a carene on a pottery \citep{Huet18a}, like relevant topologic entities to understand the iconographical composition.
-Then a new class of nodes, beside the graphical units, can be registred to create a multipartite graph (\emph{n}-mode graph).
-The plasticity of graph theory will allow to filter these different classes of nodes.
-
-Beside giving a parsimonious framework to register, manage and analyse compositions, for further statistical analysis, the \pkg{decorr} package allows also to avoid, empirical and idiosynchratic top-down hierachical typologies (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious inventories and difficult possible comparisons \emph{within} and \emph{between} these categories.
-Indeed, to quote the GUs it is more parsimonious to use the edge notation rather than multiplying hierarchical categories.
-For example, with the two types of GUs \code{casque} (helmet) and \code{personnage} (anthropomorphe) and the \emph{attribute} edge it is possible to describe: (1) a case where an anthropomorph wears a helmet (like for the Cerro Muriano steale, Figure \ref{fig:dec.1.4} left), (2) a case where an anthropomorph is separated from a helmet (like for the Zarza de Montanchez, Figure \ref{fig:dec.1.4} right).
-With a top-down hierachical typology, three types of GUs would have been required: (1) anthropomorph wearing a helmet (Cerro Muriano); (2) anthropomorph and (3) helmet (Zarza de Montanchez).
-
-<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.5630631, fig.cap="\\label{fig:dec.1.4}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
-par(mar=c(0, 0, 0, 0))
-nds.df <- read_nds(site = "Cerro Muriano", 
-                   decor = "Cerro Muriano 1",
-                   doss = system.file("extdata", 
-                                      package = "decorr"))
-eds.df <- read_eds(site = "Cerro Muriano",
-                   decor = "Cerro Muriano 1",
-                   doss = system.file("extdata", 
-                                      package = "decorr"))
-cm1 <- plot_dec_grph(nodes = nds.df,
-                     edges = eds.df,
-                     site = "Cerro Muriano",
-                     decor = "Cerro Muriano 1",
-                     nd.var = "type",
-                     lbl.size = 1.7,
-                     doss = system.file("extdata", 
-                                        package = "decorr"))
-cm1.img <- magick::image_read(cm1)
-nds.df <- read_nds(site = "Zarza de Montanchez", 
-                   decor = "Zarza De Montanchez",
-                   doss = system.file("extdata", 
-                                      package = "decorr"))
-eds.df <- read_eds(site = "Zarza de Montanchez",
-                   decor = "Zarza De Montanchez",
-                   doss = system.file("extdata", 
-                                      package = "decorr"))
-zm <- plot_dec_grph(nodes = nds.df,
-                            edges = eds.df,
-                            site = "Zarza de Montanchez",
-                            decor = "Zarza De Montanchez",
-                            nd.var = "type",
-                            lbl.size = 1.7,
-                            doss = system.file("extdata", 
-                                               package = "decorr"))
-zm.img <- magick::image_read(zm)
-out.img <- magick::image_append(c(cm1.img, zm.img))
-plot(out.img)
-@
-
-Archaeologists use quite commonly the Harris matrix to model the relative chronology of archaeological layers (in a case of an excavation) or the relative chronology of iconographical layers (in case of a decorated support).
-The \emph{overlap} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.
 
@@
 By integrating the recording, the management and the analysis process in a simple manner, within the framework of graph Theory and binary topological telationships (\emph{birel}), the \pkg{decorr} framework opens possibilities of studying graphical systems with normalized indexes over large series (Later Paleolithic cave paintings, \emph{cups-and-rings} art, Valcamonica Iron Age \emph{pittoti}, etc.) a long period of time, and on heterogeneous source data.
+\\
+The \pkg{decorr} framework is grounded on a parsimonious model, where Tobler's '\emph{near things}' \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions.
+Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges).
+The current edge notation with \emph{normal} edge, \emph{attribute} edge, and \emph{overlap} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
+For the nodes, we have only considered the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) as the studied variable (\code{nd.var}), but the package user can create and select any other structured vocabularies: \code{technique}, \code{color}, etc.
+Tree structures for hierarchical datasets allow generalization processes (up, the parent level) and specification processes (down, the children level).
+Such a formalism can be used to weight the differences between nodes, to conduct analysis with different level of precision or to overcome idiosynchratic typologies issues.
+To record the relative chronology of archaeological layers (in a case of an excavation) or the relative chronology of iconographical layers (in case of a decorated support), archaeologists use quite commonly a hierachical dataset to model : the Harris matrix.
+The \emph{overlap} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.
+The Harris matrix can be then traced with R data tree package (\pkg{data.tree}, \cite{Glur19}) and integrated into an interactive Reingold-Tilford tree diagram  (\pkg{collapsibleTree}, \cite{Khan18}).
+
+Here, the undecorated parts of the support -- the background -- is inactive and considered as homogeneous it is possible to consider but a crack on a rock or a carene on a pottery \citep{Huet18a}, like relevant topologic entities to understand the iconographical composition.
+Then a new class of nodes (eg, topographical units), beside the graphical units, can be registred to create a \emph{n}-mode graph (graph approach), or adding a new raster (eg, a Digital Elevation Model) within a decoration GIS project file (geostatistical approach).
 

### 2020-12-17 article_rvTH17.Rnw
overlap -> diachronic

@@
 
-\item[\emph{Overlap edges.}]
+\item[\emph{Diachronic edges.}]
 Sometimes a graphical composition shows decoration phases.
@@
 The column \code{edges$b} is the identifier of the \emph{ending node} or \emph{attribue node} or \emph{overlapped node}.
-The column \code{edges$type} is the type of relation (normal, attribute, overlap, etc.) between the node a and the node b.
+The column \code{edges$type} is the type of relation (normal, attribute, diachronic, etc.) between the node a and the node b.
 There is no need to record the coordinates of the edges, these coordinates are extracted from the nodes dataframe.
@@
 A latin inscription (\code{ecriture}, node 1) have been carved over a spear (\code{lance}, node 2) and overlaps partially a V-notched shield (\code{bouclier}, node 3).
-The edges between node 1 and node 2, and the edge between node 1 and node 3, are \emph{overlap} edges (Figure \ref{fig:ibah}).
+The edges between node 1 and node 2, and the edge between node 1 and node 3, are \emph{diachronic} edges (Figure \ref{fig:ibah}).
 
@@
 Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges).
-The current edge notation with \emph{normal} edge, \emph{attribute} edge, and \emph{overlap} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
+The current edge notation with \emph{normal} edge, \emph{attribute} edge, and \emph{diachronic} edge allow to describe most cases of Prehistorical iconography, but can also be extended to accept new types like, for example, the \emph{interact} edge when two graphical units interact (ie, when an action is visible).
 For the nodes, we have only considered the iconographical \code{type} (\code{personnage}, \code{casque}, etc.) as the studied variable (\code{nd.var}), but the package user can create and select any other structured vocabularies: \code{technique}, \code{color}, etc.
@@
 To record the relative chronology of archaeological layers (in a case of an excavation) or the relative chronology of iconographical layers (in case of a decorated support), archaeologists use quite commonly a hierachical dataset to model : the Harris matrix.
-The \emph{overlap} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.
+The \emph{diachronic} edge, by recording the anterior-posterior relation between pairwises of graphical units, is first operator permiting to trace the Harris matrix directly of the studied decoration.
 The Harris matrix can be then traced with R data tree package (\pkg{data.tree}, \cite{Glur19}) and integrated into an interactive Reingold-Tilford tree diagram  (\pkg{collapsibleTree}, \cite{Khan18}).

### 2020-12-18 article_rvTH17.Rnw
article

@@
 %% - \Shorttitle{} with LaTeX markup (if any), used as running title
-\title{Modelling Prehistoric Iconographic Compositions. The \proglang{R} package \pkg{decorr}}
-\Plaintitle{Modelling Prehistoric Iconographic Compositions. The R package 'decorr'}
+\title{Modelling Prehistoric Iconographic Compositions. The \proglang{R} package \pkg{iconr}}
+\Plaintitle{Modelling Prehistoric Iconographic Compositions. The R package 'iconr'}
 \Shorttitle{Modelling Prehistoric Iconographic Compositions}
@@
 At the composition level, recognition of meaningful associations of signs and reccurent patterns indicate clearly the existence of social conventions in the way to display and to read these expressions.
-We present the \pkg{decorr} \proglang{R} package which grounds concepts, methods and tools to analyse any ancient graphical systems.
+We present the \pkg{iconr} \proglang{R} package which grounds concepts, methods and tools to analyse any ancient graphical systems.
 Our asumption is that i) any graphical system is a spatial distribution of features, and ii) these features have possibly any meaningful relationships one with another depending on their pairwise spatial proximities.
@@
 
-In this article we present the \proglang{R} package \pkg{decorr}.
+In this article we present the \proglang{R} package \pkg{iconr}.
 Its purpose is to formalise a method and offer tools to analyse any graphical content (here: decorations) with methods coming from raph Theory, GIS and binary topological relationships (\emph{birel}) modeling.
@@
 <<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.5630631, fig.cap="\\label{fig:dec.1.4} Cerro Muriano 1 and Zarza de Montanchez decorations.">>=
-library(decorr)
+library(iconr)
 par(mar=c(0, 0, 0, 0))
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 eds.df <- read_eds(site = "Cerro Muriano",
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 cm1 <- plot_dec_grph(nodes = nds.df,
@@
                      doss = system.file("extdata", 
-                                        package = "decorr"))
+                                        package = "iconr"))
 cm1.img <- magick::image_read(cm1)
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 eds.df <- read_eds(site = "Zarza de Montanchez",
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 zm <- plot_dec_grph(nodes = nds.df,
@@
                             doss = system.file("extdata", 
-                                               package = "decorr"))
+                                               package = "iconr"))
 zm.img <- magick::image_read(zm)
@@
 However this presents two problems: nodes with repeated GU type in the same graph could not be handled and comparisons by different GU classifications would be hindered.
-For this reason, the R package decorr considers that the GU type is assigned in a node attribute or variable.
+For this reason, the R package iconr considers that the GU type is assigned in a node attribute or variable.
 Thus, at least one node variable must be defined in order to analyse and compare iconographical compositions.
@@
 
-\section{The R package decorr} \label{sec:models}
+\section{The R package iconr} \label{sec:models}
 
 <<results='asis', include=FALSE, echo=FALSE, warning=FALSE>>=
-library("decorr")
+library("iconr")
 library("igraph")
@@
 %
-The \pkg{decorr} package can be downloaded from GitHub
+The \pkg{iconr} package can be downloaded from GitHub
 %
@@
 
-The \pkg{decorr} package imports the following packages:
+The \pkg{iconr} package imports the following packages:
 \begin{itemize}
@@
 
-A small testing dataset is provided with the \pkg{decorr} package, stored in its \code{extdata} folder.
+A small testing dataset is provided with the \pkg{iconr} package, stored in its \code{extdata} folder.
 It contains dataframes of nodes' coordinates, edges' coordinates, decorations' listing in different formats (\code{.tsv}, \code{.csv} or \emph{shapefiles}), and the decoration drawings (\code{.jpg}) of five Late Bronze Age stelae belonging to the so-called 'Warrior stelae' family -- with about 140 stelae -- mostly concentrated in the Southwest Iberian peninsula \citep{Celestino01}.
@@
 <<results='asis', echo=TRUE, warning=FALSE>>=
-imgs <- read.table(system.file("extdata", "imgs.tsv", package = "decorr"),
+imgs <- read.table(system.file("extdata", "imgs.tsv", package = "iconr"),
                    sep="\t", stringsAsFactors = FALSE)
@@
 <<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.7507508, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
-pth <- system.file("extdata", package = "decorr")
+pth <- system.file("extdata", package = "iconr")
 lims <- list()
@@
 <<results='asis', echo=TRUE>>=
-nodes <- read.table(system.file("extdata", "nodes.csv", package = "decorr"),
+nodes <- read.table(system.file("extdata", "nodes.csv", package = "iconr"),
                     sep=";", stringsAsFactors = FALSE)
@@
 edges <- read.table(system.file("extdata", "edges.csv", 
-                                package = "decorr"),
+                                package = "iconr"),
                     sep=";",stringsAsFactors = FALSE)
@@
 \bigskip
-The \pkg{decorr} package functions can be divided into:
+The \pkg{iconr} package functions can be divided into:
 \begin{enumerate}
@@
 
-<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=6, fig.height=4, fig.cap="\\label{fig:igraph}R interface.
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:igraph}R interface.
 Plot of the first graph of the list (Cerro Muriano 1 decoration) with graph drawing on x, y coordinates (left), and force-directed graph drawing with nodes sized on their degree centralities (right)">>=
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 eds.df <- read_eds(site = "Ibahernando",
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 img.graph <- plot_dec_grph(nodes = nds.df,
@@
                            doss = system.file("extdata", 
-                                              package = "decorr"))
+                                              package = "iconr"))
 plot(magick::image_read(img.graph))
@@
                    doss = system.file("extdata", 
-                                      package = "decorr"))
+                                      package = "iconr"))
 l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd)
@@
                              doss = system.file("extdata", 
-                                                package = "decorr"))
+                                                package = "iconr"))
 Ibahernando.img <- magick::image_read(Ibahernando)
@@
                                      doss = system.file("extdata", 
-                                                        package = "decorr"))
+                                                        package = "iconr"))
 Ibahernando.contemp.img <- magick::image_read(Ibahernando.contemp)
@@
 
-The \pkg{decorr} package has two graphical functions: \code{labels_shadow()} and \code{side_plot()}.
+The \pkg{iconr} package has two graphical functions: \code{labels_shadow()} and \code{side_plot()}.
 
@@
 nds.df <- read_nds(site = sit, decor = dec, dev = ".shp",
-                   doss = system.file("extdata", package = "decorr"))
+                   doss = system.file("extdata", package = "iconr"))
 print(xtable::xtable(nds.df[nds.df$id == 6, ],
@@
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
+                   doss = system.file("extdata", package = "iconr"))
 print(xtable::xtable(eds.df[1,],
@@
 nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
+                   doss = system.file("extdata", package = "iconr"))
 eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "decorr"))
+                   doss = system.file("extdata", package = "iconr"))
 img.graph <- plot_dec_grph(nodes = nds.df,
@@
                            doss = system.file("extdata", 
-                                              package = "decorr"),
+                                              package = "iconr"),
                            nd.var = "type", lbl.size = 2)
@@
                           focus = "edges",
-                          doss = system.file("extdata", package = "decorr"))
+                          doss = system.file("extdata", package = "iconr"))
 plot(magick::image_trim(magick::image_read(eds_compar)))
@@
 <<results='asis', echo=TRUE, fig.height= 5>>=
-dataDir <- system.file("extdata", package = "decorr")
+dataDir <- system.file("extdata", package = "iconr")
 nds.df <- read_nds(site = "Ibahernando",
@@
 In Archaeology, the iconography has been considered as a privileged elements to the identification of cultural facies \citep{Renfrew91}.
-By integrating the recording, the management and the analysis process in a simple manner, within the framework of graph Theory and binary topological telationships (\emph{birel}), the \pkg{decorr} framework opens possibilities of studying graphical systems with normalized indexes over large series (Later Paleolithic cave paintings, \emph{cups-and-rings} art, Valcamonica Iron Age \emph{pittoti}, etc.) a long period of time, and on heterogeneous source data.
+By integrating the recording, the management and the analysis process in a simple manner, within the framework of graph Theory and binary topological telationships (\emph{birel}), the \pkg{iconr} framework opens possibilities of studying graphical systems with normalized indexes over large series (Later Paleolithic cave paintings, \emph{cups-and-rings} art, Valcamonica Iron Age \emph{pittoti}, etc.) a long period of time, and on heterogeneous source data.
 \\
-The \pkg{decorr} framework is grounded on a parsimonious model, where Tobler's '\emph{near things}' \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions.
+The \pkg{iconr} framework is grounded on a parsimonious model, where Tobler's '\emph{near things}' \citep{Tobler70} of any iconographical content is recorded with a minimal of \emph{a priori} definitions.
 Those definitions only concern the graph elements: graphical units variables (ie, nodes variables) and types of relations they share (edges).
@@
 
-A part of the IT development of the \pkg{decorr} package has been done in the frame of a postdoctoral position (2015-16) supported by LabEx ARCHIMEDE from "Investissement d'Avenir" program ANR-11-LABX-0032-01.
+A part of the IT development of the \pkg{iconr} package has been done in the frame of a postdoctoral position (2015-16) supported by LabEx ARCHIMEDE from "Investissement d'Avenir" program ANR-11-LABX-0032-01.
 

### 2020-12-18 article_rvTH17.Rnw
article

@@
 
-The notation of with nodes allows also to avoid, empirical and  top-down hierachical typologies .
-According to us, it is more convenient to record GUs (eg, an anthropomorph) and GUs varietes (an anthropomorph wearing an helmet) with different spatialized nodes (firstly within a GIS) rather than multiplying hierarchical levels of categories (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious idiosynchratic inventories and difficult possible comparisons \emph{within} and \emph{between} these categories.
+The notation of with nodes allows also to avoid, empirical and  top-down hierachical typologies.
+According to us, it is more convenient to record GUs (eg, an anthropomorph) and GUs varietes (an anthropomorph wearing an helmet) with discrete spatialized nodes (firstly within a GIS) rather than multiplying a-spatial levels of hierarchical categories (e.g., a category of graphical units with \emph{n} types, a type with \emph{n} subtypes, a subtype with \emph{n} varieties, etc.) that lead also to tedious idiosynchratic inventories and difficult possible comparisons \emph{within} and \emph{between} these categories.
 For example, with the two types of GUs \code{casque} (helmet) and \code{personnage} (anthropomorphe) and the \emph{attribute} edge it is possible to describe: (1) a case where an anthropomorph wears a helmet (like for the Cerro Muriano steale, Figure \ref{fig:dec.1.4} left), (2) a case where an anthropomorph is separated from a helmet (like for the Zarza de Montanchez, Figure \ref{fig:dec.1.4} right).

### 2021-01-19 article_rvTH18.Rnw
up

--- 2020-12-18 article_rvTH17.Rnw
@@
 \renewcommand{\floatpagefraction}{.8}
+%\definecolor{shadecolor}{rgb}{1, 1, 1}
+\definecolor{fgcolor}{rgb}{0, 0, 0}
+\makeatletter
+\def\verbatim@font{\ttfamily\slshape}
+\makeatother
 
@@
 library("igraph")
-opts_chunk$set(prompt=TRUE, fig.pos = '!htbp', results='asis', highlight=FALSE,
+knitr::opts_chunk$set(prompt=TRUE, fig.pos = '!htbp',
+                      background = 'white', highlight=FALSE,
                       class.source = "bg-success")
@@
 For example, the 3-letters word "\code{art}" can be modelled with three vertices (\code{a}, \code{r}, \code{t}) and two edges (one between \code{a} and \code{r}, the other between \code{r} and \code{t}).
-In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Figure \ref{fig:fart}).
-
-<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.25, fig.height=0.25, fig.cap="\\label{fig:fart}Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
+In \proglang{R}, these features, concatenated in this order with a \code{paste0()}, is \code{art}, and not \code{rat} (Figure \ref{fig:a_r_t}).
+
+<<a_r_t, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.25, fig.height=0.25, fig.cap="Concatenate of 'a', 'r' and 't' graphical units is 'art'">>=
 par(mar=c(0.1,0.1,0.1,0.1) )
@@
 
-<<echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.4, fig.height=1.4, fig.cap="\\label{fig:fratar}Potential spatial relations between graphical units">>=
+<<fratar, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=1.4, fig.height=1.4, fig.cap="Potential spatial relations between graphical units">>=
 par(mar=c(0,0,0,0))
@@
 \item consistency, proximities and relationships between these groups are often implicit and not quantified
-\item studies develop proper descriptive vocabularies, singular relationships of grouping, idosyncratic methods at site-dependent or period-dependent scales
+\item studies develop proper descriptive vocabularies, singular relationships of grouping, ido\-syn\-crat\-ic methods at site-dependent or period-dependent scales
 \end{itemize}
@@
 Whether this theoretical process includes image binarization, polygonization and \emph{birel} analysis, a simplier solution will be to create directly centroids on the GUs and to draw the edges manually.
-Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.\\
+Then \emph{x} and \emph{y} coordinates of nodes can be easily retrieved.
 
@@
 
-<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.5630631, fig.cap="\\label{fig:dec.1.4} Cerro Muriano 1 and Zarza de Montanchez decorations.">>=
+<<dec.1.4, echo=FALSE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=0.5, fig.cap="Cerro Muriano 1 and Zarza de Montanchez decorations.">>=
 library(iconr)
-par(mar=c(0, 0, 0, 0))
-nds.df <- read_nds(site = "Cerro Muriano", 
-                   decor = "Cerro Muriano 1",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-eds.df <- read_eds(site = "Cerro Muriano",
-                   decor = "Cerro Muriano 1",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-cm1 <- plot_dec_grph(nodes = nds.df,
-                     edges = eds.df,
-                     site = "Cerro Muriano",
-                     decor = "Cerro Muriano 1",
+imgs <- read.table(system.file("extdata", "imgs.tsv", package = "iconr"),
+                   sep="\t", stringsAsFactors = FALSE)
+sites = c("Cerro Muriano", "Zarza de Montanchez")
+decors = c("Cerro Muriano 1", "Zarza De Montanchez")
+par(mfrow = c(1, 2))
+for(i in 1:2) {
+  nds.df <- read_nds(sites[i], decors[i],
+                     dir = system.file("extdata", package = "iconr"))
+  eds.df <- read_eds(sites[i], decors[i],
+                     dir = system.file("extdata", package = "iconr"))
+  plot_dec_grph(nds.df, eds.df, imgs,
+                sites[i], decors[i],
                 nd.var = "type",
-                     lbl.size = 1.7,
-                     doss = system.file("extdata", 
-                                        package = "iconr"))
-cm1.img <- magick::image_read(cm1)
-nds.df <- read_nds(site = "Zarza de Montanchez", 
-                   decor = "Zarza De Montanchez",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-eds.df <- read_eds(site = "Zarza de Montanchez",
-                   decor = "Zarza De Montanchez",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-zm <- plot_dec_grph(nodes = nds.df,
-                            edges = eds.df,
-                            site = "Zarza de Montanchez",
-                            decor = "Zarza De Montanchez",
-                            nd.var = "type",
-                            lbl.size = 1.7,
-                            doss = system.file("extdata", 
-                                               package = "iconr"))
-zm.img <- magick::image_read(zm)
-out.img <- magick::image_append(c(cm1.img, zm.img))
-plot(out.img)
+                dir = system.file("extdata", package = "iconr"),
+                img.format = NULL)
+}
 @
@@
 
-<<results='asis', include=FALSE, echo=FALSE, warning=FALSE>>=
+<<include=FALSE, echo=FALSE, warning=FALSE>>=
 library("iconr")
@@
 %
-<<results='asis', echo=TRUE, warning=FALSE>>=
+<<echo=TRUE, warning=FALSE>>=
 imgs <- read.table(system.file("extdata", "imgs.tsv", package = "iconr"),
@@
 %
-<<timg, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.7507508, fig.cap="\\label{fig:dataset}Decoration images of the training dataset">>=
-pth <- system.file("extdata", package = "iconr")
-lims <- list()
+<<dataset, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=15, fig.asp=0.7507508, fig.cap="Decoration images of the training dataset">>=
+ims <- list()
 for(i in 1:nrow(imgs)){
-  i1 <- image_read(paste0(pth,"/",imgs$img[i]))
+  im <- image_read(system.file("extdata", imgs$img[i], package = "iconr"))
   lbl.txt <- paste0(paste(imgs[i,], collapse = "\n"), "\n",
-                    image_info(i1)$width, "*", image_info(i1)$height, " px")
-  lims[[i]] <- image_annotate(i1, lbl.txt, location = "northwest",
+                    image_info(im)$width, "*", image_info(im)$height,
+                    " px")
+  ims[[i]] <- image_annotate(im, lbl.txt, location = "northwest",
                              size = 28, color = "red")
 }
-out.img <- image_append(c(image_append(c(lims[[1]], lims[[2]], lims[[3]])),
-                          image_append(c(lims[[4]], lims[[5]]))),
+out.img <- image_append(c(image_append(c(ims[[1]], ims[[2]], ims[[3]])),
+                          image_append(c(ims[[4]], ims[[5]]))),
                         stack = TRUE)
@@
 
-<<results='asis', echo=TRUE>>=
-nodes <- read.table(system.file("extdata", "nodes.csv", package = "iconr"),
+<<echo=TRUE>>=
+nodes <- read.table(system.file("extdata", "nodes.csv",
+                                package = "iconr"),
                     sep=";", stringsAsFactors = FALSE)
@@
 Analogously, the edges for the five decorations are stored in different formats in \code{edges.csv}, \code{edges.tsv}, and \code{edges.shp}. This can be read by
-<<results='asis', echo=TRUE>>=
+<<echo=TRUE>>=
 edges <- read.table(system.file("extdata", "edges.csv",
@@
       include.rownames = FALSE,
-      scalebox = 0.95,
-      size = "\\renewcommand{\\arraystretch}{0.92}",
+#      scalebox = 0.95,
+#      size = "\\renewcommand{\\arraystretch}{0.92}",
 #      size = "\\small\\renewcommand{\\arraystretch}{0.989}",
@@
 This is our case since we work with geometric graphs.
-For a same graph, very different layouts (\emph{réalisations}) exist \citep{Mathis03}(\emp{graph drawing}) (Figure \ref{fig:igraph})
-
-<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=4, fig.height=4, fig.cap="\\label{fig:igraph}R interface.
+For a same graph, very different layouts (\emph{réalisations}) exist \citep{Mathis03}(\emp{graph drawing}) (Figure \ref{fig:igraphs})
+
+<<igraphs, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.asp=0.5, fig.height=4, fig.cap="R interface.
 Plot of the first graph of the list (Cerro Muriano 1 decoration) with graph drawing on x, y coordinates (left), and force-directed graph drawing with nodes sized on their degree centralities (right)">>=
@@
 par(mar = c(0, 0, 0, 0), mfrow = c(1, 2))
-coords <- layout.fruchterman.reingold(g)
 plot(g,
@@
 plot(g,
-     layout = coords,
+     layout = layout.fruchterman.reingold(g),
      vertex.size = 4 + (degree(g)*10),
@@
 
-<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.5, fig.asp=1.126126, fig.cap="\\label{fig:ibah}Ibahernando decoration, node 1 overlaps node 2 and node 3">>=
+<<ibah, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.2, fig.asp=1.126126, fig.cap="Ibahernando decoration{:} node 1 overlaps node 2 and node 3.">>=
+site = "Ibahernando"
+decor = "Ibahernando"
+dataDir = system.file("extdata", package = "iconr")
+nds.df <- read_nds(site, decor, dataDir)
+eds.df <- read_eds(site, decor, dataDir)
 par(mar=c(0,0,0,0))
-nds.df <- read_nds(site = "Ibahernando", 
-                   decor = "Ibahernando",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-eds.df <- read_eds(site = "Ibahernando",
-                   decor = "Ibahernando",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-img.graph <- plot_dec_grph(nodes = nds.df,
-                           edges = eds.df,
-                           site = "Ibahernando",
-                           decor = "Ibahernando",
-                           lbl.size = 2,
-                           doss = system.file("extdata", 
-                                              package = "iconr"))
-plot(magick::image_read(img.graph))
+plot_dec_grph(nds.df, eds.df, imgs,
+              site, decor, dir = dataDir,
+              lbl.size = 0.7, img.format = NULL)
 @
@@
 
-<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=0.5630631, fig.cap="\\label{fig:ibah2}Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
-par(mar=c(0,0,0,0))
-selected.nd <- 4
-nds.df <- read_nds(site = "Ibahernando", 
-                   decor = "Ibahernando",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-eds.df <- read_eds(site = "Ibahernando",
-                   decor = "Ibahernando",
-                   doss = system.file("extdata", 
-                                      package = "iconr"))
-l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd)
-Ibahernando <- plot_dec_grph(nodes = nds.df,
-                             edges = eds.df,
-                             site = "Ibahernando",
-                             decor = "Ibahernando",
-                             nd.var = "type",
-                             lbl.color = "brown",
-                             lbl.size=2.2,
-                             doss = system.file("extdata", 
-                                                package = "iconr"))
-Ibahernando.img <- magick::image_read(Ibahernando)
-Ibahernando.contemp <- plot_dec_grph(nodes = l_dec_df[[1]],
-                                     edges = l_dec_df[[2]],
-                                     site = "Ibahernando",
-                                     decor = "Ibahernando",
-                                     nd.var = "type",
-                                     lbl.color = "brown",
-                                     lbl.size=2.2,
-                                     doss = system.file("extdata", 
-                                                        package = "iconr"))
-Ibahernando.contemp.img <- magick::image_read(Ibahernando.contemp)
-out <- magick::image_append(c(Ibahernando.img, Ibahernando.contemp.img))
-plot(out)
+%<<ibah2, echo=TRUE, message=FALSE, warning=FALSE>>=
+%nds.df <- read_nds(site = "Ibahernando",
+%                   decor = "Ibahernando",
+%                   dir = system.file("extdata", package = "iconr"))
+%eds.df <- read_eds(site = "Ibahernando",
+%                   decor = "Ibahernando",
+%                   dir = system.file("extdata", package = "iconr"))
+%l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 4)
+%l.nodes = list(nds.df, l_dec_df[[1]])
+%l.edges = list(eds.df, l_dec_df[[2]])
+%out.file.name = paste0("figure/ibah2_", c("a", "b"), ".pdf")
+%img.files <- character(0)
+%for(i in 1:2) {
+%  img.files[i] <- plot_dec_grph(nodes = l.nodes[[i]],
+%                                  edges = l.edges[[i]],
+%                                  site = "Ibahernando",
+%                                  decor = "Ibahernando",
+%                                  nd.var = "type",
+%                                  lbl.color = "brown",
+%                                  lbl.size = 0.7,
+%                                  dir = system.file("extdata",
+%                                                     package = "iconr"),
+%#                                  img.format = "pdf")
+%                                  out.file.name = out.file.name[i])
+%}
+%@
+%\begin{figure}[!htbp]
+%  \centering
+%  \centering
+%  \includegraphics[width=0.48\textwidth]{\Sexpr{img.files[1]}}\quad
+%  \includegraphics[width=0.48\textwidth]{\Sexpr{img.files[2]}}
+%  \caption{Ibahernando stelae before and after the selection of node 4 (sword) graph component.}
+%  \label{fig:ibah2}
+%\end{figure}
+
+<<ibah2, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8, fig.asp=0.5, fig.cap="Ibahernando stelae before and after the selection of node 4 (sword) graph component">>=
+site = "Ibahernando"
+decor = "Ibahernando"
+dataDir = system.file("extdata", package = "iconr")
+nds.df <- read_nds(site, decor, dataDir)
+eds.df <- read_eds(site, decor, dataDir)
+l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 4)
+par(mfrow = c(1, 2), mar = c(0, 0, 0, 0))
+plot_dec_grph(nds.df, eds.df, imgs,
+              site, decor, dir = dataDir,
+              nd.var = "type", lbl.color = "brown", lbl.size = 0.7,
+              img.format = NULL)
+plot_dec_grph(l_dec_df$nodes, l_dec_df$edges, imgs,
+              site, decor, dir = dataDir,
+              nd.var = "type", lbl.color = "brown", lbl.size = 0.7,
+              img.format = NULL)
 @
@@
 
-\begin{CodeChunk}
-\begin{CodeInput}
-R> lgrph <- list_dec(imgs, nodes, edges)
-R> named_elements(lgrph[[2]], focus = "edges", nd.var="type")
-## [1] "epee-=-fibule" "bouclier-=-miroir"
-## [3] "epee-=-lance" "bouclier-=-chariot_char"
-## [5] "chariot_char-+-cheval" "chariot_char-+-cheval#"
-## [7] "arc-=-bouclier" "arc-=-chariot_char"
-## [9] "arc-=-miroir" "bouclier-=-fibule"
-## [11] "bouclier-=-epee" "epee-=-miroir"
-## [13] "arc-+-fleche" "chariot_char-+-roue"
-## [15] "chariot_char-+-roue#"
-R> as.character(named_elements(lgrph[[2]], focus = "nodes", nd.var="type"))
-## [1] "chariot_char" "bouclier" "arc" "miroir"
-## [5] "fibule" "epee" "lance" "cheval"
-## [9] "cheval#" "roue" "roue#" "fleche"
-\end{CodeInput}
-\end{CodeChunk}
+<<named_elements, echo=TRUE>>=
+lgrph <- list_dec(imgs, nodes, edges)
+named_elements(lgrph[[2]], focus = "edges", nd.var="type")
+named_elements(lgrph[[2]], focus = "nodes", nd.var="type")
+@
 
@@
 \begin{Code}
-side_plot(grp, doss, nd.var, focus = "nodes",
+side_plot(grp, dir, nd.var, focus = "nodes",
           nd.color = c("orange", "red"),
@@
 \code{side_plot()} allows to plot two decoration graphs side-by-side for elements (nodes or edges) comparisons.
-It take a list of two graph (\code{grp}), the path of the working directory (\code{doss}) and the name of the studied node variable (\code{nd.var}) and return an image.
+It take a list of two graph (\code{grp}), the path of the working directory (\code{dir}) and the name of the studied node variable (\code{nd.var}) and return an image.
 
@@
          decor,
-         doss = getwd(),
+         dir = getwd(),
          nodes = "nodes",
@@
 \code{read_nds()} is close to the \proglang{R} native \code{read.table()} function, but allows to read \emph{shapefiles} of nodes.
-For example, the nodes of the Torrejon Rubio 1 decoration can be read from the the \emph{shapefile}, and the 6\textsuperscript{th} node -- a sword (\code{epee}) -- displayed (Table \ref{tab:nd}).
-
-<<nd, results='asis', echo=TRUE>>=
-caption <- "Torrejon Rubio 1 decoration node 6: the sword."
-sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-nds.df <- read_nds(site = sit, decor = dec, dev = ".shp",
-                   doss = system.file("extdata", package = "iconr"))
-print(xtable::xtable(nds.df[nds.df$id == 6, ],
-                     caption = caption,
+For example, the nodes of the Torrejon Rubio 1 decoration can be read from the \emph{shapefile} (Table \ref{tab:nd} displays two nodes as example).
+
+<<nd0, echo=TRUE>>=
+nds.df <- read_nds(site = "Torrejon Rubio",
+                   decor = "Torrejon Rubio 1",
+                   dir = system.file("extdata", package = "iconr"),
+                   dev = ".shp")
+@
+<<nd, results='asis', echo=FALSE>>=
+print(xtable::xtable(nds.df[5:6, ],
+                     caption = "Torrejon Rubio 1 decoration nodes 5 (fibula) and 6 (sword).",
                      label = "tab:nd",
@@
          decor,
-         doss = getwd(),
+         dir = getwd(),
          edges = "edges",
@@
 \code{read_eds()} permits to read a \code{.csv} or \code{.tsv} dataframe, or a \emph{shapefiles}, of edges.
-When the data source is a dataframe, the coordinates of the edges are retrieved from the nodes dataframe (Table \ref{tab:ed}).
-
-<<ed, results='asis', echo=TRUE >>=
-caption <- paste0("First edge of the Torrejon Rubio 1 decoration, ",
-                  "the proximity between the sword (node 6) ",
-                  "and the fibula (node 5)")
-eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "iconr"))
-print(xtable::xtable(eds.df[1,],
-                     caption = caption,
+When the data source is a dataframe, the coordinates of the edges are retrieved from the nodes dataframe (Table \ref{tab:ed} displays the first edge as example).
+
+<<ed0, echo=TRUE >>=
+eds.df <- read_eds(site = "Torrejon Rubio",
+                   decor = "Torrejon Rubio 1",
+                   dir = system.file("extdata", package = "iconr"),
+                   dev = ".tsv")
+@
+<<ed, results='asis', echo=FALSE >>=
+caption <- paste0()
+print(xtable::xtable(eds.df[1, ],
+                     caption = "Torrejon Rubio 1 decoration first edge, representing the proximity between the sword (node 6) and the fibula (node 5)",
                      label = "tab:ed",
@@
               edges = NULL,
+              imgs,
               site,
               decor,
-              doss = getwd(),
+              dir = getwd(),
               nd.var = 'id',
-              nd.color = 'orange', nd.size = 1.7,
-              lbl.color = 'black', lbl.size = 1.2,
-              ed.color = c("orange", "blue"), ed.lwd = 4,
-              img.format = "png")
+              nd.color = 'orange', nd.size = 0.5,
+              lbl.color = 'black', lbl.size = 0.5,
+              ed.color = c("orange", "blue"), ed.lwd = 1,
+              dir.out = dir, out.file.name = NULL,
+              img.format = "png", res = 300)
 \end{Code}
@@
 
-<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.5, fig.asp=1.126126, fig.cap="\\label{fig:trst}Torrejon Rubio 1 decoration">>=
+<<trst, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=3.2, fig.asp=1.126126, fig.cap="Torrejon Rubio 1 decoration">>=
+site <- "Torrejon Rubio"
+decor <- "Torrejon Rubio 1"
+dataDir <- system.file("extdata", package = "iconr")
+nds.df <- read_nds(site, decor, dataDir, dev = ".tsv")
+eds.df <- read_eds(site, decor, dataDir, dev = ".tsv")
 par(mar=c(0, 0, 0, 0))
-sit <- "Torrejon Rubio" ; dec <- "Torrejon Rubio 1"
-nds.df <- read_nds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "iconr"))
-eds.df <- read_eds(site = sit, decor = dec, dev = ".tsv",
-                   doss = system.file("extdata", package = "iconr"))
-img.graph <- plot_dec_grph(nodes = nds.df,
-                           edges = eds.df,
-                           site = sit,
-                           decor = dec,
-                           doss = system.file("extdata", 
-                                              package = "iconr"),
-                           nd.var = "type", lbl.size = 2)
-out <- magick::image_read(img.graph)
-plot(out)
-@
-
-\subsection{Decoration comparison functions} \label{sec:functions_one}
+plot_dec_grph(nds.df, eds.df, imgs,
+              site, decor, dataDir,
+              nd.var = "type", lbl.size = 0.55,
+              img.format = NULL)
+@
+
+\subsection{Decoration comparison functions} \label{sec:functions_comp}
 
@@
 plot_compar(listg, graph2 = NULL, focus = "nodes",
-            doss = getwd(),
+            dir = getwd(),
             nd.color = c("orange", "red"), nd.size = c(0.5, 1),
             ed.color = c("orange", "red"), ed.width = c(1, 2),
-            lbl.size = 0.5, img.format = "png", res = 300)
+            lbl.size = 0.5,
+            dir.out = dir, out.file.name = NULL,
+            img.format = "png", res = 300)
 \end{Code}
@@
 
-<<echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.asp=0.5630631, fig.cap="\\label{fig:ceds}Comparison between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
-par(mar=c(0,0,0,0))
+%<<ceds, echo=TRUE, message=FALSE, warning=FALSE>>=
+%g.compar <- list_compar(lgrph, "type")
+%eds_compar <- plot_compar(g.compar, c(1, 4),
+%                          focus = "edges",
+%                          dir = system.file("extdata", package = "iconr"),
+%                          img.format = "pdf")
+%@
+%\begin{figure}[!htbp]
+%  \centering
+%  \includegraphics[width=\textwidth]{\Sexpr{eds_compar}}
+%  \caption{Comparison between decoration 1 (Cerro Muriano 1) and decoration 4
+%           (Zarza de Montsanchez)}
+%  \label{fig:ceds}
+%\end{figure}
+
+<<ceds, echo=TRUE, message=FALSE, warning=FALSE, fig.pos = '!htbp', fig.align="center", fig.width=8 , fig.asp=0.5, fig.cap="Comparison between decoration 1 (Cerro Muriano 1) and decoration 4 (Zarza de Montsanchez)">>=
 g.compar <- list_compar(lgrph, "type")
-eds_compar <- plot_compar(g.compar, c(1, 4),
+plot_compar(g.compar, c(1, 4),
             focus = "edges",
-                          doss = system.file("extdata", package = "iconr"))
-plot(magick::image_trim(magick::image_read(eds_compar)))
+            dir = system.file("extdata", package = "iconr"),
+            img.format = NULL)
 @
@@
 
-<<results='asis', echo=TRUE>>=
+<<echo=TRUE>>=
 df.same_edges <- same_elements(lgrph, "type", "edges")
+@
+<<results='asis', echo=FALSE>>=
 caption <- "Number of common edges between all graph decoration pairwises"
@@
 
-<<results='asis', echo=TRUE, fig.height= 5>>=
+<<echo=TRUE>>=
 dataDir <- system.file("extdata", package = "iconr")
-nds.df <- read_nds(site = "Ibahernando",
-                   decor = "Ibahernando",
-                   doss = dataDir)
-eds.df <- read_eds(site = "Ibahernando",
-                   decor = "Ibahernando",
-                   doss = dataDir)
+imgCase <- imgs[5, ]
+nds.df <- read_nds(imgCase$site, imgCase$decor, dataDir)
+eds.df <- read_eds(imgCase$site, imgCase$decor, dataDir)
 l_dec_df <- contemp_nds(nds.df, eds.df, selected.nd = 2)
-l_dec_df[[1]]$site <- l_dec_df[[1]]$decor <- NULL
-l_dec_df[[2]]$site <- l_dec_df[[2]]$decor <- NULL
-grph <- graph_from_data_frame(l_dec_df[[2]], vertices = l_dec_df[[1]],
-                           directed = FALSE)
-grph$site <- grph$decor <- "Ibahernando"
-grph$name <- 5
-lgrph[[5]] <- grph
-@
-
-Then, \code{same_elements()} is done on nodes and edges (Table \ref{tab:common.elts})
-
-<<results='asis', echo=TRUE>>=
+lgrph[5] <- list_dec(imgCase, l_dec_df$nodes, l_dec_df$edges)
+@
+
+Then, \code{same_elements()} can be applied on nodes and edges (Table \ref{tab:common.elts})
+
+<<echo=TRUE>>=
 df.same_edges <- same_elements(lgrph, "type", "edges")
@@
 \textbf{Common nodes}\raisebox{-1ex}{}\\
-<<echo=FALSE,results='asis' >>=
+<<echo=FALSE, results='asis'>>=
 print(xtable::xtable(df.same_nodes,
@@
 \textbf{Common edges}\raisebox{-1ex}{}\\
-<<echo=FALSE,results='asis'>>=
+<<echo=FALSE, results='asis'>>=
 print(xtable::xtable(df.same_edges,
@@
 Once the heatmap matrices calculated, the distance matrices are calculated with the native \code{dist()} function \citep{R19} (Table \ref{tab:dist.elts}).
+
+<<echo=TRUE>>=
+dist.nodes  <- dist(df.same_nodes)
+dist.edges <- dist(df.same_edges)
+@
 
@@
 \textbf{Distance matrix on nodes}\raisebox{-1ex}{}\\
-<<echo=FALSE,results='asis', warning=F >>=
+<<echo=FALSE, results='asis', warning=FALSE>>=
 library("matlib")
-dist.nodes  <- dist(as.matrix(df.same_nodes), upper=TRUE)
 mat.nodes <- as.matrix(dist.nodes)
@@
 \textbf{Distance matrix on edges}\raisebox{-1ex}{}\\
-<<echo=FALSE,results='asis'>>=
-dist.edges <- dist(as.matrix(df.same_edges))
+<<echo=FALSE, results='asis'>>=
 mat.edges <- as.matrix(dist.edges)
@@
 
-<<hclust, echo=TRUE, warning=FALSE, fig.height=4, fig.width=8, fig.pos = '!htbp', fig.cap="\\label{fig:hclust}Hierachical clusterings on decoration common elements">>=
-par(mfrow=c(1, 2), mar=c(0.2,4,1.2,0.5))
+<<hclust, echo=TRUE, warning=FALSE, fig.height=4, fig.width=8, fig.pos = '!htbp', fig.cap="Hierachical clusterings on decoration common elements">>=
 hc.nds <- hclust(dist.nodes, method = "ward.D")
 hc.eds <- hclust(dist.edges, method = "ward.D")
-plot(hc.nds, main = "common nodes", cex = 1.1)
-plot(hc.eds, main = "common edges", cex = 1.1)
+par(mfrow=c(1, 2), mar=c(0.2,4,1.2,0.5))
+plot(hc.nds, main = "Common nodes", cex = 1.1)
+plot(hc.eds, main = "Common edges", cex = 1.1)
 @
@@
 
-<<hclust2, echo=TRUE, warning=FALSE, fig.height=3, fig.pos = '!htbp', fig.cap="\\label{fig:hclust2}Comparison of hierachical clusterings on decoration common elements, nodes (left) and edges (right)">>=
-suppressPackageStartupMessages(library(dendextend))
-suppressPackageStartupMessages(library(dplyr))
-par(mfrow=c(1, 2))
-dend.nds <- as.dendrogram (hc.nds)
-dend.eds <- as.dendrogram (hc.eds)
+<<hclust2, echo=TRUE, warning=FALSE, message=FALSE, fig.height=3, fig.pos = '!htbp', fig.cap="Comparison of hierachical clusterings on decoration common elements, nodes (left) and edges (right)">>=
+library(dendextend)
+library(dplyr)
+dend.nds <- as.dendrogram(hc.nds)
+dend.eds <- as.dendrogram(hc.eds)
 dendlist(dend.nds, dend.eds) %>%
@@
   tanglegram(columns_width = c(6, 1, 6),
-             main_left = "common nodes",
-             main_right = "common edges",
+             main_left = "Common nodes",
+             main_right = "Common edges",
              lab.cex = 1.5,
              cex_main = 1.5,
-             highlight_branches_lwd = F)                    
+             highlight_branches_lwd = FALSE)
 @
